### 2023-05-02 Notes/efficiency_twophase.tex
Wrote up efficiency comparison with regression estimators. Our methods seem to outperform them.

\section*{Efficiency of Proposed Estimator}

One of our main goals is to show that we have an efficient estimator. We
want to ensure that our estimator is more efficient (lower MSE or really zero
bias and a smaller variance) than other competing estimators. We have already
demonstrated superior performance compared to the IPW estimator with known
weights. Now, we want to compare our estimator with two-phase and three-phase 
regression estimators.

\subsection*{Two and Three Phase Regression Estimators}

\begin{itemize}
  \item The simulation setup that we have been running generates monotone and
    nonmonotone missing patterns because we always observe $X$, yet we observe
    $Y_1$ and $Y_2$ with missingness. This creates the need to extend the
    traditional two-phase estimator to become a three-phase estimator. The
    two-phase regression estimator is really just a regression estimator with
    the ``finite population'' being Phase 1 of the sample. Thus, a valid
    two-phase estimator for $\theta = E[Y_2]$ is

    \[\hat \theta = \bar y_2 + (\bar x_1 - \bar x_2) \hat \beta\]

    where $\bar y_2 = n_2^{-1} \sum_{i \in U} I(i \in A_2) y_2$ and $\bar x_k =
    n_k^{-1} \sum_{i \in U} I(i \in A_k) x_i$ where $n_k$ is the number of
    elements in $A_k$ where $A_k$ is the Phase $k$ sample. In this case, $\hat
    \beta$ solves the following equation for $\beta_0$ and $\beta_1$,

    \[\sum_{i \in A_2} (y_{2i} - \beta_0 - \beta_1 x_i)^2 = 0.\]

  \item Notice that the previous construction of the two-phase estimator ignored
    the variable $Y_1$. We can incorporate this into the model using a 
    three-phase estimator. From \cite{fuller2009sampling}, the three-phase
    estimator is

    \begin{align*}
      \bar y_{2, 2p} &= \bar y_2 + (\bar x_0 - \bar x_{2}) \hat \beta_1 + 
      (\bar y_{1, reg} - \bar y_{1, 2p}) \hat \beta_2
    \end{align*}

    where 
    \begin{align*}
      \bar x_0 &= n^{-1} \sum_{i \in U} x_i \\
      \bar x_1 &= \left(\sum_{i \in A_1} \pi_{1i}^{-1}\right)^{-1} 
                  \sum_{i \in A_1} \pi_{1i}^{-1} x_i\\
      \bar x_2 &= \left(\sum_{i \in A_2} \pi_{2i}^{-1}\right)^{-1} 
                  \sum_{i \in A_2} \pi_{2i}^{-1} x_i\\
      \bar y_{1, reg} &= \bar y_{1, 1p} + (\bar x_0 - \bar x_1)\hat \beta_{1p}\\
      \bar y_{1, 1p} &= \left(\sum_{i \in A_1} \pi_{1i}^{-1}\right)^{-1} 
                        \sum_{i \in A_1} \pi_{1i}^{-1} y_{1i}\\
      \bar y_{1, 2p} &= \left(\sum_{i \in A_2} \pi_{2i}^{-1}\right)^{-1} 
                        \sum_{i \in A_2} \pi_{2i}^{-1} y_{1i}\\
      \hat \beta_{1p} &= \left(\sum_{i \in A_1} (x_i - \bar x_1)^2
      \pi_{1i}^{-1}\right)^{-1} \sum_{i \in A_1} (x_i - \bar x_1)\pi_{1i}^{-1}
      (y_{1i} - \bar y_{1, 1p})\\
      \hat \beta_1 &= \left(\sum_{i \in A_2} (x_i - \bar x_2, y_{1i} - \bar
      y_{1, 2p})' \pi_{2i}^{-1} (x_i - \bar x_2, y_{1i} - \bar y_{1,
      2p})\right)^{-1} \sum_{i \in A_2} (x_i - \bar x_2) \pi_{2i}^{-1} (y_{2i} -
      \bar y_2)\\
      \hat \beta_2 &= \left(\sum_{i \in A_2} (x_i - \bar x_2, y_{1i} - \bar
      y_{1, 2p})' \pi_{2i}^{-1} (x_i - \bar x_2, y_{1i} - \bar y_{1,
      2p})\right)^{-1} \sum_{i \in A_2} (y_{1i} - \bar y_{1, 2p}) \pi_{2i}^{-1} 
      (y_{2i} - \bar y_2)\\
    \end{align*}

  \item Notice that the three-phase estimator implicitly assumes a monotone 
    missingness model. It ignores the values of $y_2$ that have an unobserved
    $y_1$.

\end{itemize}

\subsection*{Monotone Results}

\begin{itemize}
  \item First, we test the two-phase and three-phase regression estimators with
    a monotone simulation. Like the previous monotone simulations,
    we first generate data from the following distributions:
    \begin{align*}
        X_i &\stackrel{iid}{\sim} N(0, 1) \\
        Y_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
        Y_{2i} &\stackrel{iid}{\sim} N(\theta, 1)
    \end{align*}

    Then, we create the probabilities $p_1 = \logistic(x_i)$ and 
    $p_{12} = \logistic(y_{1i})$. Including the calibration estimator from the
    previous section (note, these are the same as the previous monotone tables)
    this yields,

    \input{Tables/calimono_t-5.tex}
    \input{Tables/calimono_t0.tex}
    \input{Tables/calimono_t5.tex}

  \item The proposed semiparametric estimator and the calibration estimator
    seem to both outperform the regression estimators. With smaller bias and
    smaller variance our estimators do better.

  \item It may be puzzling to see that the three-phase estimator does worse than
    the two-phase; however, I think that there is a good reason for this. In the
    simulation, $y_2 = x + \varepsilon$. This is the regression from the
    two-phase estimator. However, the three-phase estimator measures $y_2 \sim x
    + y_1 + \varepsilon$, which just adds noise. This is why I think the
    standard deviation of the three-phase regression estimator is larger.

\end{itemize}

\subsection*{Nonmonotone Results}

\begin{itemize}
  \item Similar to the monotone results, we also use the same simulation as the
    nonmonotone calibration estimators. Repeating the simulation outline, we are
    trying to estimate $\theta = E[y_2]$ and we have

    \begin{enumerate}
      \item Generate $X_i$, $\varepsilon_{1i}$, and $\varepsilon_{2i}$ from the
        following distributions:

        \begin{align*}
          x_i &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{1i} &\stackrel{iid}{\sim} N(0, 1)\\
          \varepsilon_{2i} &\stackrel{iid}{\sim} N(\theta, 1)\\
        \end{align*}

        Then we have
        \[y_{1i} = x_i + \varepsilon_{1i} \text{ and } y_{2i} = x_i +
        \varepsilon_{2i}.\]

      \item Then we have to select the variables to observe. We always observe
        $X_i$. Then we choose to either observe $Y_1$ with probability $0.4$,
        $Y_2$ with probability $0.4$ or neither with probability $0.2$.

      \item If neither then $R_{1i} = 0$ and $R_{2i} = 0$. Otherwise, if we 
        observe $Y_1$ then $R_1 = 1$ and if we observe $Y_2$ then $R_2 = 1$.

      \item If we observe either $Y_1$ or $Y_2$ then with probability $p \propto
        \logistic(Y_k)$ where $Y_k$ is the observed $Y$ variable we choose to
        observe the other $Y$ variable.

      \item If the other $Y$ variable is observed then the corresponding $R_k =
        1$. Otherwise, $R_k = 0$.
    \end{enumerate}

    \input{Tables/calinonmono1m-5.tex}
    \input{Tables/calinonmono1m0.tex}
    \input{Tables/calinonmono1m5.tex}

  \item Overall, it seems that the proposed estimator and calibration estimator
    outperform the two regression estimators. The regression estimators display
    slight bias (perhaps because of the nonmonotonicity).

\end{itemize}

\newpage