### 2023-11-07 Notes/optest_update.tex
Added simulation 2.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
\usepackage{biblatex}
\usepackage{booktabs}
\usepackage{float}

\addbibresource{references.bib}

\hypersetup{colorlinks = true, urlcolor = black}

\headheight     15pt
\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\argmax}{{\text{argmax}}}
\newcommand{\argmin}{{\text{argmin}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\renewcommand{\arraystretch}{1.2}

\begin{document}

\lhead{Caleb Leedy}
\chead{Non-monotone Missingness}
%\chead{STAT 615 - Advanced Bayesian Methods}
%\rhead{Page \thepage\ of \pageref{LastPage}}
\rhead{\today}

\section*{Setup}

Consider the following setup. Let $(X, Y_1, Y_2, \delta)  \stackrel{ind}{\sim}
F$ for some distribution $F$ that is unknown. We define $Z = (X, Y_1, Y_2$ and
we observe the following:

\begin{table}[!ht]
  \centering
  \caption{This table shows some of our notation and some of the corresponding
  notation from \cite{tsiatis2006semiparametric}.}
\begin{tabular}{cccccccc}
  \toprule
  Segments & $X$ & $Y_1$ & $Y_2$ & Prob. Element in Segment & $\delta$ & $C$ & $G_C(Z)$ \\
  \midrule
  $A_{00}$ & $\checkmark$ &               &         & $\pi_{00}$ & $\delta_{00}$ & 1 & $\{X\}$ \\
  $A_{10}$ & $\checkmark$ & $\checkmark$  &         & $\pi_{10}$ & $\delta_{10}$ & 2 & $\{X, Y_1\}$ \\
  $A_{01}$ & $\checkmark$ &               & $\checkmark$ & $\pi_{01}$ & $\delta_{01}$ & 3 & $\{X, Y_2\}$ \\
  $A_{11}$ & $\checkmark$ & $\checkmark$  & $\checkmark$ & $\pi_{11}$ & $\delta_{11}$ & $\infty$ & $\{X, Y_1, Y_2\}$ \\
  \bottomrule
\end{tabular}
\end{table}

Define $\varpi(r, Z) = \Pr(C = r \mid Z)$. For now, we assume that $\varpi(r,
Z)$ is known. Notice that $\varpi(\infty, Z) = \pi_11$, $\varpi(3, Z) = \pi_{01}$,
$\varpi(2, Z) = \pi_{10}$, and $\varpi(1, Z) = \pi_{00}$. Since $\pi_{00} +
\pi_{10} + \pi_{01} + \pi_{11} = 1$, we only need to define three inclusion 
probabilities.\\

Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
function $g$. 
The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
we have the optimal estimator in the form:

\begin{align}
  &\hat \theta =\\ \nonumber
  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
  \frac{\delta_{10}}{\pi_{10}} (b_1(X_i, Y_{1i}) - E[g_i \mid X_i]) 
  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (b_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ \nonumber
  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - b_1(X_i, Y_{1i}) - b_2(X_i, Y_{2i}) + E[g_i \mid X_i])
\end{align}

%The proposed estimator is 
%
%\begin{align}\label{eq:propind}
%  &\hat \theta^{ind}_{prop} =\\ \nonumber
%  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
%  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
%  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
%  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
%\end{align}

\section*{Semiparametric Inference}

We know from Theorem 7.2 of \cite{tsiatis2006semiparametric} that if $\Pr(C =
\infty \mid Z) = \pi_{11} > 0$ 
then the semiparametric influence function has the form (see page 20 of notes):

\begin{equation}\label{lambda2}
 \frac{I(C = \infty) g(Z)}{\varpi(\infty, Z)} + \frac{I(C =
  \infty)}{\varpi(\infty, Z)} \left(\sum_{r \neq \infty} \varpi(r, G_r(Z))
  L_{2r}(G_r(Z))\right) - \sum_{r \neq \infty} I(C = r)L_{2r}(G_r(Z))
\end{equation}

where $L_{2r}(G_r(Z))$ is an arbitrary function of $G_r(Z)$. Notice, that this
form does not identify \textit{the} optimal estimator but a class of 
semiparametric functions. A reasonable choice of an estimator for $L_{2r}(G_r(Z))$ is

\[L_{2r}(G_r(Z)) = c_r E[g(Z) \mid G_r(Z)]\]

where each $c_r$ is a constant. This actually suggests a variety of estimators 
that each have different values of coefficients of $\delta$. Initially, we tried 
a direct projection onto the nuisance tangent space, but I got stuck. See Appendix A 
for this work.

\subsection*{Different Classes of Estimators}

All of the considered estimators have a similar form:

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

In this case, $c$ is a vector of constants $c = (c_0, c_1, c_2)$ that are 
chosen to minimize the variance of $\hat \theta$. See Appendix B for details on 
an example about choosing the values of $\hat c$.

\begin{tabular}{lrrr}
  \toprule
  Estimator & $\beta_0(\delta, c_0)$ & $\beta_1(\delta, c_1)$ & Implemented \\
  \midrule
  $\hat \theta_{prop}$ & $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta^{ind}_{prop}$ & $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta_{c}$ & $c_0\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \\
  $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $c_1\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta_{\delta}$ & $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ &
  $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ & \checkmark \\
  \bottomrule
\end{tabular}

%It turns out that $\hat \theta_{prop}$ from Equation~\ref{eq:prop} is contained within 
%the previous class of estimators with 
%
%\[L_{2r}(G_r(Z)) = \varpi(r, G_r(Z)) E[g(Z) \mid G_r(Z)].\]
%
%(See pages 81-82 of notes for proof.) \\

\section*{Simulation Studies}

\subsection*{Simulation 1}

To see which of these estimators is the best, we run a simulation study. In this 
study, we have the following variables:

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, $\pi_{11} = 0.4$ and $\pi_{00} = \pi_{10} = \pi_{01} = 0.2$.
The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

\begin{table}[ht!]
\caption{The true value of $\theta$ is 5. $\rho = 0.5$. The test statistic and 
p-value are from a one-sample t-test to see if the estimator is biased.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Algorithm & Bias & SD & Tstat & Pval\\
\midrule
Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
CC & 0.002 & 0.058 & 1.549 & 0.061\\
IPW & 0.006 & 0.210 & 1.504 & 0.066\\
$\hat \theta_{prop}$ & 0.002 & 0.051 & 1.911 & 0.028\\
$\hat \theta_{prop}^{ind}$ & 0.002 & 0.091 & 0.983 & 0.163\\
$\hat \theta_{c, ind}$ & 0.001 & 0.074 & 0.784 & 0.216\\
$\hat \theta_{\delta}$ & 0.002 & 0.051 & 1.964 & 0.025\\
\bottomrule
\end{tabular}
\end{table}

The main takeaway from this simulation study is that $\hat \theta_\delta$
and $\hat\theta_{prop}$ have the smallest standard deviation among the 
non-oracle estimators.

\subsection*{Simulation 2}
We now test a second simulation study. The setup of this simulation study 
is the same as the previous simulation. The only difference is that we 
are interested in estimating $\theta = E[Y_1^2 Y_2]$. In other words, now 
we have $g(Z) = Y_1^2 Y_2$.

\begin{table}[ht!]
  \caption{The true value of $E[g(Z)]$ is 10. $\rho = 0.5$. The test statistic and 
p-value are from a one-sample t-test to see if the estimator is biased.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Algorithm & Bias & SD & Tstat & Pval\\
\midrule
Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
CC & 0.023 & 0.824 & 1.518 & 0.065\\
IPW & 0.031 & 0.915 & 1.846 & 0.032\\
$\hat \theta_{prop}$ & 0.011 & 0.635 & 0.912 & 0.181\\
$\hat \theta_{c, ind}$ & 0.013 & 0.662 & 1.060 & 0.145\\
$\hat \theta_{\delta}$ & 0.011 & 0.636 & 0.983 & 0.163\\
\bottomrule
\end{tabular}
\end{table}

Like the first simulation, $\hat \theta_{prop}$ and $\hat \theta_{\delta}$ 
performed the best.

\newpage

\section*{Next Steps}

\begin{itemize}
  \item[1.] It turns out that \cite{tsiatis2006semiparametric} contains 
    a general technique for solving non-monotone missing data. They show 
    that the optimal estimator is found via the following double projection:

    \[L_2 = \Pi\left(\Pi\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid C,
    G_C(Z)\right) \mid Z\right)\]

    I can show that $\hat \theta_\delta$ is NOT such a double projection. 
    However, I think it is worth exploring what such an estimator (especially 
    if we consider a linear estimator) looks like.
  \item[2.] Similar to the previous point, \cite{tsiatis2006semiparametric} 
    discuss what the optimal estimator of a non-monotone problem looks like 
    and they determine that it solves the estimating equation

    \[\sum_{i = 1}^n \mathcal{L}(\mathcal{M}^{-1}(g(Z)))\]

    where $\mathcal{M}(g(Z)) = E[\mathcal{L}(g(Z)) \mid Z]$ and 
    $\mathcal{L}(g(Z)) = E[g(Z) \mid C, G_C(Z)]$. The difficult part is 
    understanding $\mathcal{M}^{-1}$, but there is conveniently the fact that

    \[\mathcal{M}^{-1}(g(Z)) = \lim_{n \to \infty} \phi_{n + 1}(Z) \text{ where } 
      \phi_{n + 1}(Z) = (I - \mathcal{M}) \phi_n(Z) + g(Z) \text{ and }
    \phi_0(Z) = g(Z).\]

    I have been investigating what this estimating equation yields if we 
    approximate $\mathcal{M}^{-1}$ with $\phi_1$. However, I have not 
    finished this yet.
  \item[3.] A different method to find the projection onto $\Lambda_2$ 
    could be to try to minimize the (KL) divergence between 
    $L_2$ and $\frac{\delta_{11}}{\pi_{11}}g(Z)$.
\end{itemize}

\newpage

\printbibliography

\newpage

\appendix

\section{Linear Expectations}

To simplify this problem, we consider the following estimator\footnote{This
estimator has slightly different coefficients compared to the initial
estimator.}:

\begin{align}
  &\hat \theta_c \\ \nonumber
  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
    \left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) -
    \left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) + 
  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
  &+
  \left(\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}}\right) - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
  \left(\left(\frac{\delta_{01} + \delta_{11}}{\pi_{01} + \pi_{11}}\right) - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
\end{align}

\subsection{Projection onto Nuisance Tangent Space}

The goal is now to find the coefficients $c_0, c_1$, and $c_2$ such that 
$\langle \hat \theta_c, L_2\rangle \equiv E[\hat \theta_c L_2] = 0$ for 
all $L_2 \in \Lambda_2$ (see \cite{tsiatis2006semiparametric} for definition 
of $\Lambda_2$). If we can find such coefficients that the estimator $\hat
\theta_c$ will be orthogonal to $\Lambda_2$ and hence by Theorem 10.1 of 
\cite{tsiatis2006semiparametric} semiparametrically optimal. The good news 
is that we know (from Theorem 7.2) that any element $L_2 \in \Lambda_2$ has a
form:

\begin{align}
  L_2 &= 
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{00} - \delta_{00}\right) L_{20}(X) +
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{10} - \delta_{10}\right) L_{21}(X, Y_1) +
  \left(\frac{\delta_{11}}{\pi_{11}}\pi_{01} - \delta_{01}\right) L_{22}(X, Y_2).
\end{align}

Then expanding and solving $E[\hat \theta_c L_2] = 0$ yields:

{\tiny
\begin{align*}
  0 &= E[\hat \theta_c L_2] \\ 
    &E\left[\left(\frac{\pi_{00}}{\pi_{11}} + \left(\frac{\pi_{10}}{\pi_{10} + \pi_{11}}\right)\frac{\pi_{00}c_1}{\pi_{11}} +
      \left(\frac{\pi_{01}}{\pi_{01} + \pi_{11}}\right) \frac{\pi_{00}c_2}{\pi_{11}} + \frac{\pi_{00}(\pi_{10}\pi_{01} -
  \pi_{11}^2)c_0}{(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})\pi_{11}}\right) E[g \mid X] L_{20}(X)\right. \\ 
    &+ \frac{\pi_{10}}{\pi_{11}} \left(E[g(Z) L_{21}(X, Y_1) \mid X] - c_1 E[E[g(Z) \mid X, Y_1] L_{21}(X, Y_1) \mid X] +
      \frac{\pi_{01}c_2}{\pi_{10} + \pi_{11}} E[E[g \mid X, Y_2] L_{21}(X, Y_1) \mid X] +
      \frac{\pi_{10} \pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})} E[g \mid X]
      E[L_{21}(X, Y_1) \mid X]c_0\right)\\
    &+ \left.\frac{\pi_{01}}{\pi_{11}} \left(E[g(Z) L_{22}(X, Y_2) \mid X] + \frac{\pi_{10} c_1}{\pi_{10} + \pi_{11}} E[E[g(Z) \mid X, Y_1] L_{22}(X, Y_2) \mid X] -
      c_2 E[E[g \mid X, Y_2] L_{22}(X, Y_2) \mid X] +
      \frac{\pi_{10}}{(\pi_{01} + \pi_{11})} E[g \mid X]
      E[L_{22}(X, Y_2) \mid X]c_0\right)\right]
\end{align*}
}

To solve for $c_0, c_1$, and $c_2$ we need the following to hold for any
$L_{21}(X, Y_1)$ and $L_{22}(X, Y_2)$:

{\small
\begin{align*}
  &1 + c_0 \frac{\pi_{01} \pi_{10} - \pi_{11}^2}{(\pi_{10} + \pi_{11})(\pi_{01}
  + \pi_{11})} + c_1 \frac{\pi_{10}}{\pi_{01} + \pi_{11}} + c_2
  \frac{\pi_{01}}{\pi_{01} + \pi_{11}} = 0\\ 
  &E\left[\left(g(Z) + c_0 \frac{\pi_{01}}{\pi_{01} + \pi_{11}}E[g(Z) \mid X] - c_1 E[g(Z) \mid X, Y_1] + c_2 \frac{\pi_{01}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X, Y_2]\right)L_{21}(X, Y_1) \mid X\right] = 0 \\ 
  &E\left[\left(g(Z) + c_0 \frac{\pi_{10}}{\pi_{10} + \pi_{11}}E[g(Z) \mid X] + c_1 \frac{\pi_{10}}{\pi_{10} + \pi_{11}} E[g(Z) \mid X, Y_1] - c_2 E[g(Z) \mid X, Y_2]\right)L_{22}(X, Y_2) \mid X\right] = 0
\end{align*}
}

Unfortunately, I am now stuck because it is unclear to me how to use any actual 
data to solve this problem.

\section{Solving for $\hat c$}

We can find the values of 
$c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
estimator. We can find these values by differentiating by $c_i$ and solving 
for $c_i$:

\begin{align*}
  \begin{bmatrix} \hat c_0 \\ \hat c_1 \\ \hat c_2 \end{bmatrix} 
  &= -
  \begin{bmatrix} 
    M_{11} & M_{12} & M_{13} \\
    M_{21} & M_{22} & M_{23} \\
    M_{31} & M_{32} & M_{33} \\
  \end{bmatrix}^{-1}
  \times 
  \begin{bmatrix}
    E[E[g \mid X]^2] \left(1 + \frac{\pi_{10}\pi_{01} - \pi_{11}^2}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
    E[E[g \mid X, Y_1]^2] \left(\frac{-\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
    E[E[g \mid X, Y_2]^2] \left(\frac{-\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right) \\
  \end{bmatrix}
\end{align*}

where

\begin{align*}
  M_{11} &= E[E[g \mid X]^2] \left(\frac{\pi_{11}^2 + \pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})} - 1\right) \\
  M_{12} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{13} &= E[E[g \mid X]^2] \left(\frac{-\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{22} &= E[V(E[g \mid X, Y_1] \mid X)] \left(\frac{\pi_{10}}{\pi_{11}(\pi_{10} + \pi_{11})}\right) \\
  M_{23} &= E[E[g \mid X, Y_1]E[g \mid X, Y_2]] \left(\frac{\pi_{10}\pi_{01}}{\pi_{11}(\pi_{10} + \pi_{11})(\pi_{01} + \pi_{11})}\right) \\
  M_{33} &= E[V(E[g \mid X, Y_2] \mid X)] \left(\frac{\pi_{01}}{\pi_{11}(\pi_{01} + \pi_{11})}\right)
\end{align*}

This estimator is similar to several other estimators:

\begin{align}\label{thetaind}
  \hat \theta^{ind}_c 
  &= \frac{\delta_{11}}{\pi_{11}} g(Z) + \left(1 -
    \left(\frac{\delta_{10}}{\pi_{10}}\right) -
    \left(\frac{\delta_{01}}{\pi_{01}}\right) + 
  \frac{\delta_{11}}{\pi_{11}}\right) c_0 E[g \mid X] \\ \nonumber
  &+
  \left(\frac{\delta_{10}}{\pi_{10}} - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_1 E[g \mid X, Y_1] +
  \left(\frac{\delta_{01}}{\pi_{01}} - 
  \frac{\delta_{11}}{\pi_{11}}\right) c_2 E[g \mid X, Y_2]
\end{align}

\begin{align}\label{thetadelta}
  \hat \theta^{\delta}_c 
  &= \frac{\delta_{11}}{\pi_{11}} g(Z) 
  + \left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right) 
   c_0 E[g \mid X] \\ \nonumber
  &+
  \left(\frac{\delta_{11}}{\pi_{11}} - 
  \frac{\delta_{10}}{\pi_{10}}\right) c_1 E[g \mid X, Y_1] +
  \left(\frac{\delta_{11}}{\pi_{11}} - 
  \frac{\delta_{01}}{\pi_{01}}\right) c_2 E[g \mid X, Y_2]
\end{align}

The first expression (Equation~\ref{thetaind}) is the proposed estimator 
with independent differences in each segment, while the second expression 
(Equation~\ref{thetadelta}) is the optimal estimator with values of $\delta$
such that $\hat \theta^{\delta}_c \in \Lambda_2$, which means that it has 
the form of the semiparametric in Equation~\ref{lambda2}.

\end{document}