### 2024-01-22 Notes/fstat_question.qmd
Added Quarto document with semiparametric constraint simulation.

---
title: "The F Test"
author:
  # Add additional authors here if needed, by replicating the above entries.
  - name: Caleb Leedy
# The short* versions of title and author are used for constructing page headers
# and are optional
date: today
date-format: "D MMMM YYYY"
#abstract: |
format:
  simple-article-pdf:
    include-in-header: "latex_header.tex"
    keep-tex: true
    indent: false
    classoption: 
     - 12pt
# reference-section-title: References
# bibliography: ../references.bib
---

# Computing the F-Test

To compute an F test we can use @eq-fstat.

$$ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
SSE_{Full} / DFE_{Full}}. $$ {#eq-fstat}

Let's assess what this would mean to compare models `Parametric 1` and 
`Outcome Robust` from @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

Consider the case where we run estimate each model on a data set with $n = 1000$
observations, and define the following notation: let $\hat \theta^{(P)}$ and 
$\hat \theta^{(OR)}$ be the estimated values of $\theta = E[Y_2]$ for the
`Parametric 1` model and the `Outcome Robust` model respectively. Define 
the estimated coefficients to be $\hat c_j^{(P)}$ and $\hat c_j^{(OR)}$ where 
$j = 1, \dots 9$ for the `Parametric 1` and `Outcome Robust` models
respectively. The one can compute the SSE with the following:

$$ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j $$

and $\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2
\mid x_i]$, $\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2
\mid x_i, y_{1i}]$, etc. Note that this is different from the previous estimate
of the variance which used the Monte Carlo variance (standard deviation) defined
by $$ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 $$ where
$B$ is the number of Monte Carlo estimates and $\bar \theta_B = \frac{1}{B}
\sum_{b = 1}^B \hat \theta_b$.

Likewise, one can compute the degrees of freedom by noticing that each model has
a degrees of freedom equal to nine minus the number of constraints. This means 
that we can compute the model degrees of freedom and error degrees of freedom
for each model type, which we do in @tbl-df

| Model | Model Degrees of Freedom | Error Degrees of Freedom |
| -----| ------------------------ | ------------------------ |
| Parametric 1    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Parametric 2    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Outcome Robust  | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Response Robust | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Double Robust   | $9 - 8 = 1$ | $n - 1 - 1 = n - 2$ |
: This table displays the degrees of freedom for each model {#tbl-df}

So continuing our first example, if $SSE^{(P)} = 993$ and $SSE^{(OR)} = 3428$ 
with $n = 1000$ then the F statistic is

$$F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.$$

The critical value we want to compare this with is the $0.95$ quantile of 
$F_{3, 993}$ which is $2.61$. So there *is* a significant difference 
between the fit of these two model. (The p-value is basically zero.)