### 2024-01-22 Notes/semiconstraints.qmd

---
title: "Understanding Semiparametric Constraints"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: "latex_header.tex"
    keep-tex: true
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

```

# Introduction {#sec-intro}

The goal of this report is to understand how additional constraints effect
semiparametric models. 
Inspired by Dr. Fuller's note, I set out to see if we could assess the amount 
of efficiency loss due to using semiparametric inference. First, let's define 
some notation.

| Segment | Quantity of Interest | Estimator | Coefficient |
| --------|----------------------| --------- | ------------ |
| $A_{00}$ | $E[g \mid X]$           | $\hat \gamma_{11}$ | $c_{11}$ |
| $A_{10}$ | $E[g \mid X]$           | $\hat \gamma_{21}$ | $c_{21}$ |
| $A_{10}$ | $E[g \mid X, Y_1]$      | $\hat \gamma_{22}$ | $c_{22}$ |
| $A_{01}$ | $E[g \mid X]$           | $\hat \gamma_{31}$ | $c_{31}$ |
| $A_{01}$ | $E[g \mid X, Y_2]$      | $\hat \gamma_{33}$ | $c_{33}$ |
| $A_{11}$ | $E[g \mid X]$           | $\hat \gamma_{41}$ | $c_{41}$ |
| $A_{11}$ | $E[g \mid X, Y_1]$      | $\hat \gamma_{42}$ | $c_{42}$ |
| $A_{11}$ | $E[g \mid X, Y_2]$      | $\hat \gamma_{43}$ | $c_{43}$ |
| $A_{11}$ | $E[g \mid X, Y_1, Y_2]$ | $\hat \gamma_{44}$ | $c_{44}$ |
: This table shows the relationship between different quantities of
interest in each segment, their estimators, and the corresponding coefficients
of their estimators. {#tbl-reftab}

This is a similar framework to understanding the class of estimators 
that we discussed in \verb|optest_update.pdf|, where we write the estimator 
as a weighted average of different expectation with functional coefficients.

$$\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].$$

However, now we write the estimator as

$$\hat \theta = \sum_{k,t} c_{kt}\hat \gamma_{kt}$$

where $\hat \gamma_{kt} = \frac{\delta_{ij}}{\pi_{ij}}E[g \mid G_{ij}(X, Y_1,
Y_2)]$ and $ij$ corresponds to the segment $A_{ij}$ associated with $\hat
\gamma_{kt}$ in @tbl-reftab.

# Methods {#sec-methods}

As noted by Dr. Fuller, we can understand parametric estimation as solving 
the following constrained optimization problems:

\begin{align*}
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{k,t} c_{kt} = 1 \\
       &\text{or} \\
  \min &\Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that } 
  \sum_{(k,t): (k,t) \neq (4, 4)} c_{kt} = 0 \text{ and } c_{44} = 1. \\
\end{align*}

To be robust to the outcome model, we can construct a similar optimization
problem with different constraints:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0,\\
       &\text{and } c_{44} = 1.
\end{align*}

Likewise to be robust to the response model, we can have the problem:

\begin{align*}
  \min \Var\left(\sum_{k, t} c_{kt} \hat \gamma_{kt}\right) \text{ such that }& 
  c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \\
       &\text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}.
\end{align*}

To be double robust (robust to the outcome and response model) we can combine 
the last two constraints. This is summarized in @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

# Simulation Study {#sec-simulations}

## Simulation 1 {.unnumbered}

We use the following simulation setup

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, $\pi_{11} = 0.2$ and $\pi_{00} = 0.3, \pi_{10} = 0.4$, and 
$\pi_{01} = 0.1$.
The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

Furthermore, we include four existing estimators that we have already 
proposed in the past: WLS, Prop, PropInd, and SemiDelta.

The estimator WLS is a a weight least square estimator derived in the 
following manner. We now consider a normal model:

$$ 
  \begin{pmatrix}
    x_i \\ e_{1i} \\ e_{2i}
  \end{pmatrix} \stackrel{ind}{\sim}
  N \left(
  \begin{bmatrix}
    0 \\ 0 \\ 0
  \end{bmatrix},
  \begin{bmatrix}
    1 & 0 & 0 \\
    0 & \sigma_{11} & \sigma_{12} \\ 
    0 & \sigma_{12} & \sigma_{22}
  \end{bmatrix},
  \right)
$$

and define $y_{1i} = \theta_1 + x_i + e_{1i}$ and $y_{2i} = \theta_2 + x_i
+ e_{2i}$. Then $b_1 = b_2 = 1$. We define $\bar z_k^{(ij)}$ as the mean of
$y_k$ in segment $A_{ij}$. This means that we have means $\bar z_1^{(11)}$,
$\bar z_2^{(11)}$, $\bar z_1^{(10)}$, and $\bar z_2^{(01)}$. Let $W = [\bar
z_1^{(11)}, \bar z_2^{(11)}, \bar z_1^{(10)}, \bar z_2^{(01)}]'$, then for
$n_{ij} = |A_{ij}|$, we have

$$Z - M \mu \sim N(\vec 0, V)$$

where

$$M = 
  \begin{bmatrix}
    1 & 0 \\
    0 & 1 \\
    1 & 0 \\
    0 & 1 \\
  \end{bmatrix}
  \text{ and }
  V = 
  \begin{bmatrix}
    \frac{\sigma_{11}}{n_{11}} & \frac{\sigma_{12}}{n_{11}} & 0 & 0 \\
    \frac{\sigma_{12}}{n_{11}} & \frac{\sigma_{22}}{n_{11}} & 0 & 0 \\
    0 & 0 & \frac{\sigma_{11}}{n_{10}} & 0 \\
    0 & 0 & 0 & \frac{\sigma_{22}}{n_{01}} \\
  \end{bmatrix}.
$$

Thus, the BLUE for $\mu = [\mu_1, \mu_2]'$ is

$$\hat \mu = (M' V^{-1} M)^{-1} M' V^{-1} W.$$

Hence, WLS is $\mu_2$ as $g(X, Y_1, Y_2) = Y_2$.

The remaining three estimators are drived from the following expression
where the values for $\beta$ are provided in @tbl-beta. These 
are good estimators to compare to because Prop is the original proposed 
estimator. PropInd has the same form as Prop except the values for $\beta$ 
is different. PropInd shares the same values of $\beta$ as all of the new 
models with constraints. SemiDelta is useful because it is the best 
estimator in general so far.

$$
\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + 
\beta_0(\delta, c_0)E[g(Z) \mid X] + 
\beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + 
\beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].
$$

| Estimator | $\beta_0(\delta, c_0)$ | $\beta_1(\delta, c_1)$ |
| --------- | ---------------------- | ---------------------- |
  Prop | $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ | $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ |
  PropInd | $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ | $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ |
  SemiDelta | $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ | $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ |

: This table displays the values of $\beta$ for different estimator types. {#tbl-beta}

In the simulation results in @tbl-sim1, the new results have the 
same label as the value from the `Type` column in @tbl-mods.

```{r}
#| label: libraries
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(CVXR)
library(ggplot2)

library(doParallel)
library(doRNG)
library(parallelly)

source("R/opt_est.R")
source("Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")

```

```{r}
#| label: mcmc E[g] = Y2
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

n_obs <- 1000
true_theta <- 5 
cov_e1e2 <- 0.5

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR")) %dorng% {

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    oracle_est <- mean(df$Y2)
    oraclex_est <- mean(df$Y2 - df$X)
    ipw_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_est = Y2 / prob_11) |>
      pull(g_est) |>
      sum() / nrow(df)
    cc_est <- 
      filter(df, delta_2 == 1) |>
      pull(Y2) |>
      mean()
    # em_est <- em_optsim(df)
    wls_est <- opt_lin_est(df, cov_y1y2 = cov_e1e2)
    wlsalt_est <- comb_lin_est_lin(df, theta2 = true_theta, cov_e1e2 = cov_e1e2)
    prop_est <- prop_nmono_est(df)
    propind_est <- prop_nmono_est(df, prop_ind = TRUE)
    propopt_est <- opt_theta_c(df)
    propoptdef_est <- opt_theta_c(df, est = "default")
    semiopt_est <- opt_semi_est(df)
    semidef_est <- opt_semi_est(df, est = "default")
    semidel_est <- opt_delta_c(df)
    semideldef_est <- opt_delta_c(df, est = "default")

    conalg_para <- min_var_alg(df, mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- min_var_alg(df, mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- min_var_alg(df, mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- min_var_alg(df, mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- min_var_alg(df, mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(tibble(oracle = oracle_est,
                  oraclex = oraclex_est,
                  cc = cc_est,
                  ipw = ipw_est,
                  # em = em_est,
                  wls = wls_est,
                  wlsalt = wlsalt_est,
                  prop = prop_est,
                  propind = propind_est,
                  propopt = propopt_est,
                  propoptdef = propoptdef_est,
                  semiopt = semiopt_est,
                  semidef = semidef_est,
                  semidel = semidel_est,
                  semideldef = semideldef_est,
                  conpara = conalg_para$theta_est,
                  conzero = conalg_sumzero$theta_est,
                  conout = conalg_outrob$theta_est,
                  conresp = conalg_resprob$theta_est,
                  condoub = conalg_doubrob$theta_est
    ))
  } |>
  bind_rows()

stopCluster(clust)

```

```{r}
#| label: tbl-sim1
#| echo: false
#| tbl-cap: |
#|    Results from Simulation 1. True values: $\theta = 5, \rho = 0.5$. 
#|    The test conducted for the T-statistic and P-value is a two sample test 
#|    to see if the estimator is unbiased. This simulation uses 
#|    $Y_1 = x + \varepsilon_1$, $Y_2 = \theta + x + \varepsilon_2$ where 
#|    $X \sim N(0, 1)$ and $(\varepsilon_1, \varepsilon_2)$ come from a mean 
#|    zero bivariate normal distribution with unit variance and covariance 
#|    $\rho$. The segments are unbalanced with: $\pi_{11} = 0.2$, 
#|    $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and $\pi_{00} = 0.3$.

mc_theta |>
  summarize(
    bias_oracle = mean(oracle) - true_theta,
    bias_cc = mean(cc) - true_theta,
    bias_ipw = mean(ipw) - true_theta,
    bias_wls = mean(wls) - true_theta,
    bias_prop = mean(prop) - true_theta,
    bias_semidel = mean(semidel) - true_theta,
    bias_conpara = mean(conpara) - true_theta,
    bias_conzero = mean(conzero) - true_theta,
    bias_conout = mean(conout) - true_theta,
    bias_conresp = mean(conresp) - true_theta,
    bias_condoub = mean(condoub) - true_theta,
    sd_oracle = sd(oracle),
    sd_cc = sd(cc),
    sd_ipw = sd(ipw),
    sd_wls = sd(wls),
    sd_prop = sd(prop),
    sd_semidel = sd(semidel),
    sd_conpara = sd(conpara),
    sd_conzero = sd(conzero),
    sd_conout = sd(conout),
    sd_conresp = sd(conresp),
    sd_condoub = sd(condoub),
    tstat_oracle = (mean(oracle) - true_theta) / sqrt(var(oracle) / B),
    tstat_cc = (mean(cc) - true_theta) / sqrt(var(cc) / B),
    tstat_ipw = (mean(ipw) - true_theta) / sqrt(var(ipw) / B),
    tstat_wls = (mean(wls) - true_theta) / sqrt(var(wls) / B),
    tstat_prop = (mean(prop) - true_theta) / sqrt(var(prop) / B),
    tstat_semidel = (mean(semidel) - true_theta) / sqrt(var(semidel) / B),
    tstat_conpara = (mean(conpara) - true_theta) / sqrt(var(conpara) / B),
    tstat_conzero = (mean(conzero) - true_theta) / sqrt(var(conzero) / B),
    tstat_conout = (mean(conout) - true_theta) / sqrt(var(conout) / B),
    tstat_conresp = (mean(conresp) - true_theta) / sqrt(var(conresp) / B),
    tstat_condoub = (mean(condoub) - true_theta) / sqrt(var(condoub) / B)
  ) |>
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(pval = pt(-abs(tstat), df = B)) |>
  mutate(Algorithm = case_when(
    Algorithm == "oracle" ~ "Oracle",
    Algorithm == "cc" ~ "CC",
    Algorithm == "ipw" ~ "IPW",
    Algorithm == "wls" ~ "WLS",
    Algorithm == "prop" ~ "$\\hat \\theta_{prop}$",
    Algorithm == "semidel" ~ "$\\hat \\theta_{\\delta}$",
    Algorithm == "conpara" ~ "Parametric 1",
    Algorithm == "conzero" ~ "Parametric 2",
    Algorithm == "conout" ~ "Outcome Robust",
    Algorithm == "conresp" ~ "Response Robust",
    Algorithm == "condoub" ~ "Double Robust",
    TRUE ~ "ERROR"
  )) |>
  knitr::kable(digits = 3)

```

## Simulation 2 {.unnumbered}

Next, we used the same simulation setup except that we are interested in 
estimating $\theta = E[g] = E[Y_1^2 Y_2]$. We do not use the WLS estimator 
because it does not work.

```{r}
#| label: mcmc E[g] = Y1^2 Y2
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR")) %dorng% {

    # g_fun <- "Y1^2Y2"

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    oracle_est <- mean(df$Y1^2 * df$Y2)
    oraclex_est <- mean((df$Y1)^2 * (df$Y2 - df$X))
    ipw_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_f = Y1^2 * Y2) |>
      mutate(g_est = g_f / prob_11) |>
      pull(g_est) |>
      sum() / nrow(df)
    cc_est <- 
      filter(df, delta_1 == 1, delta_2 == 1) |>
      mutate(g_f = Y1^2 * Y2) |>
      pull(g_f) |>
      mean()
    wls_est <- comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cov_e1e2)
    wlstt_est <- 
      comb_lin_est(df, gfun = "Y1^2 * Y2", cov_e1e2 = cov_e1e2, theta2 = true_theta)
    prop_est <- prop_nmono_est(df, gfun = "Y1^2 * Y2", pow = 3)
    propopt_est <- opt_theta_c(df, gfun = "Y1^2 * Y2", pow = 3)
    semiopt_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", pow = 3)
    semidef_est <- opt_semi_est(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)
    semidel_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", pow = 3)
    semideldef_est <- opt_delta_c(df, gfun = "Y1^2 * Y2", est = "default", pow = 3)

    conalg_para <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- 
      min_var_alg(df, gfun = "Y1^2 * Y2", mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(tibble(oracle = oracle_est,
                  oraclex = oraclex_est,
                  cc = cc_est,
                  ipw = ipw_est,
                  wls = wls_est,
                  wlstt = wlstt_est,
                  prop = prop_est,
                  propopt = propopt_est,
                  semiopt = semiopt_est,
                  semidef = semidef_est,
                  semidel = semidel_est,
                  semideldef = semideldef_est,
                  conpara = conalg_para$theta_est,
                  conzero = conalg_sumzero$theta_est,
                  conout = conalg_outrob$theta_est,
                  conresp = conalg_resprob$theta_est,
                  condoub = conalg_doubrob$theta_est
    ))
  } |> 
  bind_rows()

stopCluster(clust)

```

```{r}
#| label: tbl-sim2
#| echo: false
#| tbl-cap: |
#|    Results from Simulation 2. True values: $\theta = 10, \rho = 0.5$. 
#|    The test conducted for the T-statistic and P-value is a two sample test 
#|    to see if the estimator is unbiased. This simulation uses the same
#|    setup as Simulation 1 with
#|    $Y_1 = x + \varepsilon_1$, $Y_2 = \theta + x + \varepsilon_2$ where 
#|    $X \sim N(0, 1)$ and $(\varepsilon_1, \varepsilon_2)$ come from a mean 
#|    zero bivariate normal distribution with unit variance and covariance 
#|    $\rho$. The segments are unbalanced with: $\pi_{11} = 0.2$, 
#|    $\pi_{10} = 0.4$, $\pi_{01} = 0.1$, and $\pi_{00} = 0.3$.

true_g <- 2 * true_theta

mc_theta |>
  summarize(
    bias_oracle = mean(oracle) - true_g,
    bias_oraclex = mean(oraclex) - true_g,
    bias_cc = mean(cc) - true_g,
    bias_ipw = mean(ipw) - true_g,
    bias_wls = mean(wls) - true_g,
    bias_wlstt = mean(wlstt) - true_g,
    bias_prop = mean(prop) - true_g,
    bias_propopt = mean(propopt) - true_g,
    bias_semiopt = mean(semiopt) - true_g,
    bias_semidef = mean(semidef) - true_g,
    bias_semidel = mean(semidel) - true_g,
    bias_semideldef = mean(semideldef) - true_g,
    bias_conpara = mean(conpara) - true_g,
    bias_conzero = mean(conzero) - true_g,
    bias_conout = mean(conout) - true_g,
    bias_conresp = mean(conresp) - true_g,
    bias_condoub = mean(condoub) - true_g,
    sd_oracle = sd(oracle),
    sd_oraclex = sd(oraclex),
    sd_cc = sd(cc),
    sd_ipw = sd(ipw),
    sd_wls = sd(wls),
    sd_wlstt = sd(wlstt),
    sd_prop = sd(prop),
    sd_propopt = sd(propopt),
    sd_semiopt = sd(semiopt),
    sd_semidef = sd(semidef),
    sd_semidel = sd(semidel),
    sd_semideldef = sd(semideldef),
    sd_conpara = sd(conpara),
    sd_conzero = sd(conzero),
    sd_conout = sd(conout),
    sd_conresp = sd(conresp),
    sd_condoub = sd(condoub),
    tstat_oracle = (mean(oracle) - true_g) / sqrt(var(oracle) / B),
    tstat_oraclex = (mean(oraclex) - true_g) / sqrt(var(oraclex) / B),
    tstat_cc = (mean(cc) - true_g) / sqrt(var(cc) / B),
    tstat_ipw = (mean(ipw) - true_g) / sqrt(var(ipw) / B),
    tstat_wls = (mean(wls) - true_g) / sqrt(var(wls) / B),
    tstat_wlstt = (mean(wlstt) - true_g) / sqrt(var(wlstt) / B),
    tstat_prop = (mean(prop) - true_g) / sqrt(var(prop) / B),
    tstat_propopt = (mean(propopt) - true_g) / sqrt(var(propopt) / B),
    tstat_semiopt = (mean(semiopt) - true_g) / sqrt(var(semiopt) / B),
    tstat_semidef = (mean(semidef) - true_g) / sqrt(var(semidef) / B),
    tstat_semidel = (mean(semidel) - true_g) / sqrt(var(semidel) / B),
    tstat_semideldef = (mean(semideldef) - true_g) / sqrt(var(semideldef) / B),
    tstat_conpara = (mean(conpara) - true_g) / sqrt(var(conpara) / B),
    tstat_conzero = (mean(conzero) - true_g) / sqrt(var(conzero) / B),
    tstat_conout = (mean(conout) - true_g) / sqrt(var(conout) / B),
    tstat_conresp = (mean(conresp) - true_g) / sqrt(var(conresp) / B),
    tstat_condoub = (mean(condoub) - true_g) / sqrt(var(condoub) / B)
  ) |>
  tidyr::pivot_longer(cols = everything(),
               names_to = c(".value", "Algorithm"),
               names_pattern = "(.*)_(.*)") |>
  mutate(pval = pt(-abs(tstat), df = B)) |>
  filter(Algorithm %in% c("oracle", "cc", "ipw", "prop", "semidel", "conpara",
           "conzero", "conout", "conresp", "condoub")) |>
  mutate(Algorithm = case_when(
    Algorithm == "oracle" ~ "Oracle",
    Algorithm == "cc" ~ "CC",
    Algorithm == "ipw" ~ "IPW",
    Algorithm == "prop" ~ "$\\hat \\theta_{prop}$",
    Algorithm == "semidel" ~ "$\\hat \\theta_{\\delta}$",
    Algorithm == "conpara" ~ "Parametric 1",
    Algorithm == "conzero" ~ "Parametric 2",
    Algorithm == "conout" ~ "Outcome Robust",
    Algorithm == "conresp" ~ "Response Robust",
    Algorithm == "condoub" ~ "Double Robust",
    TRUE ~ "ERROR"
  )) |>
  knitr::kable(digits = 3)

```

## Discussion {.unnumbered}

There has been concern about why the constrained results are the same in
@tbl-sim1. This section aims to address these concerns.

```{r}
#| label: MCMC extended Sim 1
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

n_obs <- 1000
true_theta <- 5 
cov_e1e2 <- 0.5

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr", "CVXR", "tidyr")) %dorng% {

    # Generate Data
    # df <- gen_optsim_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
    df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)

    # Get Estimates
    conalg_para <- min_var_alg(df, mod = "para", cov_e1e2 = cov_e1e2)
    conalg_sumzero <- min_var_alg(df, mod = "sumzero", cov_e1e2 = cov_e1e2)
    conalg_outrob <- min_var_alg(df, mod = "out_rob", cov_e1e2 = cov_e1e2)
    conalg_resprob <- min_var_alg(df, mod = "resp_rob", cov_e1e2 = cov_e1e2)
    conalg_doubrob <- min_var_alg(df, mod = "double_rob", cov_e1e2 = cov_e1e2)

    return(
      tibble(conpara = conalg_para$theta_est,
             conzero = conalg_sumzero$theta_est,
             conout = conalg_outrob$theta_est,
             conresp = conalg_resprob$theta_est,
             condoub = conalg_doubrob$theta_est,
             parac = conalg_para$c_hat,
             zeroc = conalg_sumzero$c_hat,
             outc = conalg_outrob$c_hat,
             respc = conalg_resprob$c_hat,
             doubc = conalg_doubrob$c_hat,
             iter = iter,
             c_i = 1:9
      ) |> 
      group_by(conpara, conzero, conout, conresp, condoub) |> 
      nest()
    )
  } |>
  bind_rows()

stopCluster(clust)

```

Interestingly enough, the optimization algorithm does not have a lot of
variability in the output. The variance for each coefficient is found 
in @tbl-estvar. Since these values are all basically zero, we can see that
for each algorithm we are getting basically the same coefficients each
iteration. The coefficients that we see are displayed in @tbl-estc.

```{r}
#| label: tbl-estvar
#| echo: false
#| tbl-cap: |
#|    This table displays the variance of each coefficient for each estimator.
#|    A number close to zero means that the optimization algorithm chose 
#|    very similar values for each Monte Carlo simulation. A value of exactly 
#|    zero indicated that the optimization algorithm chose the exact same
#|    number.

c_tab <-
  bind_rows(mc_theta$data) |>
  pivot_longer(cols = parac:doubc, names_to = "estimator", values_to = "c_val") |>
  mutate(c_i = paste0("c_", c_i)) |>
  pivot_wider(names_from = "c_i", values_from = "c_val")

# Variance of each coefficient per estimator
alg_vec <- c("Para. 1", "Para. 2", "Outcome",
             "Response", "Double")
names(alg_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

ord_vec <- 1:5
names(ord_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

c_tab |>
  group_by(estimator) |>
  summarize(across(starts_with("c_"), var), .groups = "drop") |>
  rename_with(~paste0("$", .x, "$"), starts_with("c_")) |>
  mutate(order = ord_vec[estimator]) |>
  mutate(estimator = alg_vec[estimator]) |>
  rename(Estimator = estimator) |>
  arrange(order) |>
  select(-order) |>
  knitr::kable(digits = 3)

```

```{r}
#| label: tbl-estc
#| echo: false
#| tbl-cap: This table displays the estimated coefficients for each algorithm on
#|  iteration 1. Since the variance between iterations is small, these are
#|  reasonable for subsequent iterations.

c_tab |>
  filter(iter == 1) |>
  rename_with(~paste0("$", .x, "$"), starts_with("c_")) |>
  mutate(estimator = alg_vec[estimator]) |>
  rename(Estimator = estimator) |>
  select(-iter) |>
  knitr::kable(digits = 3)

```

```{r}
#| echo: false

#' get_f_statistic is a function to compute the F-statistic between two of 
#' the constrained models.
#'
#' @param c_tab - A table with values for the estimated c coefficients
#' @param reduced_mod - A string indicating which model is the reduced model
#' @param full_mod - A string indicating which model is the full model
#' @param df - A data set simulated from the same data generating process that 
#'    the c_tab coefficients were estimated from.
#'
#' @note We have previously shown that the variance of the estimated c_tab 
#' variables is close to or exactly zero. Hence, conducting a new simulation 
#' should not effect the results of the model much.
get_f_statistic <- function(c_tab, reduced_mod, full_mod,
                            df, theta = true_theta, cov_e1e2 = cov_e1e2) {

  vars_lst <- list(reduced_mod, full_mod)
  # SSE reduced 
  red_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[1]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # SSE full
  full_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[2]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # Constructing gam_vec
  Eg <- theta
  Egx <- theta + df$X 
  Egxy1 <- theta + df$X + cov_e1e2 * (df$Y1 - df$X)
  Egxy2 <- df$Y2
  EEg2 <- theta^2 + 2
  EEgx2 <- theta^2 + 1
  EEgxy12 <- theta^2 + 1 + cov_e1e2^2
  EEgxy22 <- theta^2 + 2
  EEgxy1Egxy2 <- theta^2 + 1 + cov_e1e2^2

  g_11 <- df$delta_00 / df$prob_00 * (Egx)
  g_21 <- df$delta_10 / df$prob_10 * (Egx)
  g_22 <- df$delta_10 / df$prob_10 * (Egxy1)
  g_31 <- df$delta_01 / df$prob_01 * (Egx)
  g_33 <- df$delta_01 / df$prob_01 * (Egxy2)
  g_41 <- df$delta_11 / df$prob_11 * (Egx)
  g_42 <- df$delta_11 / df$prob_11 * (Egxy1)
  g_43 <- df$delta_11 / df$prob_11 * (Egxy2)
  g_44 <- df$delta_11 / df$prob_11 * (df$Y2)

  gam_mat <- base::cbind(g_11, g_21, g_22, g_31, g_33, g_41, g_42, g_43, g_44)

  # SSE
  sse_red <- sum((as.numeric(gam_mat %*% t(t(red_cvec))) - theta)^2)
  sse_full <- sum((as.numeric(gam_mat %*% t(t(full_cvec))) - theta)^2)

  # Error Degrees of Freedom
  df_vec <- nrow(df) - 1 - c(9 - 1, 9 - 2, 9 - 4, 9 - 4, 9 - 8)
  names(df_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

  df_red <- df_vec[reduced_mod]
  df_full <- df_vec[full_mod]

  fstat <- ((sse_red - sse_full) / (df_red - df_full)) / (sse_full / df_full)

  pval <- 1 - pf(fstat, df_red - df_full, df_full)
  return(list(f = fstat, p = pval))

}

```

```{r}
#| label: tbl-fstat
#| echo: false
#| tbl-cap: This table computes the F-statistic and P-value for an F-test
#|    comparing the `Full Model` with a nested model in `Reduced Model`. The
#|    computation of the F-statistic is found in the `get_f_statistic` function.

df <- gen_unbal_data(n = n_obs, theta = true_theta, cov_e1e2 = cov_e1e2)
rf_tab <- tibble(rmod = c("zeroc", "outc",  "respc", "doubc",
                 "outc",  "respc", "doubc", "doubc", "doubc"),
                 fmod = c("parac", "parac", "parac", "parac",
                 "zeroc", "zeroc", "zeroc", "outc",  "respc"))

fstat_tab <-
  purrr::pmap(rf_tab, function(rmod, fmod) {
    get_f_statistic(c_tab, rmod, fmod, df = df, cov_e1e2 = cov_e1e2)
  }) |>
  bind_rows()

change_tab_names <- c("Parametric 1", "Parametric 2", "Outcome Robust",
                      "Response Robust", "Double Robust")
names(change_tab_names) <- c("parac", "zeroc", "outc", "respc", "doubc")

rf_tab |>
  mutate(Fstat = fstat_tab$f) |>
  mutate(fmod = change_tab_names[fmod]) |>
  mutate(rmod = change_tab_names[rmod]) |>
  select(fmod, rmod, Fstat) |>
  rename(`Reduced Model` = rmod) |>
  rename(`Full Model` = fmod) |>
  mutate(`P-Value` = fstat_tab$p) |>
  knitr::kable(digits = 2)

```

From the analysis in @tbl-fstat, we can see that there are significant
differences between several of the models. Why is this different from the 
previous analysis in @tbl-sim1? The answer comes from how we compute the
variance versus the sum of squares of error. The variance is computed from each
estimator created in a Monte Carlo trial; whereas the sum of squares of error is
summed from each individuals estimate for the mean of $Y_2$. For more details
about how the F-test is computed see Appendix A and Appendix B.

{{< pagebreak >}}

# Appendix A: Computing the F-Test {.appendix}

To compute an F test we can use @eq-fstat.

$$ F = \frac{(SSE_{Reduced} - SSE_{Full}) / (DFE_{Reduced} - DFE_{Full})}{
SSE_{Full} / DFE_{Full}}. $$ {#eq-fstat}

Let's assess what this would mean to compare models `Parametric 1` and 
`Outcome Robust` from @tbl-mods.

| Type | Constraints |
| -----| ------------|
| Parametric 1 | $\sum_{k, t} c_{kt} = 1$ |
| Parametric 2 |  $\sum_{k, t: (k, t) \neq (4, 4)} c_{kt} = 0, c_{44} = 1$ |
| Outcome Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, \text{ and } c_{44} = 1.$ |
| Response Robust | $c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
| Double Robust | $c_{11} + c_{21} + c_{31} + c_{41} = 0, c_{22} + c_{42} = 0, c_{33} + c_{43} = 0, c_{44} = 1, c_{11} = \pi_{00}, c_{21} + c_{22} = \pi_{10}, c_{31} + c_{33} = \pi_{01}, \text{ and } c_{41} + c_{42} + c_{43} + c_{44} = \pi_{11}$ |
: This table identifies the different constraints for each model type. {#tbl-mods}

Consider the case where we run estimate each model on a data set with $n = 1000$
observations, and define the following notation: let $\hat \theta^{(P)}$ and 
$\hat \theta^{(OR)}$ be the estimated values of $\theta = E[Y_2]$ for the
`Parametric 1` model and the `Outcome Robust` model respectively. Define 
the estimated coefficients to be $\hat c_j^{(P)}$ and $\hat c_j^{(OR)}$ where 
$j = 1, \dots 9$ for the `Parametric 1` and `Outcome Robust` models
respectively. The one can compute the SSE with the following:

$$ SSE = n^{-1} \sum_{i = 1}^n (\hat y_{2i} - \theta)^2 \text{ where }
\hat y_{2i} = \sum_{j = 1}^9 \hat c_j \hat \gamma_j $$

and $\hat \gamma_0 := \hat \gamma_{00} = \frac{\delta_{00i}}{\pi_{00}} E[Y_2
\mid x_i]$, $\hat \gamma_1 := \hat \gamma_{11} = \frac{\delta_{10i}}{\pi_{10}} E[Y_2
\mid x_i, y_{1i}]$, etc. Note that this is different from the previous estimate
of the variance which used the Monte Carlo variance (standard deviation) defined
by $$ \frac{1}{n - 1} \sum_{b = 1}^B (\hat \theta_b - \bar \theta_B)^2 $$ where
$B$ is the number of Monte Carlo estimates and $\bar \theta_B = \frac{1}{B}
\sum_{b = 1}^B \hat \theta_b$.

Likewise, one can compute the degrees of freedom by noticing that each model has
a degrees of freedom equal to nine minus the number of constraints. This means 
that we can compute the model degrees of freedom and error degrees of freedom
for each model type, which we do in @tbl-df

| Model | Model Degrees of Freedom | Error Degrees of Freedom |
| -----| ------------------------ | ------------------------ |
| Parametric 1    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Parametric 2    | $9 - 1 = 8$ | $n - 1 - 8 = n - 9$ |
| Outcome Robust  | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Response Robust | $9 - 4 = 5$ | $n - 1 - 5 = n - 6$ |
| Double Robust   | $9 - 8 = 1$ | $n - 1 - 1 = n - 2$ |
: This table displays the degrees of freedom for each model {#tbl-df}

So continuing our first example^[SSE values might differ in @tbl-fstat slightly
depending on the actual data values.], if $SSE^{(P)} = 993$ and $SSE^{(OR)} =
3428$ with $n = 1000$ then the F statistic is

$$F = \frac{(3428 - 993) / (3)}{993 / (1000 - 9)} = 810.$$

The critical value we want to compare this with is the $0.95$ quantile of 
$F_{3, 993}$ which is $2.61$. So there *is* a significant difference 
between the fit of these two model. (The p-value is basically zero.) The 
code to do this automatically is in Appendix B.

{{< pagebreak >}}

# Appendix B: R Functions {.appendix}

```{r}
#| lst-label: lst-getfstat
#| lst-cap: The R code for estimating the F statistic.
#| echo: true

#' get_f_statistic is a function to compute the F-statistic between two of 
#' the constrained models.
#'
#' @param c_tab - A table with values for the estimated c coefficients
#' @param reduced_mod - A string indicating which model is the reduced model
#' @param full_mod - A string indicating which model is the full model
#' @param df - A data set simulated from the same data generating process that 
#'    the c_tab coefficients were estimated from.
#'
#' @note We have previously shown that the variance of the estimated c_tab 
#' variables is close to or exactly zero. Hence, conducting a new simulation 
#' should not effect the results of the model much.
get_f_statistic <- 
  function(c_tab, reduced_mod, full_mod, df, theta = true_theta, cov_e1e2) {

  vars_lst <- list(reduced_mod, full_mod)

  # SSE reduced 
  red_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[1]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # SSE full
  full_cvec <-
    c_tab |>
    filter(iter == 1) |>
    filter(.data[["estimator"]] == vars_lst[[2]]) |>
    select(starts_with("c_")) |>
    as.numeric()

  # Constructing gam_vec
  Eg <- theta
  Egx <- theta + df$X 
  Egxy1 <- theta + df$X + cov_e1e2 * (df$Y1 - df$X)
  Egxy2 <- df$Y2
  EEg2 <- theta^2 + 2
  EEgx2 <- theta^2 + 1
  EEgxy12 <- theta^2 + 1 + cov_e1e2^2
  EEgxy22 <- theta^2 + 2
  EEgxy1Egxy2 <- theta^2 + 1 + cov_e1e2^2

  g_11 <- df$delta_00 / df$prob_00 * (Egx)
  g_21 <- df$delta_10 / df$prob_10 * (Egx)
  g_22 <- df$delta_10 / df$prob_10 * (Egxy1)
  g_31 <- df$delta_01 / df$prob_01 * (Egx)
  g_33 <- df$delta_01 / df$prob_01 * (Egxy2)
  g_41 <- df$delta_11 / df$prob_11 * (Egx)
  g_42 <- df$delta_11 / df$prob_11 * (Egxy1)
  g_43 <- df$delta_11 / df$prob_11 * (Egxy2)
  g_44 <- df$delta_11 / df$prob_11 * (df$Y2)

  gam_mat <- base::cbind(g_11, g_21, g_22, g_31, g_33, g_41, g_42, g_43, g_44)

  # SSE
  sse_red <- sum((as.numeric(gam_mat %*% t(t(red_cvec))) - theta)^2)
  sse_full <- sum((as.numeric(gam_mat %*% t(t(full_cvec))) - theta)^2)

  # Error Degrees of Freedom
  df_vec <- nrow(df) - 1 - c(9 - 1, 9 - 2, 9 - 4, 9 - 4, 9 - 8)
  names(df_vec) <- c("parac", "zeroc", "outc", "respc", "doubc")

  df_red <- df_vec[reduced_mod]
  df_full <- df_vec[full_mod]

  fstat <- ((sse_red - sse_full) / (df_red - df_full)) / (sse_full / df_full)

  pval <- 1 - pf(fstat, df_red - df_full, df_full)
  return(list(f = fstat, p = pval))

}

```