### 2024-01-30 Simulations/T4GetVarianceMin/variance_of_minimum.qmd

---
title: "Non-Monotone Missingness: The Minimal Variance of the Linear Estimate"
author: "Caleb Leedy"
date: today
date-format: "D MMMM YYYY"
format: 
  pdf:
    keep-tex: true
    include-in-header: latex_header.tex
---

# Introduction

The goal of this project is to get an optimal (or near optimal) estimator of a
general estimator when the data has a non-monotone missing structure. So far we
have reduced the problem to estimating the following:

$$
\hat \theta = \frac{\delta_{11}}{\pi_{11}} g(Z) + 
\lambda_0(\delta) \alpha_0(X) +
\lambda_1(\delta) \alpha_1(X, Y_1) +
\lambda_2(\delta) \alpha_2(X, Y_2).
$$ {#eq-estclass}

For notation we have variables $Z = (X, Y_1, Y_2)$ that are observed in the
segment noted in @tbl-segments.

| Segment | Variables | 
| ------- | ----------|
| $A_{00}$| $X$       |
| $A_{10}$| $X, Y_1$  |
| $A_{01}$| $X, Y_2$  |
| $A_{11}$| $X, Y_1, Y_2$ |
: This table matches the segment with its associated observed variables. {#tbl-segments}

For each segment $A_{d_1 d_2}$ the probability of an observation being in a specific
segment is known to be $\pi_{d_1 d_2}$, and the associated random variable that
indicates whether observation $i$ is in $A_{d_1 d_2}$ is $\delta_{d_1 d_2 i}$.
For brevity, it will sometimes be convienent to drop the subscript $i$. To get
an optimal estimator, we need to choose values of $\lambda$ and $\alpha$ to
minimize the variance of $\hat \theta$.

# Simplifications

To simplify this model consider the case in which each $\lambda$ *and* $\alpha$
are linear functions. This means that we have the following: 
$\lambda_0 = \lambda_0^{(0)} + \lambda_1^{(0)}\delta_{00} + \lambda_2^{(0)}
\delta_{10} + \lambda_3^{(0)} \delta_{01} + \lambda_4^{(0)}\delta_{11}$,
$\lambda_1 = \lambda_0^{(1)} + \lambda_2^{(1)} \delta_{10} +
\lambda_4^{(1)}\delta_{11}$, and 
$\lambda_2 = \lambda_0^{(2)} + \lambda_3^{(2)} \delta_{01} +
\lambda_4^{(2)}\delta_{11}$. Also, 
$\alpha_0 = \alpha_0^{(0)} + \alpha_1^{(0)} x$, 
$\alpha_1 = \alpha_0^{(1)} + \alpha_1^{(1)} x + \alpha_2^{(1)} y_1$, and
$\alpha_2 = \alpha_0^{(2)} + \alpha_1^{(1)} x + \alpha_3^{(1)} y_2$. This means
that we can express the estimator in @eq-estclass as the following matrix
equation:

$$
\hat \theta = n^{-1} \bf{1}_n' \left(\frac{\delta_{11}}{\pi_{11}} g(Z) + 
(\pmb{\alpha} \pmb{\lambda}' \pmb{\delta})' \bf Z\right).
$$

where

$$
\pmb{\alpha} = 
\begin{bmatrix}
\alpha_0^{(0)} & \alpha_0^{(1)} & \alpha_0^{(2)} \\
\alpha_1^{(0)} & \alpha_1^{(1)} & \alpha_1^{(2)} \\
0 & \alpha_2^{(1)} & 0 \\
0 & 0 & \alpha_3^{(2)} \\
\end{bmatrix},
\pmb{\lambda} = 
\begin{bmatrix}
\lambda_0^{(0)} & \lambda_0^{(1)} & \lambda_0^{(2)} \\
\lambda_1^{(0)} & 0 & 0 \\
\lambda_2^{(0)} & \lambda_2^{(1)} & 0 \\
\lambda_3^{(0)} & 0 & \lambda_3^{(2)} \\
\lambda_4^{(0)} & \lambda_4^{(1)} & \lambda_4^{(2)} \\
\end{bmatrix},
\pmb{\delta}= 
\begin{bmatrix}
1 \\ \delta_{00} \\ \delta_{10} \\ \delta_{01} \\ \delta_{11}
\end{bmatrix}
\bf{Z}_i = 
\begin{bmatrix}
1 \\ X_i \\ Y_{1i} \\ Y_{2i}
\end{bmatrix}
$$

This is a linear estimator and the goal is to find the optimal $\pmb{\alpha}$
and $\pmb{\lambda}$ that,

$$ \text{minimize } \Var(\hat \theta) \text{ such that } 
\pmb{\lambda}' E[\pmb{\delta}] = 0.$$

# Simpliying the Variance

We know that

\begin{align*}
\Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)) + 
  (\pmb \alpha \pmb \lambda' \pmb \delta)' \bf Z\right) 
&= \Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)\right) +
  \Var((\pmb \alpha \pmb \lambda' \pmb \delta)'\bf Z) +
  2 \Cov\left(\frac{\delta_{11}}{\pi_{11}}g(Z), 
  (\pmb \alpha \pmb \lambda' \pmb \delta)' \bf Z\right)\\
&:= V_1 + V_2 + V_3
\end{align*}

Hence, we have for $\Pi := E[\delta]$,

\begin{align*}
V_1 &= \Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z)\right) \\ 
&= \Var\left(E\left[\frac{\delta_{11}}{\pi_{11}}g(Z) \mid Z\right]\right)
+ E\left[\Var\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid Z\right)\right]\\
  &= E[g^2(Z)] - E[g(Z)]^2 + \frac{1}{\pi_{11}}E[g^2(Z)] - E[g^2(Z)] \\ 
  &= \frac{1}{\pi_{11}} E[g^2(Z)] - E[g(Z)]^2
\end{align*}

\begin{align*}
V_2 &= \Var((\alpha \lambda' \delta)'Z) \\ 
&= \Var(E[(\alpha \lambda' \delta)'Z \mid Z]) + 
E[\Var((\alpha \lambda' \delta)'Z \mid Z)] \\ 
&= \Pi' \lambda \alpha'  \Var(Z) \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' \Var(\delta)\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  (E[ZZ'] - E[Z]E[Z']) \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' (E[\delta \delta'] - \Pi\Pi')\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  E[ZZ'] \alpha \lambda' \Pi -
\Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] - 
E[Z'\alpha \lambda' \Pi\Pi'\lambda \alpha' Z]\\
&= \Pi' \lambda \alpha'  E[ZZ'] \alpha \lambda' \Pi -
\Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] - 
\Pi' \lambda \alpha' E[ZZ'] \alpha \lambda' \Pi \\
&= \Pi' \lambda \alpha' E[Z]E[Z'] \alpha \lambda' \Pi + 
E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] \\
&= E[Z'\alpha \lambda' E[\delta \delta']\lambda \alpha' Z] \\
\end{align*}

The third to last equality holds because a $1 \times 1$ matrix is always
symmetric and the last equality holds because $\Pi' \lambda = (\lambda' \Pi)' =
0'$. We can expand this expression but we get the horribly ugly result:

\begin{RaggedRight}

```{python}
#| echo: false
#| output: asis

import sympy

l00, l01, l02 = sympy.symbols('\lambda_0^{(0)} \lambda_0^{(1)} \lambda_0^{(2)}')
l10, l20, l21 = sympy.symbols('\lambda_1^{(0)} \lambda_2^{(0)} \lambda_0^{(1)}')
l30, l32 = sympy.symbols('\lambda_3^{(0)} \lambda_3^{(2)}')
l40, l41, l42 = sympy.symbols('\lambda_4^{(0)} \lambda_4^{(1)} \lambda_4^{(2)}')
lmat = sympy.Matrix([
  [l00, l01, l02],
  [l10, 0,   0 ],
  [l20, l21, 0],
  [l30, 0,   l32],
  [l40, l41, l42]
])

a00, a01, a02 = sympy.symbols('\\alpha_0^{(0)} \\alpha_0^{(1)} \\alpha_0^{(2)}')
a10, a11, a12 = sympy.symbols('\\alpha_1^{(0)} \\alpha_1^{(1)} \\alpha_1^{(2)}')
a21, a32 = sympy.symbols('\\alpha_2^{(1)} \\alpha_3^{(2)}')
amat = sympy.Matrix([
  [a00, a01, a02],
  [a10, a11, a12],
  [0,   a21, 0],
  [0,   0,   a32]
])

p00, p10, p01, p11 = sympy.symbols('\pi_{00} \pi_{10} \pi_{01} \pi_{11}')
ddmat = sympy.Matrix([
  [1,   p00, p10, p01, p11],
  [p00, p00, 0,   0,   0],
  [p10, 0,   p10, 0,   0],
  [p01, 0,   0,   p01, 0],
  [p11, 0,   0,   0,   p11]
])

x, y1, y2 = sympy.symbols('X Y_1 Y_2')
zmat = sympy.Matrix([[1, x, y1, y2]]).T

ans = zmat.T * amat * lmat.T * ddmat * lmat * amat.T * zmat
print(sympy.latex(sympy.expand(ans[0, 0]), itex = True, mode = "inline"))

```
\end{RaggedRight}

Finally, to understand $V_3$ we can solve for the third covariance term.

\begin{align*}
\Cov&\left(\frac{\delta_{11}}{\pi_{11}}g(Z), (\alpha \lambda' \delta)' Z\right)\\
&= E\left[\frac{\delta_{11}}{\pi_{11}}g(Z)(\alpha \lambda' \delta)' Z\right] - 
E\left[\frac{\delta_{11}}{\pi_{11}} g(Z)\right] E[(\alpha \lambda' \delta)'Z]\\
&= E\left[g(Z)\frac{\delta_{11}}{\pi_{11}}\delta' \lambda \alpha' Z\right] - 
E[g(Z)] (\alpha \lambda' \Pi)' E[Z]\\
&= E[g(Z) \begin{bmatrix}1 & 0 & 0 & 0 & 1\end{bmatrix} \lambda \alpha' Z] \\
&= (\lambda_0^{(0)} + \lambda_4^{(0)})(\alpha_0^{(0)} + \alpha_1^{(0)}E[g(Z)x])
+ (\lambda_0^{(1)} + \lambda_4^{(1)})(\alpha_0^{(1)} + \alpha_1^{(1)}E[g(Z)x] +
\alpha_2^{(1)}E[g(Z)y_1]) \\
&\qquad \qquad+ (\lambda_0^{(2)} + \lambda_4^{(2)})(\alpha_0^{(2)} +
\alpha_1^{(2)}E[g(Z)x] + \alpha_3^{(2)}E[g(Z)y_2])
\end{align*}

Since we have an understanding of $\Var(\hat \theta)$, we can find the minimum 
by differentiating with respect to each coefficient in $\lambda$ and $\alpha$.

```{python}
#| echo: false

egx, egy1, egy2 = sympy.symbols('E[gX] E[gY_1] E[gY_2]')
cov = (l00 + l40) * (a00 + a10 * egx) + (l01 + l41) * (a01 + a11 * egx + a21 *
  egy1) + (l02 + l42) * (a02 + a12 * egx + a32 * egy2)

var = ans[0, 0] + 2 * cov

# sympy.diff(var, l00)
# sympy.diff(var, l01)
# sympy.diff(var, l02)
# sympy.diff(var, l10)
# sympy.diff(var, l20)
# sympy.diff(var, l21)
# sympy.diff(var, l30)
# sympy.diff(var, l32)
# sympy.diff(var, l40)
# sympy.diff(var, l41)
# sympy.diff(var, l42)
# 
# sympy.diff(var, a00)
# sympy.diff(var, a01)
# sympy.diff(var, a02)
# sympy.diff(var, a10)
# sympy.diff(var, a11)
# sympy.diff(var, a12)
# sympy.diff(var, a21)
# sympy.diff(var, a32)

```

This is still a work in progress. I need to run a simulation and test it out.
I also believe that the terms $\alpha_0^{(0)}$, $\alpha_0^{(1)}$, and 
$\alpha_0^{(2)}$ hinder identifiability. But I will work on it.

```{r}
#| echo: false

library(dplyr)
library(CVXR)

source("~/Research/Data_Integration/R/opt_est.R")
source("~/Research/Data_Integration/Simulations/T3Assess_Semiparametric_Cost/compare_algs.R")

```

```{r}

linear_const_min <- function(df, gfun = "Y2") {

  df <- mutate(df, g_i = eval(rlang::parse_expr(gfun)))
  df_11 <- filter(df, delta_1 == 1, delta_2 == 1)
  df_10 <- filter(df, delta_1 == 1, delta_2 == 0)
  df_01 <- filter(df, delta_1 == 0, delta_2 == 1)
  df_00 <- filter(df, delta_1 == 0, delta_2 == 0)

  # Set up objective
  lambda <- Variable(rows = 5, cols = 3)
  alpha <- Variable(rows = 3, cols = 3)
  Z <- cbind(df$X, df$Y1, df$Y2)
  dd_mat <- matrix(c(1, pi_00, pi_10, pi_01, pi_11,
                     pi_00, pi_00, 0, 0, 0,
                     pi_10, 0, pi_10, 0, 0,
                     pi_01, 0, 0, pi_01, 0,
                     pi_11, 0, 0, 0, pi_11), nrow = 5, byrow = TRUE)

  mat <- quad_form(quad_form(t(Z %*% alpha %*% t(lambda)), dd_mat), diag(3))
  gz_mat <- matrix(df$g_i, ncol = 1)
  s_mat <- matrix(c(1, 0, 0, 0, 1), ncol = 1)
  covariance <- 
    matrix_trace(quad_form(Z %*% alpha %*% t(lambda) %*% s_mat %*% gz_mat),
                           diag(nrow(df)))
  obj <- Minimize(multiply(matrix_trace(mat), 1 / nrow(df)) + covariance))

}

```