### 2024-02-02 final/paper/old_paper/missing_mechanism.tex
Changed project structure.

\section*{Missingness Mechanism}

\begin{itemize}
  \item First I am going to reproduce the proof of double robustness that we
    talked about during our last meeting. I think it is insightful for future
    comments:

    \begin{align*}
      E[\hat \theta_{eff} - \theta_n] 
      &= E\left[n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] - g_i\right]\\
      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}
      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i])\right]\\
      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}
      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i])\right]\\
      &\quad+ E\left[n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)} 
      (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i \mid X_i])\right]\\
      &= n^{-1} \sum_{i = 1}^n (E[E[g_i \mid X_i]] - E[g_i])\\
      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} 
      (b_2(X_i, Y_{1i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
      (a_2(X_i, Y_{2i}) - E[g_i \mid X_i]) \mid X_i\right]\right] \\
      &\quad+ n^{-1} \sum_{i = 1}^n E\left[E\left[\frac{R_{1i}
      R_{2i}}{\pi_{11}(X_i)} (g_i - a_2(X_i, Y_{2i}) - b_2(X_i, Y_{1i}) + E[g_i
      \mid X_i]) \mid X_i \right]\right]\\
      \intertext{Since $R_{1i} \perp Y_{1i} \mid X_i$, $R_{2i} \perp Y_{2i} \mid
      X_i$, $(R_{1i}, R_{2i}) \perp (Y_{1i}, Y_{2i}) \mid X_i$ and $\pi_{1+},
      \pi_{2+}$ and $\pi_{11}$ are all free of $Y_{1i}$ and $Y_{2i}$.}
      &= n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}}{\pi_{1+}(X_i)} E\left[
        (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) \mid X_i\right]\right] \\ 
      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{2i}}{\pi_{2+}(X_i)} 
      E\left[E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i] \mid X_i\right]\right] \\
      &\quad+ n^{-1} \sum_{i = 1}^n E\left[\frac{R_{1i}
      R_{2i}}{\pi_{11}(X_i)} E\left[(g_i - E[g_i \mid X_i, Y_{2i}] - 
      E[g_i \mid X_i, Y_{1i}]+ E[g_i \mid X_i]) \mid X_i \right]\right]\\
      \intertext{Since $E[E[g_i \mid X_i, Y_{ki}] \mid X_i] = E[g_i \mid X_i] =
      0$,}
      &= 0.
    \end{align*}

    Thus, if the outcome models are correctly specified $\hat \theta_{eff}$ is
    unbiased. If the response models are correctly specified it is easy to see
    that $\hat \theta_{eff}$ is also unbiased. This means that $\hat
    \theta_{eff}$ is doubly robust.

  \item However, one of the key steps is that \textit{all} the response
    models are free of $Y$. In a previous iteration of Simulation 4, we had
    adopted the framework of \cite{robins1997non} where we first to observe the
    first variable and see if we observe the second variable. In this case, the
    second step can depend on the result of the first step and this is what we
    did. However, this makes it the case that $\pi_11$ is a function of $X_i$
    and $Y_1$ and $Y_2$. In this case $\hat \theta_{eff}$ is not unbiased if the
    response model is misspecified (or even just estimated).

  \item If we modify Simulation 4, such that the second step of observed the
    second variable is proportional to $\logistic(y_k)$ then we get the same
    result as before:

    \input{Tables/nonmonosim5c0.tex}
    \input{Tables/nonmonosim5c0.1.tex}
    \input{Tables/nonmonosim5c0.5.tex}

\end{itemize}

\newpage