### 2024-02-02 final/presentations/old_kimfuller_intro.qmd

---
title: "Combining Information from Multiple Surveys Using Probability Sampling"
author: "Caleb Leedy"
date: "August 15, 2023"
date-format: long
format: 
  revealjs:
    slide-number: c/t
    theme: [default, pres.scss]
bibliography: references.bib
---
<!---
Outline:
* Introduce notation

* Introduce problem:
  * semiparametrically efficient estimator to combine non-nested samples

* Proposed approach:
  * Use model-assisted semiparametric inference to get better / more robust results
  * i.e. use both a regression estimator and a model for the response probabilities

* Alternative Approaches:
  * Modified regression estimation of Merkouris (2004), Merkouris et al. (2023)
  * Multiphase regression estimator of Fuller (2009)

* Goals:
  * TODO:

* Next steps:
  * TODO:

Notation:
* Hidiroglou: Population universe (U), Sample i (s_i)
* Merkouris: same as above

* Inclusion probabilities: $\pi_k$
* Inclusion indicator R_i
* Variable names: x, y, z

--->

## Overall Goal

* We want to combine information from separate surveys and use 
semiparametric theory to get an optimal estimator.

# Existing Work

## Surveys with Missingness {.smaller}

Non-monotone Missingness

| Sample | $X$ | $Y_1$ | $Y_2$ | $R_1$ | $R_2$ |
|:------:|:---:|:-----:|:-----:|:-----:|:-----:|
|$A_{11}$| ✔️   | ✔️     | ✔️     | 1     | 1     |
|$A_{10}$| ✔️   | ✔️     |       | 1     | 0     |
|$A_{01}$| ✔️   |       | ✔️     | 0     | 1     |
|$A_{00}$| ✔️   |       |       | 0     | 0     |

* The goal is the estimate $\theta = E[g(X, Y_1, Y_2)]$.

## Estimating $\theta$ {.smaller}

Then for

\begin{align*}
b(X_i, Y_{1i}) &= E[g_i \mid X_i, Y_{1i}], \\ 
a(X_i, Y_{2i}) &= E[g_i \mid X_i, Y_{2i}], \text{ and } \\
i &= \{1, \dots, n\},
\end{align*}

\begin{align*}
\hat \theta_{\text{eff}} &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i}}{\pi_{1+}(X_i)}(b(X_i, Y_{1i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{2i}}{\pi_{2+}(X_i)}(a(X_i, Y_{2i}) - E[g_i \mid X_i]) \\ 
&+ n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_{11}(X_i)}(g_i - b_i - a_i + E[g_i \mid X_i])
\end{align*}

# Simulation Studies

## Simulation 1

We use the following simulation setup

$$
\begin{bmatrix}
X_i \\ 
\varepsilon_{1i} \\
\varepsilon_{2i} \\
\end{bmatrix} \stackrel{iid}{\sim}
N\left(
\begin{bmatrix}
0 \\ 0 \\ \theta
\end{bmatrix},
\begin{bmatrix}
1 & 0 & 0 \\ 
0 & 1 & \sigma_{yy} \\
0 & \sigma_{yy} & 1
\end{bmatrix}
\right)
$$

* $y_{1i} = x_i + \varepsilon_{1i}$
* $y_{2i} = x_i + \varepsilon_{2i}$

## Simulation 1

* Following the graphical MAR of [@robins1997non], for Stage 1, we have,

$$ p_0 = 0.2, p_1 = 0.4, \text{ and } p_2 = 0.4. $$

* For Stage 2, we have,

$$ p_{12} = \text{logistic}(y_1) \text{ and } p_{21} = \text{logistic}(y_2). $$

* The goal is to estimate $g = Y_2 - \theta$.

## Simulation 1

Then we compare three estimators:

* Oracle: we use both the observed and missing values.
* IPW-Oracle: we use 
$$ \hat \theta_{IPW} = n^{-1} \sum_{i = 1}^n \frac{R_{1i} R_{2i}}{\pi_i}y_{2i}.$$
* The proposed estimator $\hat \theta_{eff}$.

## Results

```{r}

library(MASS)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)

library(doParallel)
library(doRNG)
library(parallelly)

# *****************
# * Generate Data *
# *****************

source("../Simulations/generate_data.R")

# ************************
# * Estimating Functions *
# ************************

source("../Simulations/nmono_est_funs.R")

run_init_sims <- function(n_sim, B, true_mean, cor_y1y2,
                          miss_out, est_weights, seed, r_ind_y = FALSE) {

  clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
  registerDoParallel(clust)

  mc_theta <- 
    foreach(b = 1:B, 
            .options.RNG = seed,
            .export = c("nonmono_mar", "ipw_cc_oracle", "est_nonmono", "expit"),
            .packages = c("dplyr")) %dorng% {

    # Parameters
    df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
                      r_ind_y = r_ind_y, miss_out = miss_out)

    # Estimation
    # We want to compute:
    # 1. Oracle estimate
    # 2. pi_11 estimate (ipw using fully observed cases)
    # 3. New estimate

    tibble(oracle = mean(df$y2),
           ipworacle = ipw_cc_oracle(df),
           proposed = est_nonmono(df, oracle = !est_weights))

  } %>% bind_rows()

  stopCluster(clust)

  # ****************
  # * Analyze Data *
  # ****************

  mc_theta %>%
   summarize(
     bias_oracle = mean(oracle) - true_mean,
     bias_ipworacle = mean(ipworacle) - true_mean,
     bias_proposed = mean(proposed) - true_mean,
     sd_oracle = sd(oracle),
     sd_ipworacle = sd(ipworacle),
     sd_proposed = sd(proposed),
     tstat_oracle = (mean(oracle) - true_mean) / sqrt(var(oracle) / B),
     tstat_ipworacle = (mean(ipworacle) - true_mean) / sqrt(var(ipworacle) / B),
     tstat_proposed = (mean(proposed) - true_mean) / sqrt(var(proposed) / B))%>%
    tidyr::pivot_longer(cols = everything(),
                 names_to = c(".value", "algorithm"),
                 names_pattern = "(.*)_(.*)") %>%
    mutate(pval = pt(-abs(tstat), df = B))

}

# **********************************
# * Simulations: Sim 1 Nonmonotone *
# **********************************

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = -5, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is -5. Cor(eps1, eps2) = 0")

```

## Results

```{r}
run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(eps1, eps2) = 0")

```

## Results

```{r}
run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 5, cor_y1y2 = 0,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 5. Cor(eps1, eps2) = 0")

```

## Simulation 2

Simulation 2 follows the same structure are Simulation 1 except that 
instead of testing variation win the mean, we test differences in the covariance
of $\varepsilon_1$ and $\varepsilon_2$.

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.1,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.1")

```

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.5,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.5")

```

## Results

```{r}

run_init_sims(n_sim = 2000, B = 2000,
              true_mean = 0, cor_y1y2 = 0.9,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = "True Value is 0. Cor(Y1, Y2) = 0.9")

```

## Simulation 3

<!--- FIXME: -->

## Results

```{r}

library(dplyr)
library(CVXR)

library(doParallel)
library(doRNG)
library(parallelly)

# *****************************
# * Data Generation Functions *
# *****************************
source("generate_data.R")

# *************************************
# * Non/monotone Estimation Functions *
# *************************************
source("nmono_est_funs.R")

run_cali_sims <- function(n_sim, B, true_mean, cor_y1y2,
                          miss_out, est_weights, seed, r_ind_y = FALSE) {

  clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
  registerDoParallel(clust)

  mc_theta <- 
    foreach(b = 1:B, 
           .options.RNG = seed,
           .export = c("nonmono_mar", "ipw_cc_oracle", "est_nonmono", "expit",
                       "nonmono_est_weights", "twophase_reg", "threephase_reg"),
           .packages = c("dplyr", "CVXR")) %dorng% {

    # Parameters
    df <- nonmono_mar(n_sim, mean_y2 = true_mean, cor_y1y2 = cor_y1y2,
                      r_ind_y = r_ind_y, miss_out = miss_out)

    # Estimation
    # We want to compute:
    # 1. Oracle estimate
    # 2. pi_11 estimate (ipw using fully observed cases)
    # 3. New estimate

    tibble(oracle = mean(df$y2),
           ipworacle = ipw_cc_oracle(df),
           reg2p = twophase_reg(df, reg_on_y1 = FALSE),
           reg3p = threephase_reg(df),
           proposed = est_nonmono(df, oracle = !est_weights),
           calib = nonmono_est_weights(df))

  } %>% bind_rows()

  stopCluster(clust)

  # ****************
  # * Analyze Data *
  # ****************

  mc_theta %>%
   summarize(
     bias_oracle = mean(oracle) - true_mean,
     bias_ipworacle = mean(ipworacle) - true_mean,
     bias_proposed = mean(proposed) - true_mean,
     bias_reg2p = mean(reg2p) - true_mean,
     bias_reg3p = mean(reg3p) - true_mean,
     bias_calib = mean(calib) - true_mean,
     sd_oracle = sd(oracle),
     sd_ipworacle = sd(ipworacle),
     sd_proposed = sd(proposed),
     sd_reg2p = sd(reg2p),
     sd_reg3p = sd(reg3p),
     sd_calib = sd(calib),
     tstat_oracle = (mean(oracle) - true_mean) / sqrt(var(oracle) / B),
     tstat_ipworacle = (mean(ipworacle) - true_mean) / sqrt(var(ipworacle) / B),
     tstat_proposed = (mean(proposed) - true_mean) / sqrt(var(proposed) / B),
     tstat_reg2p = (mean(reg2p) - true_mean) / sqrt(var(reg2p) / B),
     tstat_reg3p = (mean(reg3p) - true_mean) / sqrt(var(reg3p) / B),
     tstat_calib = (mean(calib) - true_mean) / sqrt(var(calib) / B)) %>%
    tidyr::pivot_longer(cols = everything(),
                 names_to = c(".value", "algorithm"),
                 names_pattern = "(.*)_(.*)") %>%
    mutate(pval = pt(-abs(tstat), df = B))

}

tr_m <- -5
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 1) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2)) %>%

```

## Results

```{r}

tr_m <- 0
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 2) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2))

```

## Results

```{r}

tr_m <- 5
cy1y2 <- 0
run_cali_sims(n_sim = 1000, B = 1000,
              true_mean = tr_m, cor_y1y2 = cy1y2,
              miss_out = FALSE, est_weights = FALSE, seed = 4) %>%
  knitr::kable(booktabs = TRUE, digits = 3,
  caption = paste0("True Value is ", tr_m, ". Cor(Y1, Y2) = ", cy1y2))

```

## Problem Setup

::: {.incremental}
* Let $U$ be a finite population.
* Let $s_1$ and $s_2$ be samples of $U$
  - It may be the case that $s_1 \cap s_2 = \emptyset$ or $s_1 \cap s_2 \neq \emptyset$
  but $s_1 \subsetneq s_2$ and $s_2 \subsetneq s_1$.
  - This is an example of *non-nested* sampling [@hidiroglou2001double]
:::

## {}

::: {.incremental}
* Let $X$ be observed in both $s_1$ and $s_2$.
* Let $Y_1$ be only observed in $s_1$.
* Let $Y_2$ be only observed in $s_2$.
* We denote the first-order inclusion probabilities as $\pi^{(i)}_{j}$ of 
element $Y_j$ in survey $s_i$.
* If element $Y_j$ is in $s_i$, then let $R_j^{(i)} = 1$. Otherwise, $R_j^{(i)} = 0$.
:::

## {}

* This setup is very similar to the NRI project with $X$ consisting of core data
points and $Y_1$ and $Y_2$ being part of the supplemental samples.

## Research Problem

* How to estimate $\bar Y_1$ efficiently from information in
samples $s_1$ and $s_2$?

## Similar Methods

* Weight adjustment [@hidiroglou2001double]
* GREG Estimation [@merkouris2004combining, @merkouris2023combining]
* Projection estimator [@kim2012combining]
* Mass imputation [@chen2022nonparametric]

## Proposed Approach

* Use semiparametric theory to construct an efficient estimator.
* We fist consider the case of non-informative sampling.

# Comparison with Existing Methods

## Weight adjustment [@hidiroglou2001double]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Incorporates a regression model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not have semiparametric theory.
* Does not use a response model.

:::

::::

## GREG estimator [@merkouris2004combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Constructs optimal regression model (conditions out calibrated variables).

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not use a response model.

:::

::::

## Projection estimator [@kim2012combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Considers a non-nested sampling model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Imputes missing values of $Y_2$ with a mass imputation estimator.

:::

::::

## Mass imputation [@chen2022nonparametric]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Can be used to estimate $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Non-nested model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Nonparametric estimator.
* No response model.
* Considers $s_2$ to be non-probability sample.

:::

::::

## Proposed Next Steps

* Propose semiparametric estimator.

## References

## Simulation 2

## Simulation 3

## Problem Setup

::: {.incremental}
* Let $U$ be a finite population.
* Let $s_1$ and $s_2$ be samples of $U$
  - It may be the case that $s_1 \cap s_2 = \emptyset$ or $s_1 \cap s_2 \neq \emptyset$
  but $s_1 \subsetneq s_2$ and $s_2 \subsetneq s_1$.
  - This is an example of *non-nested* sampling [@hidiroglou2001double]
:::

## {}
::: {.incremental}
* Let $X$ be observed in both $s_1$ and $s_2$.
* Let $Y_1$ be only observed in $s_1$.
* Let $Y_2$ be only observed in $s_2$.
* We denote the first-order inclusion probabilities as $\pi^{(i)}_{j}$ of 
element $Y_j$ in survey $s_i$.
* If element $Y_j$ is in $s_i$, then let $R_j^{(i)} = 1$. Otherwise, $R_j^{(i)} = 0$.
:::

## {}

* This setup is very similar to the NRI project with $X$ consisting of core data
points and $Y_1$ and $Y_2$ being part of the supplemental samples.

## Research Problem

* How to estimate $\bar Y_1$ efficiently from information in
samples $s_1$ and $s_2$?

## Similar Methods

* Weight adjustment [@hidiroglou2001double]
* GREG Estimation [@merkouris2004combining, @merkouris2023combining]
* Projection estimator [@kim2012combining]
* Mass imputation [@chen2022nonparametric]

## Proposed Approach

* Use semiparametric theory to construct an efficient estimator.
* We fist consider the case of non-informative sampling.

# Comparison with Existing Methods

## Weight adjustment [@hidiroglou2001double]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Incorporates a regression model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not have semiparametric theory.
* Does not use a response model.

:::

::::

## GREG estimator [@merkouris2004combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$. 
* Considers the non-nested sampling model.
* Constructs optimal regression model (conditions out calibrated variables).

:::

::: {.column width="50%"}
Differences with Proposed Method

* No $Y_2$ in $s_2$.
* Does not use a response model.

:::

::::

## Projection estimator [@kim2012combining]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Estimates $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Considers a non-nested sampling model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Imputes missing values of $Y_2$ with a mass imputation estimator.

:::

::::

## Mass imputation [@chen2022nonparametric]

:::: {.columns}

::: {.column width="50%"}
Similarities with Proposed Method

* Can be used to estimate $\bar Y_1$.
* Considers $Y_2 \in s_2$.
* Non-nested model.

:::

::: {.column width="50%"}
Differences with Proposed Method

* Nonparametric estimator.
* No response model.
* Considers $s_2$ to be non-probability sample.

:::

::::

## Proposed Next Steps

* Propose semiparametric estimator.

## References