### 2024-02-05 src/explore/20240202-estcovmat.qmd
I realized that the Gram-Schmidt process was not going to work. We resort to estimating the covariance and variance directly.

---
title: "Estimating the Covariance Matrix"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we

1. Propose a new technique for estimating a covariance matrix, and 
2. Show via simulation studies that it works.

# Problem and Proposal

## Problem

Previously, have a model $\hat g = Z g + e$ where $\hat g$ is defined by

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_i) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(y_{1i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{2i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_i) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(y_{1i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_i) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{2i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_i) \\
\end{align*}

and

$$\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
$$

We also have

$$
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
$$
$$
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
$$
$$
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
$$

However, this challenge with actually using this model is that we have to use
the matrix $V$, which we assume to be *known*. In this write up, we will not
make this assumption.

## Proposals

### Proposal 1: REML

One potential solution would be to directly estimate $V$ and use some sort of
$\hat V$. We can do this with a REML estimate.

### Proposal 2: Orthogonal g-functions

For this idea we choose functions that after estimation are approximately
independent from each
other with a standard variance. If this is true, the covariance matrix $V$ is
approximately the identity matrix. Not only do we not have to estimate $V$, but
we do not even need to invert it! I now give a more detailed explanation of this
method.

Suppose that we observe variables $Z = (X_{m_1}, Y_{m_2})$ with an objective of
estimating $\theta = E[g(Y_{m_2})$ for some function $g$ such that the
variables are subject to missingness. We assume that there are $R$ unique
combinations of observed variables including a fully observed case. We can index
the combinations of observed variables by $r$ and assume that the fully observed
case occurs at $r = 1$. Let $G_r(Z)$ be the variables that are observed at a
particular value of $r$. We can choose a sequence of functions $f_1, \dots, f_K$
that we want to estimate. Each $f_k$ is assumed to be a function of a subset of
$Z$ and it makes sense to assume (since these are chosen by the analyst) that
there is at least one $f_k$ for each combination of observed variables $G_r(Z)$.
Let $A_k$ be the sets of observed variable combinations that can be evaluate by
the function $f_k$. If $f_k$ can be evaluated by the observed variables
$G_r(Z)$, this will consist of all of the combinations of variables
$r'$ such that $G_r(Z) \subseteq G_r'(Z)$. Because we assume that $G_1(Z) = Z$,
$A_k$ is always non-empty.

Once we have the original $f_1, \dots, f_K$ we orthogonalize them using a
Gram-Schmidt process. Let $g_1 = f_1 / \hat \Var(f_1)$. Then for $k > 1$
we can continue using the sequence,

$$\tilde g_k = f_k - \sum_{i = 1}^k \frac{\hat \Cov(f_k, g_i)}{\hat \Var(f_k)}
g_i \text{ followed by } g_k = \tilde g_k / \hat \Var(\tilde g_k).$$

However, the parameters for $\tilde g_k$ and $k > 1$ are all regression
coefficients and the multiplier for $g_k$ is simply computing a variance.
Then to
achieve efficiency, we propose estimating the regression coefficients by running
the coresponding regression on all of the data points in $A_k$ as well as
computing the variance of $\tilde g_k$ within $A_k$.

# Simulation

<!---# Compare with: Known V in non-monotone case --->
## Non-Monotone Case

First, we consider the following non-monotone simulation setup.

\begin{align*}
  \begin{bmatrix} x_1 \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  x_2 &= x + e_1 \\
  y &= \mu + x + e_2 \\
\end{align*}

This yields outcome variables $X_1$ and $X_2$ that are correlated both with $Y$
and additionally with each other. To generate the missingness pattern, we select
`r obs_seg` observations into the four segments independently using simple
random sampling. The goal of this estimation problem is to estimate $\theta = E[Y]$.

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/
source("R/data_generation.R")

```

```{r}
#| label: gls functions
#| echo: false

#' This function is also used in 20240202-gls_sim.qmd. I have modified the code
#' to account for the change from X, Y1, Y2 to X1, X2, Y.
gls_est <- function(df, class = "basic", rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, 2), nrow = 3)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_00 <- 1 / df$pi_00[1]

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: proposed estimation
#| echo: false

ortho_gls <- function(df, class = "basic") {

  # For class == 'basic', we consider the functions f_k to be originally,
  # f1 = X1, f2 = X2, f3 = Y.

  if (class == "basic") {

    # Get functions g_k
    g1 <- df$X1 / var(df$X1)
    df <- mutate(df, g1 = g1)
    df_1 <- filter(df, delta_11 == 1 | delta_10 == 1)
    t2 <- lm(X2 ~ g1, data = df_1)
    g2 <- df_1$X2 - t2$coefficients[2] * df_1$g1
    g2 <- g2 / var(g2)
    df_1 <- mutate(df_1, g2 = g2)
    df_11 <- filter(df_1, delta_11 == 1)
    t3 <- lm(Y ~ X1 + X2, data = df_11)
    g3 <- 
      df_11$Y - t3$coefficients[2] * df_11$g1 - t3$coefficients[3] * df_11$g2
    g3 <- g3 / var(g3)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * g1) |> pull(tmp) |> sum(),
        mutate(df_1, tmp = delta_11 / pi_11 * g2) |> pull(tmp) |> sum(),
        mutate(df_11, tmp = delta_11 / pi_11 * g3) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_10 / pi_10 * g1) |> pull(tmp) |> sum(),
        mutate(df_1, tmp = delta_10 / pi_10 * g2) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_01 / pi_01 * g1) |> pull(tmp) |> sum(),
        # ERROR: This does not work!
        mutate(df_11, tmp = delta_01 / pi_01 * g3) |> pull(tmp) |> sum(),
        mutate(df, tmp = delta_00 / pi_00 * g1) |> pull(tmp) |> sum())

    g_hat <- g_hat / nrow(df)

  } else {
    stop("Only class == 'basic' has been implemented.")
  }
}

```

<!---# Compare with: Best case in monotone data --->
## Monotone Case

# Variance Estimation

# Conclusion

Did it work?