### 2024-02-07 src/explore/20240206-covmat.qmd
Finished CSSM presentation.

---
title: "Estimating the Covariance Matrix"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

```{r}
#| label: global variables
#| echo: false

B <- 2000
obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we do the following:

1. Verify numerically the true covariance matrix in the simple
normal model.

# Problem

In `20240202-estcovmat.pdf` and `20240202-gls_sim.pdf`, we ran simulations in
which we used the *true* covariance matrix to get the best estimator.
Unfortunately, further analysis suggests that the covariance matrix used
previously was slightly incorrect. This report tests to see if this new result
is correct.

The previously accepted true variance estimate^[See Page 219 of notes] was the
following:

$$\Cov(\hat f_1, \hat f_2) = N^{-1} \pi^{-1} I(\delta_1 = \delta_2) \Cov(f_1,
f_2) $$

and the new variance estimate^[See Page 221 of notes] is

$$\Cov(\hat f_1, \hat f_2) = N^{-1}(I(\delta_1 \neq \delta_2) \Cov(f_1, f_2) +
\pi^{-1} I(\delta_1 = \delta_2) \Cov(f_1, f_2)).$$

The only difference between the estimates is the covariance terms between
segments. Are these independent or not?

# Simulation

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/ or src/explore/
source("R/data_generation.R")

```

```{r}
#| label: MCMC
#| echo: false

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_g <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

# Generate Data
  df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)
  g_hat <- 
    c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

  return(tibble(g1 = g_hat[1],
                g2 = g_hat[2],
                g3 = g_hat[3],
                g4 = g_hat[4],
                g5 = g_hat[5],
                g6 = g_hat[6],
                g7 = g_hat[7],
                g8 = g_hat[8]))
  } |>
  bind_rows()

stopCluster(clust)
cov_vec <- 
  c(cov(mc_g$g1, mc_g$g1),
    cov(mc_g$g1, mc_g$g2),
    cov(mc_g$g1, mc_g$g3),
    cov(mc_g$g1, mc_g$g4),
    cov(mc_g$g1, mc_g$g5),
    cov(mc_g$g1, mc_g$g6),
    cov(mc_g$g1, mc_g$g7),
    cov(mc_g$g1, mc_g$g8),
    cov(mc_g$g2, mc_g$g2),
    cov(mc_g$g2, mc_g$g3),
    cov(mc_g$g2, mc_g$g4),
    cov(mc_g$g2, mc_g$g5),
    cov(mc_g$g2, mc_g$g6),
    cov(mc_g$g2, mc_g$g7),
    cov(mc_g$g2, mc_g$g8),
    cov(mc_g$g3, mc_g$g3),
    cov(mc_g$g3, mc_g$g4),
    cov(mc_g$g3, mc_g$g5),
    cov(mc_g$g3, mc_g$g6),
    cov(mc_g$g3, mc_g$g7),
    cov(mc_g$g3, mc_g$g8),
    cov(mc_g$g4, mc_g$g4),
    cov(mc_g$g4, mc_g$g5),
    cov(mc_g$g4, mc_g$g6),
    cov(mc_g$g4, mc_g$g7),
    cov(mc_g$g4, mc_g$g8),
    cov(mc_g$g5, mc_g$g5),
    cov(mc_g$g5, mc_g$g6),
    cov(mc_g$g5, mc_g$g7),
    cov(mc_g$g5, mc_g$g8),
    cov(mc_g$g6, mc_g$g6),
    cov(mc_g$g6, mc_g$g7),
    cov(mc_g$g6, mc_g$g8),
    cov(mc_g$g7, mc_g$g7),
    cov(mc_g$g7, mc_g$g8),
    cov(mc_g$g8, mc_g$g8))

cov_f1f2 <- 
  c(1, 1, 1, 1, 1, 1, 1, 1,
    2, 1 + cov_e1e2, 1, 2, 1, 1 + cov_e1e2, 1,
    2, 1, 1 + cov_e1e2, 1, 2, 1,
    1, 1, 1, 1, 1,
    2, 1, 1 + cov_e1e2, 1,
    1, 1, 1,
    2, 1,
    1)
exp1_cov <- 
  c(4, 4, 4, 0, 0, 0, 0, 0,
    4, 4, 0, 0, 0, 0, 0,
    4, 0, 0, 0, 0, 0,
    4, 4, 0, 0, 0,
    4, 0, 0, 0,
    4, 4, 0,
    4, 0,
    4) / (4 * obs_seg)
exp2_cov <- 
  c(4, 4, 4, 1, 1, 1, 1, 1,
    4, 4, 1, 1, 1, 1, 1,
    4, 1, 1, 1, 1, 1,
    4, 4, 1, 1, 1,
    4, 1, 1, 1,
    4, 4, 1,
    4, 1,
    4) / (4 * obs_seg)

exp1 <- exp1_cov * cov_f1f2
exp2 <- exp2_cov * cov_f1f2

tibble(Actual = cov_vec, Exp1 = exp1, Exp2 = exp2) |>
  knitr::kable(digits = 5)

```

# Conclusion

Which is correct? It looks like the first derivation is correct. This does make
sense as I view the different segments as independent. The question is then what
mistake did I make in my own analysis?