### 2024-02-13 references/literature/papers.qmd
It looks like we can estimate the variance. This holds if we just use X1 and X2 or if we use f(X1) and f(X2).

---
title: "Papers Read"
author: "Caleb Leedy"
date: February 12, 2024
format: pdf
bibliography: references.bib
---

# @zhou2012efficient

Title: An efficient method of estimation for longitudinal surveys with monotone
missing data

Author(s): Ming Zhou and Jae-Kwang Kim

Journal (Year): Biometrika (2012)

Importance: I think that this paper does something very similar to our project
except with monotone data rather than nonmonotone data. The newness in the paper
is that the MAR assumption depends on variables oberseved in the previous time
period.

Innovation: They use a likelihood method for estimating the propensity score. I
am also curious about how they got Equation 9, which they claim is an optimal
estimator.

{{< pagebreak >}}

# References