### 2024-02-13 src/explore/20240208-optimalf.qmd

---
title: "Finding Optimal f-Functions"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

# Summary

This document takes the case of observing three variables $(X_1, X_2, Y)$ under
nonmonotone missingness and discusses the optimal function $f_1$, $f_2$, and
$f_3$ under a simple random sampling design.

# Notation and Problem Setup

Consider the case where $Z = (X_1, X_2, Y)$ and we want to estimate the
parameter $\theta = E[Y]$. Suppose that we have a finite population of size $N$.
Instead of observing the entire data set we observe 
the segments in @tbl-vars1. Each of the segments is a simple random sample of
size $n$ and independent from the other segments. Hence this means that the
first order selection probability $\pi_i = \pi = n / N$ for every $i$.

| Segment  | Variables Observed |
| -------  | ------------------ |
| $A_{00}$ | $X_1$              |
| $A_{10}$ | $X_1, X_2$         |
| $A_{01}$ | $X_1, Y$           |
| $A_{11}$ | $X_1, X_2, Y$      | 
: This table identifies which variables are observed in each segment. Since $X_1$
is always observed, the subscript for each segment identifies which of variables
$X_2$ and $Y$ are in the segment based on the position of a 1. {#tbl-vars1}

As an analyst, we can choose functions $f_1, f_2, f_3$ such that $g = Zf + e$
where

\begin{align*}
g_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_1(x_{1i}) \\
g_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_2(x_{2i}) \\
g_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} f_3(y_{i}) \\
g_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} f_1(x_{1i}) \\
g_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} f_2(x_{2i}) \\
g_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} f_1(x_{1i}) \\
g_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} f_2(y_{i}) \\
g_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} f_1(x_{1i}) \\
\end{align*}

where

$$\hat g = 
\begin{bmatrix}
g_1^{(11)} \\
g_2^{(11)} \\
g_3^{(11)} \\
g_1^{(10)} \\
g_2^{(10)} \\
g_1^{(01)} \\
g_3^{(01)} \\
g_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = V.$$

For now assume that $V$ is known. Due to the construction of $g$, we know that

$$V = 
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}
$$

where

$$
V_{11} = N^{-1} \pi^{-1} \Cov(f, f), 
V_{10} = N^{-1} \pi^{-1} 
\begin{bmatrix}
\Cov(f_1, f_1) & \Cov(f_1, f_2) \\ \Cov(f_1, f_2) & \Cov(f_2, f_2) 
\end{bmatrix},
$$

$$
V_{01} = N^{-1} \pi^{-1} 
\begin{bmatrix}
\Cov(f_1, f_1) & \Cov(f_1, f_3) \\ \Cov(f_1, f_3) & \Cov(f_3, f_3) 
\end{bmatrix}, \text{ and }
V_{00} = N^{-1} \pi^{-1} \Cov(f_3, f_3).
$$

Previously, we have shown that the optimal estimator is the GLS estimator $\hat
f = (Z'V^{-1}Z)^{-1} Z'V^{-1}g$ and by linear model theory 
$\Var(\hat f) = (Z'V^{-1}Z)^{-1}$ since $\Var(g) = V$ by construction.

```{python}
#| echo: false

import sympy

s1, s2, s3 = sympy.symbols('\sigma_1 \sigma_2 \sigma_3')
p12, p13, p23 = sympy.symbols('\\rho_{12} \\rho_{13} \\rho_{23}')

v_11 = sympy.Matrix([
  [s1 * s1, p12 * s1 * s2, p13 * s1 * s3],
  [p12 * s1 * s2, s2 * s2, p23 * s2 * s3],
  [p13 * s1 * s3, p23 * s2 * s3, s3 * s3]
])
v_10 = sympy.Matrix([
  [s1 * s1, p12 * s1 * s2],
  [p12 * s1 * s2, s2 * s2]
])
v_01 = sympy.Matrix([
  [s1 * s1, p13 * s1 * s3],
  [p13 * s1 * s3, s3 * s3]
])

v_00 = sympy.Matrix([s1 * s1])

v_mat = diag(v_11, v_10, v_01, v_00)

z_mat = sympy.Matrix([
  [1, 0, 0],
  [0, 1, 0],
  [0, 0, 1],
  [1, 0, 0],
  [0, 1, 0],
  [1, 0, 0],
  [0, 0, 1],
  [1, 0, 0]
])
```