### 2024-02-13 src/explore/20240213-optcoefs.qmd

---
title: "Finding the Optimal Coefficient Forms"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    include-in-header: 
      - "latex_header.tex"
---

```{r}
#| label: global variables
#| echo: false

B <- 2000
obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we do the following:

1. Vary the functional forms of the coefficients, 
2. Check to ensure that our estimated variance is still accurate.

# Idea

In `20240202-estcovmat.qmd`, we tested an estimated covariance matrix. The results
suggest that estimating the covariance matrix suffers little penalty. In this
report, we want to see if choosing different functional forms of $f_1$ and $f_2$ 
(that depend on the data) cause the covariance estimation to be biased.

Instead of using $f_1(X_1) = X_1$ and $f_2(X_2) = X_2$ as we did in 
`20240202-estcovmat.qmd` we want to use $f_1(X_1) = E[Y \mid X_1]$ and 
$f_2(X_2) = E[Y \mid X_2]$, *and* we need to also test them with estimated
versions of these functions.

# Simulation

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/ or src/explore/
source("R/data_generation.R")

```

```{r}
#| label: gls functions
#| echo: false

#' This function is also used in 20240202-gls_sim.qmd. I have modified the code
#' to account for the change from X, Y1, Y2 to X1, X2, Y.
gls_est <- function(df, class = "basic", rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, 2), nrow = 3)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_00 <- 1 / df$pi_00[1]

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: estimated variance basic
#| echo: false

est_var_gls <- function(df, class = "basic") {

  if (class == "basic") {

    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    # Estimate Covariance Matrix
    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
    df_11 <- filter(df, delta_11 == 1)
    v_x1 <- var(df$X1)
    v_x2 <- var(df_1p$X2)
    v_y <- var(df_p1$Y)
    cov_x1x2 <- cov(df_1p$X1, df_1p$X2)
    cov_x1y <- cov(df_p1$X1, df_p1$Y)
    cov_x2y <- cov(df_11$X2, df_11$Y)

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(v_x1, cov_x1x2, cov_x1y,
               cov_x1x2, v_x2, cov_x2y,
               cov_x1y, cov_x2y, v_y), nrow = 3)

    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
    v_00 <- 1 / df$pi_00[1] * v_x1

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

    } else {
    stop("Only class == 'basic' has been implemented.")
  }
}

```

```{r}
#| label: estimated variance opt
#| echo: false

estvar_glsopt <- function(df, class = "opt") {

  if (class == "opt") {

    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    z_mat <- matrix(rep(1, 8), ncol = 1)

    # Get Regression Estimates
    mx1 <- lm(Y ~ X1, data = filter(df, delta_11 == 1 | delta_01 == 1))
    mx2 <- lm(Y ~ X2, data = filter(df, delta_11 == 1))
    df <- df |>
      mutate(mx1 = predict(mx1, newdata = df)) |>
      mutate(mx2 = predict(mx2, newdata = df))

    # Put Estimates into g_hat
    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * mx1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * mx2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * mx1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * mx2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * mx1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * mx1) |> pull(tmp) |> mean())

    # Estimate Covariance Matrix
    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
    df_11 <- filter(df, delta_11 == 1)
    v_x1 <- var(df$mx1)
    v_x2 <- var(df_1p$mx2)
    v_y <- var(df_p1$Y)
    cov_x1x2 <- cov(df_1p$mx1, df_1p$mx2)
    cov_x1y <- cov(df_p1$mx1, df_p1$Y)
    cov_x2y <- cov(df_11$mx2, df_11$Y)

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(v_x1, cov_x1x2, cov_x1y,
               cov_x1x2, v_x2, cov_x2y,
               cov_x1y, cov_x2y, v_y), nrow = 3)

    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
    v_00 <- 1 / df$pi_00[1] * v_x1

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(ghat))
  } else {
    stop("Only class == 'opt' has been implemented.")
  }
}

```

```{r}
#| label: MCMC
#| echo: false
#| tbl-cap: |
#|   Results from simulations study with independent equally sized segments
#|   $A_{11}$, $A_{10}$, $A_{01}$, and $A_{00}$ all of size $n = 250$. In this
#|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
#|   between $e_1$ and $e_2$ is $\rho = 0.5$. The goal is to estimate 
#|   $E[Y_2] = \mu$. For the GLS estimation, we use the
#|   estimated covariance matrix $\hat V$ with 
#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$. For the GLS optimal
#|   estimation we are trying $f_1 = E[Y \mid X_1]$, $f_2 = E[Y \mid X_2]$ and 
#|   $f_3 = Y$. Each of these expectations are estimated with using linear
#|   regression. So $E[Y \mid X_i]$ is implemented as a regression estimator
#|

clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

    # Generate Data
    df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)

    # Estimators
    oracle <- mean(df$Y)
    cc <- df |>
      filter(delta_11 == 1) |>
      pull(Y) |>
      mean()
    ipw <- df |>
      mutate(tmp = delta_11 / pi_11 * Y) |>
      pull(tmp) |>
      mean()

    gls <- gls_est(df, rho = cov_e1e2)
    gls_ev <- est_var_gls(df)
    gls_opt <- estvar_glsopt(df)

    return(tibble(Oracle = oracle,
                  CC = cc,
                  IPW = ipw,
                  GLS = gls,
                  GLSEstVar = gls_ev,
                  GLSOpt = gls_opt
    ))

  } |>
  bind_rows()

stopCluster(clust)

# Analyze Results
mc_theta |>
  summarize(
    across(everything(),
    list(Bias = ~mean(.) - true_mu,
         SD = ~sd(.),
         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
    .names = "{.fn}_{.col}")) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(Pval = pt(-abs(Tstat), df = B)) |>
  knitr::kable(digits = 3)

```
# Conclusion

Is the covariance estimation still working? (By working we mean causing the
estimator to be unbiased.)

* It looks like estimating the variance is not causing additional bias in the
  estimator. This is great. 
* We may, however, need to check more complicated simulation setups because the
  simpicity of the current setup might make it look better than it actually is.
* One suprising thing is that the "optimal" estimator is not better than the
  other estimators. This could be because of the simplicity of the setup, but we
  may need to check it out.