### 2024-02-17 src/explore/20240215-sympy.qmd
Updated simulation with true and estimated marginal and conditional GLS. See 20240216-optest.qmd.

---
title: "SymPy Calculations"
author:
  - name: Caleb Leedy
date: today
date-format: "D MMMM YYYY"
format:
  pdf:
    classoption: [landscape]
    include-in-header: 
      - "latex_header.tex"
---

# Summary

This document records computations that I think would be easier to make in SymPy
than on paper. For now it does the following:

1. Compute the optimal values of $\theta$, $\eta_1$ and $\eta_2$ for the simple
nonmonotone GLS estimator (Notes: Pages 237-243+).

\tiny
```{python}
#| label: sympyres1
#| echo: false
#| output: asis

import sympy

# Steps:
# 1. Solve for eta_1
# 2. Solve for eta_2
# 3. Solve for theta

# We need variables for what we observe:
x1_1, x2_1, x3_1, x4_1 = sympy.symbols('X^1_1 X^2_1 X^3_1 X^4_1')
x1_2, x2_2, y1, y2 = sympy.symbols('X^1_2 X^2_2 Y^1 Y^2')

# Variances 
vx1_1, vx2_1 = sympy.symbols('V(X^1_{1}) V(X^2_{1})')
vx3_1, vx4_1 = sympy.symbols('V(X^3_{1}) V(X^4_{1})')
vx1_2, vx2_2, vy1, vy2 = sympy.symbols('V(X^1_{2}) V(X^2_{2}) V(Y^{1}) V(Y^{2})')

# Covariances
cx11x12, cx21x22 = sympy.symbols('C(X^1_{1}X^1_{2}) C(X^2_{1}X^2_{2})')
cx11y1 = sympy.symbols('C(X^1_{1}Y^{1})')
cx12y1, cx31y2 = sympy.symbols('C(X^1_{2}Y^{1}) C(X^3_{1}Y^{2})')

# 1. Solve for eta_1
x_1_mat = sympy.Matrix([[x1_1], [x2_1], [x3_1], [x4_1]])
vq1 = sympy.diag(vx1_1, vx2_1, vx3_1, vx4_1)
denom_q1 = (sympy.ones(1, 4) * (vq1**(-1)) * sympy.ones(4, 1))**(-1)
numer_q1 = sympy.ones(1, 4) * (vq1**(-1)) * x_1_mat
eta1 = (denom_q1 * numer_q1)[0, 0]

eta1 = sympy.symbols('\eta_1')

# 2. Solve for eta_2
x_2_mat = sympy.Matrix([[x1_2], [x2_2]])
q2adj_mat = sympy.Matrix([
  [cx11x12 / vx1_1 * (x1_1 - eta1)],
  [cx21x22 / vx2_1 * (x2_1 - eta1)]
])
vq2 = sympy.diag(vx1_2 - cx11x12**2 / vx1_1, vx2_2 - cx21x22**2 / vx2_1)
denom_q2 = (sympy.ones(1, 2) * (vq2**(-1)) * sympy.ones(2, 1))**(-1)
numer_q2 = sympy.ones(1, 2) * (vq2**(-1)) * (x_2_mat - q2adj_mat)
eta2 = (denom_q2 * numer_q2)[0, 0]

# 3. Solve for theta
qv11 = sympy.diag(vx1_1, vx2_1, vx3_1, vx4_1, vx1_2, vx2_2)
qv11[0, 4] = cx11x12
qv11[4, 0] = cx11x12
qv11[1, 5] = cx21x22
qv11[5, 1] = cx21x22

qv11_inv = qv11**(-1)
qv21 = sympy.Matrix([
  [cx11y1, 0, 0, 0, cx12y1, 0],
  [0, 0, cx31y2, 0, 0, 0]
])
qv22 = sympy.diag(vy1, vy2)
qv = qv22 - qv21 * qv11_inv * qv21.T

denom_q = (sympy.ones(1, 2) * qv**(-1) * sympy.ones(2, 1))**(-1)
diff_mat = sympy.Matrix([
  [x1_1 - eta1],
  [x2_1 - eta1],
  [x3_1 - eta1],
  [x4_1 - eta1],
  [x1_2 - eta2],
  [x2_2 - eta2]
])
y_mat = sympy.Matrix([[y1], [y2]])
numer_q = sympy.ones(1, 2) * qv**(-1) * (y_mat - qv21 * qv11_inv * diff_mat)
theta = (denom_q * numer_q)[0, 0]
# sympy.simplify(theta)

print(sympy.latex(sympy.simplify(theta), itex = True, mode = "inline"))

```