### 2024-02-17 src/explore/20240216-optest.qmd

---
title: "Optimal Estimation with Nuisance Parameters"
author: "Caleb Leedy"
date: "February 16, 2024"
format: 
  pdf:
    include-in-header: 
      - "latex_header.tex"
bibliography: references.bib
---

```{r}
#| label: global variables
#| echo: false
B <- 1000

obs_seg <- 250
true_mu <- 5
cov_e1e2 <- 0.5

```

# Summary

In this document, we do the following:

1. Estimate the optimal $\theta$ with a known covariance function,
2. Estimate the optimal $\theta$ with an estimated covariance function, and
3. Compare these results to other estimates.

# Simulation Setup

We consider a normal model for $(X_1, X_2, Y)$ where $X_1 \sim N(0, 1)$, and 
$e_1, e_2$ are from a bivariate normal distribution each with mean zero,
variance one, and covariance $\rho = $ `r cov_e1e2`. Then $X_2 = X_1 + e_1$ and
$Y = \mu + X_1 + e_1$ where $\mu = $ `{r} true_mu`. Each observation is
independent of the others. Instead of always observing all three variables, we
observe them with missingness. The observed segments can be viewed in @tbl-segs.
Each segment is independent of each other and of the same size. We can view each
segment as an independent simple random sample of the superpopulation of size
$n = $ `{r} obs_seg`.

| Segment | Observed Variables |
| --------| -------------------|
| $A_{11}$| $X_1, X_2, Y$      |
| $A_{10}$| $X_1, X_2$         |
| $A_{01}$| $X_1, Y$           |
| $A_{00}$| $X_1$              |
: This table shows the segment names with each variable that they contain. {#tbl-segs}

# Estimation

The goal of the simulation is to estimate $\theta = E[Y]$. We also estimate
several nuisance parameters $\eta_1 = E[X_1]$ and $\eta_2 = E[X_2]$. All of
these parameters can be estimated using GLS estimation. The model that we
propose is the following. Consider the model, for 
$g = (\eta_1, \eta_2, \theta)'$

$$ \hat f = Zg + e \text{ where } $$

\begin{align*}
f_1^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_1(x_{1i}) \\
f_2^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_2(x_{2i}) \\
f_3^{(11)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g_3(y_{i}) \\
f_1^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_1(x_{1i}) \\
f_2^{(10)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{10}}{\pi_{10}} g_2(x_{2i}) \\
f_1^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_1(x_{1i}) \\
f_3^{(01)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} g_2(y_{i}) \\
f_1^{(00)} &= n^{-1} \sum_{i = 1}^n \frac{\delta_{00}}{\pi_{00}} g_1(x_{1i}) \\
\end{align*}

and

$$\hat f = 
\begin{bmatrix}
f_1^{(11)} \\
f_2^{(11)} \\
f_3^{(11)} \\
f_1^{(10)} \\
f_2^{(10)} \\
f_1^{(01)} \\
f_3^{(01)} \\
f_1^{(00)} \\
\end{bmatrix},
Z = 
\begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
0 & 1 & 0 \\
1 & 0 & 0 \\
0 & 0 & 1 \\
1 & 0 & 0 \\
\end{bmatrix},
E[e] = 0,
\text{ and } 
\Var(e) = n^{-1}
\begin{bmatrix}
V_{11} & 0 & 0 & 0 \\
0 & V_{10} & 0 & 0 \\
0 & 0 & V_{01} & 0 \\
0 & 0 & 0 & V_{00} \\
\end{bmatrix}.
$$

We also have

$$
V_{11} = 
\begin{bmatrix}
\frac{1}{\pi_{11}} E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{11}} E[g_1g_2] -
E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{11}} E[g_2^2] -
E[g_2]^2 & \frac{1}{\pi_{11}} E[g_2g_3] - E[g_2]E[g_3] \\
\frac{1}{\pi_{11}} E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{11}} E[g_2g_3] -
E[g_2]E[g_3] & \frac{1}{\pi_{11}} E[g_3^2] - E[g_3]^2 \\
\end{bmatrix},
$$
$$
V_{10} = 
\begin{bmatrix}
\frac{1}{\pi_{10}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2]\\
\frac{1}{\pi_{10}}E[g_1g_2] - E[g_1]E[g_2] & \frac{1}{\pi_{10}}E[g_2^2] - E[g_2]^2 \\
\end{bmatrix},
$$
$$
V_{01} = 
\begin{bmatrix}
\frac{1}{\pi_{01}}E[g_1^2] - E[g_1]^2 & \frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3]\\
\frac{1}{\pi_{01}}E[g_1g_3] - E[g_1]E[g_3] & \frac{1}{\pi_{01}}E[g_3^2] - E[g_3]^2 \\
\end{bmatrix}, \text{ and }
V_{00} = 
\begin{bmatrix}
\frac{1}{\pi_{00}}E[g_1^2] - E[g_1]^2
\end{bmatrix}.
$$

Using GLS directly, yields an estimator
$\hat g = (\hat \eta_1, \hat \eta_2, \hat \theta)$. Previously, we have used 
the estimator^[The "M" is short for marginal.]

$$ \hat \theta_{M} := [0, 0, 1] \cdot \hat g. $$

However, we can do better if we assume that the distribution of $\hat g$ is a
joint multivariate normal model and compute $\hat \theta \mid \hat \eta_1, \hat
\eta_2$. We can represent this model^[The "C" is for conditional.] as 
$\hat \theta_C := E[\hat \theta \mid \hat \eta]$. For each of these models we
use the true covariance matrix and the estimated covariance matrix. For a full
discussion of the derivation of $\hat \theta_C$ see Appendix A.

# Simulation Study

```{r}
#| label: setup
#| echo: false
#| warning: false
#| message: false

library(dplyr)

library(doParallel)
library(doRNG)
library(parallelly)

# Run from src/ or src/explore/
source("R/data_generation.R")

```

```{r}
#| label: gls M, true
#| echo: false

#' This function is also used in 20240202-gls_sim.qmd. I have modified the code
#' to account for the change from X, Y1, Y2 to X1, X2, Y.
gls_m_true <- function(df, class = "basic", rho) {

  if (class == "basic") {
    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(1, 1, 1,
               1, 2, 1 + rho,
               1, 1 + rho, 2), nrow = 3)
    v_10 <- 1 / df$pi_10[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(1, 1, 1, 2), nrow = 2)
    v_00 <- 1 / df$pi_00[1]

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))

  } else {
    stop("We have only 1 class: 'basic'.")
  }
}

```

```{r}
#| label: GLS M est
#| echo: false

gls_m_est <- function(df, class = "basic") {

  if (class == "basic") {

    z_mat <- matrix(c(1, 0, 0,
                      0, 1, 0,
                      0, 0, 1,
                      1, 0, 0,
                      0, 1, 0,
                      1, 0, 0,
                      0, 0, 1,
                      1, 0, 0), ncol = 3, byrow = TRUE)

    g_hat <- 
      c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean(),
        mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean())

    # Estimate Covariance Matrix
    df_1p <- filter(df, delta_11 == 1 | delta_10 == 1)
    df_p1 <- filter(df, delta_11 == 1 | delta_01 == 1)
    df_11 <- filter(df, delta_11 == 1)
    v_x1 <- var(df$X1)
    v_x2 <- var(df_1p$X2)
    v_y <- var(df_p1$Y)
    cov_x1x2 <- cov(df_1p$X1, df_1p$X2)
    cov_x1y <- cov(df_p1$X1, df_p1$Y)
    cov_x2y <- cov(df_11$X2, df_11$Y)

    v_11 <- 1 / df$pi_11[1] *
      matrix(c(v_x1, cov_x1x2, cov_x1y,
               cov_x1x2, v_x2, cov_x2y,
               cov_x1y, cov_x2y, v_y), nrow = 3)

    v_10 <- 1 / df$pi_10[1] * matrix(c(v_x1, cov_x1x2, cov_x1x2, v_x2), nrow = 2)
    v_01 <- 1 / df$pi_01[1] * matrix(c(v_x1, cov_x1y, cov_x1y, v_y), nrow = 2)
    v_00 <- 1 / df$pi_00[1] * v_x1

    v_mat <- matrix(rep(0, 64), ncol = 8)
    v_mat[1:3, 1:3] <- v_11
    v_mat[4:5, 4:5] <- v_10
    v_mat[6:7, 6:7] <- v_01
    v_mat[8, 8] <- v_00

    v_mat <- v_mat / nrow(df)

    ghat <- MASS::ginv(t(z_mat) %*% MASS::ginv(v_mat) %*% z_mat) %*% 
      t(z_mat) %*% MASS::ginv(v_mat) %*% g_hat

    return(as.numeric(t(c(0, 0, 1) %*% ghat)))
  } else {
    stop("Only class == 'basic' has been implemented.")
  }
}

```

```{r}
#| label: GLS C true
#| echo: false

gls_c <- function(df, case = "true") {

  # Steps:
  # 1. Get fhat
  # 2. Get covariance matrix
  # 3. Solve for eta1
  # 4. Solve for eta2
  # 5. Solve for theta

  # 1. Get fhat
  fhat <- 
    c(mutate(df, tmp = delta_11 / pi_11 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_00 / pi_00 * X1) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_10 / pi_10 * X2) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_11 / pi_11 * Y) |> pull(tmp) |> mean(),
      mutate(df, tmp = delta_01 / pi_01 * Y) |> pull(tmp) |> mean()
    )

  # 2. Get covariance matrix

  if (case == "true") {

    Ex1 <- 0
    Ex2 <- 0
    Ey <- true_mu
    Ex12 <- 1
    Ex22 <- 2
    Ey2 <- true_mu^2 + 2
    Ex1x2 <- 1
    Ex1y <- 1
    Ex2y <- 1 + cov_e1e2

  } else if (case == "est") {

    Ex1 <- mean(df$X1)
    Ex2 <- 
      mutate(df, tmp = (delta_10 + delta_11) * X2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ey <-
      mutate(df, tmp = (delta_01 + delta_11) * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex12 <- mean(df$X1^2)
    Ex22 <-
      mutate(df, tmp = (delta_10 + delta_11) * X2^2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ey2 <-
      mutate(df, tmp = (delta_01 + delta_11) * Y^2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex1x2 <-
      mutate(df, tmp = (delta_10 + delta_11) * X1 * X2) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex1y <-
      mutate(df, tmp = (delta_01 + delta_11) * X1 * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()
    Ex2y <-
      mutate(df, tmp = (delta_11) * X2 * Y) %>%
      filter(tmp != 0) %>% 
      pull(tmp) %>% 
      mean()

  } else {
    stop("We have only implemented case == 'true' and 'est'.")
  }

  big_v <- matrix(0, nrow = 8, ncol = 8)
  big_v[1:4, 1:4] <- 
    diag(c(1 / df$pi_11[1] * Ex12 - Ex1^2,
           1 / df$pi_10[1] * Ex12 - Ex1^2,
           1 / df$pi_01[1] * Ex12 - Ex1^2,
           1 / df$pi_00[1] * Ex12 - Ex1^2))
  big_v[5:6, 5:6] <- 
  diag(c(1 / df$pi_11[1] * Ex22 - Ex2^2,
         1 / df$pi_10[1] * Ex22 - Ex2^2))
  big_v[7:8, 7:8] <- 
    diag(c(1 / df$pi_11[1] * Ey2 - Ey^2,
           1 / df$pi_01[1] * Ey2 - Ey^2))

  big_v[1, 5] <- 1 / df$pi_11[1] * Ex1x2 - Ex1 * Ex2
  big_v[1, 7] <- 1 / df$pi_11[1] * Ex1y - Ex1 * Ey
  big_v[5, 7] <- 1 / df$pi_11[1] * Ex2y - Ex2 * Ey
  big_v[2, 6] <- 1 / df$pi_10[1] * Ex1x2 - Ex1 * Ex2
  big_v[3, 8] <- 1 / df$pi_01[1] *  Ex1y - Ex1 * Ey

  big_v[5, 1] <- big_v[1, 5]
  big_v[7, 1] <- big_v[1, 7]
  big_v[7, 5] <- big_v[5, 7]
  big_v[6, 2] <- big_v[2, 6]
  big_v[8, 3] <- big_v[3, 8]

  # 3. Solve for eta1
  x <- t(t(c(1, 1, 1, 1)))
  v <- big_v[1:4, 1:4]
  y <- fhat[1:4]

  eta1 <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  # 4. Solve for eta2
  x <- t(t(c(1, 1)))
  v <- 
  big_v[5:6, 5:6] - big_v[5:6, 1:4] %*% MASS::ginv(big_v[1:4, 1:4]) %*% big_v[1:4, 5:6]
  y <- 
    fhat[5:6] - big_v[5:6, 1:4] %*% MASS::ginv(big_v[1:4, 1:4]) %*% 
    (fhat[1:4] - c(eta1, eta1, eta1, eta1))

  eta2 <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  # 5. Solve for theta
  x <- t(t(c(1, 1)))
  v <-
  big_v[7:8, 7:8] - big_v[7:8, 1:6] %*% MASS::ginv(big_v[1:6, 1:6]) %*% big_v[1:6, 7:8]
  y <-
    fhat[7:8] - big_v[7:8, 1:6] %*% MASS::ginv(big_v[1:6, 1:6]) %*% 
    (fhat[1:6] - c(eta1, eta1, eta1, eta1, eta2, eta2))

  theta <- 
    MASS::ginv(t(x) %*% MASS::ginv(v) %*% x) %*% t(x) %*% MASS::ginv(v) %*% y

  return(as.numeric(theta))
}

```

```{r}
#| label: MCMC
#| echo: false
#| tbl-cap: |
#|   Results from simulations study with independent equally sized segments
#|   $A_{11}$, $A_{10}$, $A_{01}$, and $A_{00}$ all of size $n = 250$. In this
#|   simulation we have the true mean of $Y_2$ equal to $\mu = 5$ and the covariance
#|   between $e_1$ and $e_2$ is $\rho = 0.5$. The goal is to estimate 
#|   $E[Y_2] = \mu$. For the estimators with 'est', we use the
#|   estimated covariance matrix $\hat V$ with 
#|   f-functions $f_1 = X_1$, $f_2 = X_2$ and $f_3 = Y$. For the estimators
#|   without we use the true covariance matrix. GLSM is the marginal model while
#|   GLSC is the conditional model.
clust <- parallel::makeCluster(min(parallelly::availableCores() - 2, 100))
registerDoParallel(clust)

mc_theta <-
  foreach(iter = 1:B,
          .options.RNG = 1,
          .packages = c("dplyr", "stringr")) %dorng% {

    # Generate Data
    df <- gen_simple_data(obs_seg, mu = true_mu, rho = cov_e1e2)

    # Estimators
    oracle <- mean(df$Y)
    cc <- df |>
      filter(delta_11 == 1) |>
      pull(Y) |>
      mean()
    ipw <- df |>
      mutate(tmp = delta_11 / pi_11 * Y) |>
      pull(tmp) |>
      mean()

    est_m_true <- gls_m_true(df, rho = cov_e1e2)
    est_m_est <- gls_m_est(df)
    est_c_true <- gls_c(df, case = "true")
    est_c_est <- gls_c(df, case = "est")

    return(tibble(Oracle = oracle,
                  CC = cc,
                  IPW = ipw,
                  GLSM = est_m_true,
                  GLSMEst = est_m_est,
                  GLSC = est_c_true,
                  GLSCEst = est_c_est
    ))

  } |>
  bind_rows()

stopCluster(clust)

# Analyze Results
mc_theta |>
  summarize(
    across(everything(),
    list(Bias = ~mean(.) - true_mu,
         SD = ~sd(.),
         Tstat = ~(mean(.) - true_mu) / sqrt(var(.) / B)),
    .names = "{.fn}_{.col}")) |> 
  tidyr::pivot_longer(cols = everything(),
                      names_to = c(".value", "Algorithm"),
                      names_pattern = "(.*)_(.*)") |>
  mutate(Pval = pt(-abs(Tstat), df = B)) |>
  knitr::kable(digits = 3)

```

# Results and Discussion

1. Estimating the covariance in our conditional model cause dependence issues.
2. This particular simulation does not differentiate between the GLS models. We
   might need a different setup to distinguish between the models (such as
   unequal segment size).

# Next Steps

1. Find the minimum variance estimator GLSC as a function of $g$.
2. Estimate the covariance matrix so that it does not interfere with estimation
   of $\hat f$.

{{< pagebreak >}}

# Appendix A: Derivation of $\hat \theta_C$

The derivation in this appendix follows the idea in @zhou2012efficient. The
initial goal is to minimize $Q(\eta_1, \eta_2, \theta)$ where

$$
Q(\eta_1, \eta_2, \theta) = z'V^{-1}z \text{ for }
z = 
\begin{bmatrix}
f_1^{(11)} - \eta_1 \\
f_1^{(10)} - \eta_1 \\
f_1^{(01)} - \eta_1 \\
f_1^{(00)} - \eta_1 \\
f_2^{(11)} - \eta_2 \\
f_2^{(10)} - \eta_2 \\
f_3^{(11)} - \theta \\
f_3^{(01)} - \theta \\
\end{bmatrix} \text{ and }
V = V(z).
$$

We can notice that $Q$ is the kernel of a multivariate normal distribution,
which means that we can decompose it into a marginal and conditional part,

$$ 
\min_{\eta, \theta} Q(\eta_1, \eta_2, \theta) = 
\min_{\eta, \theta} (Q(\eta) + Q(\theta \mid \eta)).
$$

Since $Q(\eta)$ is the kernal of a normal distribution we can apply the same
technique to get

$$ 
\min_{\eta, \theta} Q(\eta_1, \eta_2, \theta) = 
\min_{\eta, \theta} \{Q(\eta_1) + Q(\eta_2 \mid \eta_1) + Q(\theta \mid \eta_1,
\eta_2)\}.
$$

Then we can solve these sequentially.

1. We find $\hat \eta_1 = \argmin Q(\eta_1)$.
2. We can substitute this back into $Q(\eta_2 \mid \eta_1)$ to get the 
  optimal $\hat \eta_2$ of 
  $$ \hat \eta_2 = \argmin Q(\eta_2 \mid \hat \eta_1). $$
3. We can use the results of the previous two steps to get,
  $$ \hat \theta = \argmin Q(\theta \mid \hat \eta_1, \hat \eta_2).$$

Unfortunately, the nonmonotone case does not have any nice simplifications like
how the monotone case reduces to a regression estimator. This is due to the fact
that the marginal and conditional distributions are generally multivariate
normal distributions instead of a one-dimensional distribution. This causes the
result to be weighted averages instead of difference estimators.

## Finding $\hat \eta_1$

We start with the fact that

$$
\hat \eta_1 = \argmin Q(\eta_1) \text{ where } 
Q(\eta_1) = z_{1:4}' V^{-1}_{1:4, 1:4} z_{1:4}.
$$

This means that for the simulation study in this report, since each segment is
independent,

$$
Q(\eta_1) = 
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix}'
\begin{bmatrix}
(\Var(\hat f_1^{(11)}))^{-1} & 0 & 0 & 0 \\
0 & (\Var(\hat f_1^{(10)}))^{-1} & 0 & 0 \\
0 & 0 & (\Var(\hat f_1^{(01)}))^{-1} & 0 \\
0 & 0 & 0 & (\Var(\hat f_1^{(00)}))^{-1} \\
\end{bmatrix}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix}.
$$

Since the dual of this problem is a GLS estimator with a diagonal variance
matrix, $\hat \eta_1$ is the inverse-variance weighted average of all of the
individual estimators,

$$
\hat \eta_1 = 
\frac{\frac{\hat f_1^{(11)}}{\Var(\hat f_1^{(11)})} + 
\frac{\hat f_1^{(10)}}{\Var(\hat f_1^{(10)})} + 
\frac{\hat f_1^{(01)}}{\Var(\hat f_1^{(01)})} + 
\frac{\hat f_1^{(00)}}{\Var(\hat f_1^{(00)})}}{
\frac{1}{\Var(\hat f_1^{(11)})} + 
\frac{1}{\Var(\hat f_1^{(10)})} + 
\frac{1}{\Var(\hat f_1^{(01)})} + 
\frac{1}{\Var(\hat f_1^{(00)})}
}.
$$

## Finding $\hat \eta_2$

Similar to the previous step, we can notice that $Q(\eta_2 \mid \eta_1)$ is the
kernel of the conditional distribution

$$
\begin{bmatrix}
\hat f_2^{11} \\
\hat f_2^{10}
\end{bmatrix} 
\sim N \left(
\begin{bmatrix}
\eta_2 \\ \eta_2
\end{bmatrix}  + 
\Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\end{bmatrix},
\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1} \Sigma_{\eta_1, \eta_2}
\right).
$$

Since we have $\Sigma_{\eta_1} = V_{1:4, 1:4}$ and

$$
\Sigma_{\eta_2, \eta_1} = 
\begin{bmatrix}
\Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 & 0 & 0 \\
0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) & 0 & 0 \\
\end{bmatrix},
$$

we have

$$
\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1} 
\Sigma_{\eta_2, \eta_1} = 
\begin{bmatrix}
\Var(\hat f_2^{(11)}) - \frac{\Cov(\hat f_1^{(11)}, \hat{f}_2^{(11)})^2}{\Var(\hat f_1^{(11)})} & 0 \\
0 & \Var(\hat f_2^{(10)}) - \frac{\Cov(\hat f_1^{(10)}, \hat{f}_2^{(10)})^2}{\Var(\hat f_1^{(10)})} \\
\end{bmatrix}.
$$

This means that

$$
\footnotesize
\hat \eta_2 =
\left(\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}\Sigma_{\eta_1, \eta_2})^{-1} 
\begin{bmatrix} 1 \\ 1 \end{bmatrix} 
\right)^{-1}
\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\eta_2} - \Sigma_{\eta_2, \eta_1} \Sigma_{\eta_1}^{-1}\Sigma_{\eta_1, \eta_2})^{-1} 
\begin{bmatrix} 
\hat f_2^{(11)} - \frac{\Cov(\hat f_2^{(11)}, \hat f_1^{(11)})}{\Var(\hat f_1^{(11)})} (\hat f_1^{(11)} -
\hat \eta_1) \\
\hat f_2^{(10)} - \frac{\Cov(\hat f_2^{(10)}, \hat f_1^{(10)})}{\Var(\hat f_1^{(10)})} (\hat f_1^{(10)} -
\hat \eta_1) \\
\end{bmatrix}.
$$

## Finding $\hat \theta$

Finally, to find $\hat \theta$ we use a similar process as shown in the previous
step. This time it is used to compute $Q(\theta \mid \hat \eta_1, \hat \eta_2)$.
We can notice that $Q(\theta \mid \hat \eta)$ is the kernel of a conditional
multivariate normal distribution,

$$
\begin{bmatrix}
\hat f_3^{(11)} \\
\hat f_3^{(01)}
\end{bmatrix} 
\sim N \left(
\begin{bmatrix}
\theta \\ \theta
\end{bmatrix}  + 
\Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \eta_1 \\
\hat f_1^{(10)} - \eta_1 \\
\hat f_1^{(01)} - \eta_1 \\
\hat f_1^{(00)} - \eta_1 \\
\hat f_2^{(11)} - \eta_2 \\
\hat f_2^{(10)} - \eta_2 \\
\end{bmatrix},
\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1} \Sigma_{\eta, \theta}
\right).
$$

In this case we have,

$$
\Sigma_{\theta} = 
\begin{bmatrix}
V(\hat f_3^{(11)}) & 0 \\
0 & V(\hat f_3^{(10)})
\end{bmatrix},
\Sigma_{\theta, \eta} = 
\begin{bmatrix}
\Cov(\hat f_3^{(11)}, \hat f_1^{(11)}) & 0 & 0 & 0 & \Cov(\hat f_3^{(11)}, \hat
f_2^{(10)}) & 0 \\
0 & 0 & \Cov(\hat f_3^{(01)}, \hat f_1^{(01)}) & 0 & 0 & 0
\end{bmatrix},
$$

and,

$$
\Sigma_{\eta} = 
\begin{bmatrix}
V(\hat f_1^{(11)}) & 0 & 0 & 0 & \Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 \\
0 & V(\hat f_1^{(10)}) & 0 & 0 & 0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) \\
0 & 0 & V(\hat f_1^{(01)}) & 0 & 0 & 0 \\
0 & 0 & 0 & V(\hat f_1^{(00)}) & 0 & 0 \\
\Cov(\hat f_1^{(11)}, \hat f_2^{(11)}) & 0 & 0 & 0 & V(\hat f_2^{(11)}) & 0 \\
0 & \Cov(\hat f_1^{(10)}, \hat f_2^{(10)}) & 0 & 0 & 0 & V(\hat f_2^{(10)}) \\
\end{bmatrix}.
$$

Hence,

$$
\footnotesize
\hat \theta =
\left(\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}\Sigma_{\eta, \theta})^{-1} 
\begin{bmatrix} 1 \\ 1 \end{bmatrix} 
\right)^{-1}
\begin{bmatrix} 1 & 1 \end{bmatrix} 
(\Sigma_{\theta} - \Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}\Sigma_{\eta, \theta})^{-1} 
\left(
\begin{bmatrix}
\hat f_3^{(11)} \\
\hat f_3^{(01)}
\end{bmatrix}
-\Sigma_{\theta, \eta} \Sigma_{\eta}^{-1}
\begin{bmatrix}
\hat f_1^{(11)} - \hat \eta_1 \\
\hat f_1^{(10)} - \hat \eta_1 \\
\hat f_1^{(01)} - \hat \eta_1 \\
\hat f_1^{(00)} - \hat \eta_1 \\
\hat f_2^{(11)} - \hat \eta_2 \\
\hat f_2^{(10)} - \hat \eta_2 \\
\end{bmatrix}
\right).
$$

While this might still be slightly unclear, I have been able to write down the
closed form solution to this estimator, which is given below in @fig-sympy as a
function of $\eta_1$. (Including $\hat \eta_1$ made it too big for the page.)

![A closed form solution to $\hat \theta$.](sympy_opttheta.png){#fig-sympy}

{{< pagebreak >}}

# References