### 2024-02-19 references/literature/papers.qmd
Added REML papers.

---
title: "Papers Read"
author: "Caleb Leedy"
date: February 12, 2024
format: pdf
bibliography: references.bib
---

# @zhou2012efficient

Title: An efficient method of estimation for longitudinal surveys with monotone
missing data

Author(s): Ming Zhou and Jae-Kwang Kim

Journal (Year): Biometrika (2012)

Importance: I think that this paper does something very similar to our project
except with monotone data rather than nonmonotone data. The newness in the paper
is that the MAR assumption depends on variables oberseved in the previous time
period.

Innovation: They use a likelihood method for estimating the propensity score. I
am also curious about how they got Equation 9, which they claim is an optimal
estimator.

# @patterson1971recovery

Title: Recovery of inter-block information when block sizes are unequal

Author(s): H.D. Patterson and R. Thompson

Journal (Year): Biometrika (1971)

Importance: This describes REML for unbalanced designs.

Innovation: They decompose the likelihood into error constrasts and other. (These
are the $S$ and $Q$ matrices.)

# @harville1977maximum

Title: Maximum Likelihood Approaches to Variance Component Estimation and to
Related Problems

Author(s): David A. Harville

Journal (Year): JASA (1977)

Importance: This paper summarizes the use of ML and REML methods and focuses on
their use and computation.

Innovation: This is a good summary of the problems that REML solves.

Title:

Author(s):

Journal (Year):

Importance:

Innovation:

Title:

Author(s):

Journal (Year):

Importance:

Innovation:

Title:

Author(s):

Journal (Year):

Importance:

Innovation:

{{< pagebreak >}}

# References