### 2024-03-06 src/explore/20240222-sympyvarest.qmd
I showed that the conditional expectation is the optimal function.

---
title: "SymPy Representation for Optimal Regression Estimation"
author: "Caleb Leedy"
date: "February 22, 2024"
format: html
---

# Summary: Part 1

This part prints out the optimal estimator for a general $f(x)$ and for when
$f(x)  = E[Y \mid X]$.

```{python}
#| label: sympyres1
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y = sympy.symbols('f_11 f_10 y')
s11, s10, sy, sfy = sympy.symbols('s11 s10 sy sfy')

v_all = sympy.Matrix([
  [s11, 0, sfy],
  [0, s10, 0],
  [sfy, 0, sy]
])

one3 = sympy.ones(3, 1)
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y]
])

out = (one3.T * v_all**(-1) * one3)**(-1) * one3.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])

sympy.factor(4 - 4 * y - 3 * y**2 + 2 * y**3 + y**4)

```

# Summary: Part 2

We consider the model where $f(X) = E[Y \mid X] + \ell(X)$. We want to get the
variance of $\hat \mu_y$. Then we want to show

1. $V(\hat \mu_y) = a + b \mu_{\ell}$
2. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$
3. $0 = \argmin_{\sigma_{\ell, y}} V(\hat \mu_y)$

```{python}
#| label: sympyres2
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y = sympy.symbols('f_11 f_10 y')
sl, sx, sy = sympy.symbols('sl sx sy')
pcy, ply = sympy.symbols('pcy ply')

v_all = sympy.Matrix([
  [sx**2 + sl**2 + 2 * ply * sl * sy, 0, sx**2 + ply * sl * sy],
  [0, sx**2 + sl**2 + 2 * ply * sl * sy, 0],
  [sx**2 + ply * sl * sy, 0, sy**2]
])

mod_mat = sympy.Matrix([
  [1, 1],
  [1, 1],
  [1, 0]
])
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y]
])

out = (mod_mat.T * v_all**(-1) * mod_mat)**(-1) * mod_mat.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])
sympy.simplify((mod_mat.T * v_all**(-1) * mod_mat)**(-1))[0, 0]

```

# Summary: Part 3

We consider the model where $f(X) = E[Y \mid X] + \ell(X)$. We want to get the
variance of $\hat \mu_y$. Then we want to show

1. $V(\hat \mu_y) = a + b \mu_{\ell}$
2. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$
3. $0 = \argmin_{\sigma_{\ell, y}} V(\hat \mu_y)$

Unlike Part 2, in this model we have a true nonmonotone missing problem with the
following parts observed:

| Segment | $X$ | $Y$ |
| ------- | --- | --- |
|$A_{11}$ | ✔️   | ✔️   |
|$A_{10}$ | ✔️   |     |
|$A_{01}$ |     | ✔️   |

We assume that $|A_{11}| = |A_{10}| = |A_{01}|$.

```{python}
#| label: sympyres3
#| echo: false
#| output: asis

import sympy

# We need variables for what we observe:
f11, f10, y11, y01 = sympy.symbols('f_11 f_10 y11 y01')
sl, sy = sympy.symbols('sl sy')
pcy, ply = sympy.symbols('pcy ply')

v_all = sympy.Matrix([
  [pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0, pcy**2 * sy**2 + ply * sl * sy, 0],
  [0, pcy**2 * sy**2 + sl**2 + 2 * ply * sl * sy, 0, 0],
  [pcy**2 * sy**2 + ply * sl * sy, 0, sy**2, 0],
  [0, 0, 0, sy**2]
])

mod_mat = sympy.Matrix([
  [1, 1],
  [1, 1],
  [1, 0],
  [1, 0]
])
y_vec = sympy.Matrix([
  [f11],
  [f10],
  [y11],
  [y01]
])

xvinvx = mod_mat.T * v_all**(-1) * mod_mat 
xvinvx_inv = sympy.Matrix([
  [xvinvx[1, 1], -xvinvx[0, 1]],
  [-xvinvx[1, 0], xvinvx[0, 0]]
]) * 1 / (xvinvx[0, 0] * xvinvx[1, 1] - xvinvx[1, 0] * xvinvx[0, 1])

out = xvinvx_inv * mod_mat.T * v_all**(-1) * y_vec
out = xvinvx**(-1) * mod_mat.T * v_all**(-1) * y_vec
sympy.simplify(out[0, 0])
sympy.simplify(out[1, 0])
vhatmuy = sympy.simplify(xvinvx_inv[0, 0])
sympy.simplify(sympy.diff(vhatmuy, sl))

alt_vhatmuy = sy**2 * (1 + (2 * pcy**2 * sy**2 + 4 * ply * sl * sy + 2 * sl**2) * (pcy**4 * sy**2 + 2 * pcy**2 * ply * sl * sy - 4 * pcy**2 * sy**2 + ply**2 * sl**2 - 8 * ply * sl * sy - 4 * sl**2)**(-1))

# These are the same
sympy.simplify(vhatmuy - alt_vhatmuy)

sympy.simplify(sympy.diff(alt_vhatmuy, sl))

```

# Summary: Part 4

We consider the model where $f(X_1) = E[Y \mid X_1] + \ell(X_1)$, $f(X_2) = E[Y
\mid X_2]$ and $f(X_1, X_2) = E[Y \mid X_1, X_2] + \ell(X_1, X_2)$. We want to
get the variance of $\hat \mu_y$. Then we want to show

1. $0 = \argmin_{\sigma_\ell^2} V(\hat \mu_y)$ for each $\sigma_\ell$

This model we have a true nonmonotone missing problem with the
following parts observed:

| Segment | $X_1$ | $X_2$ | $Y$ |
| ------- | ----- | --- | --- |
|$A_{11}$ | ✔️     | ✔️   |     |
|$A_{10}$ | ✔️     |     | ✔️   |

We assume that $|A_{11}| = |A_{10}|$.

```{python}
#| label: sympyres3
#| echo: false
#| output: asis

import sympy

def matinv_2x2(mat):
  assert mat.shape == (2, 2)

  det_mat = mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0]
  ret = sympy.Matrix([
    [mat[1, 1], -mat[0, 1]],
    [-mat[1, 0], mat[0, 0]]
  ])

  return(ret / det_mat)

def matinv_3x3(mat):
  assert mat.shape == (3, 3)

  mat = mat.T

  adj_mat = sympy.Matrix([
    [mat[[1, 2], [1, 2]].det(), -mat[[1, 2], [0, 2]].det(),  mat[[1, 2], [0, 1]].det()],
    [-mat[[0, 2], [1, 2]].det(), mat[[0, 2], [0, 2]].det(), -mat[[0, 2], [0, 1]].det()],
    [mat[[0, 1], [1, 2]].det(), -mat[[0, 1], [0, 2]].det(),  mat[[0, 1], [0, 1]].det()],
  ])

  return(adj_mat / mat.det())

def matdet_2x2(mat):
  assert mat.shape == (2, 2)
  return(mat[0, 0] * mat[1, 1] - mat[0, 1] * mat[1, 0])

def matdet_3x3(mat):
  assert mat.shape == (3, 3)
  return(mat[0, 0] * matdet_2x2(mat[[1, 2], [1, 2]]) \
    - mat[0, 1] * matdet_2x2(mat[[1, 2], [0, 2]]) \
    + mat[0, 2] * matdet_2x2(mat[[1, 2], [0, 1]]))

def matdet_4x4(mat):
  assert mat.shape == (4, 4)
  return(mat[0, 0] * matdet_3x3(mat[[1, 2, 3], [1, 2, 3]]) \
    - mat[0, 1] * matdet_3x3(mat[[1, 2, 3], [0, 2, 3]]) \
    + mat[0, 2] * matdet_3x3(mat[[1, 2, 3], [0, 1, 3]]) \
    - mat[0, 3] * matdet_3x3(mat[[1, 2, 3], [0, 1, 2]]))

# We need variables for what we observe:
f11x1, f10x1, f11x2, f11x12, y01 = sympy.symbols('f11x1 f10x1 f11x2 f11x12 y01')
sx1, sx2, sx12, sl1, sl2, sl12, sy = sympy.symbols('sx1 sx2 sx12 sl1 sl2 sl12 sy')
pl1y, px1x2, px1l2, px2l1 = sympy.symbols('pl1y px1x2 px1l2 px2l1')
pl1l2, pl2y, pl12y = sympy.symbols('pl1l2 pl2y pl12y')
px1x12, px1l12, px12l1, pl1l12 = sympy.symbols('px1x12 px1l12 px12l1 pl1l12')
px2x12, px2l12, px12l2, pl2l12 = sympy.symbols('px2x12 px2l12 px12l2 pl2l12')

c01 = px1x2 * sx1 * sx2 + px2l1 * sx1 * sl2 + px2l1 * sx2 * sl1 + pl1l2 * sl1 * sl2
c02 = px1x12 * sx1 * sx12 + px1l12 * sx1 * sl12 + px12l1 * sx12 * sl1 + pl1l12 * sl1 * sl12
c12 = px2x12 * sx2 * sx12 + px2l12 * sx2 * sl12 + px12l2 * sx12 * sl2 + pl2l12 * sl2 * sl12

v_11 = sympy.Matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, c01, c02],
  [c01, sx2**2 + sl2**2 + 2 * pl2y * sl2 * sy, c12],
  [c02, c12, sx12**2 + sl12**2 + 2 * pl12y * sl12 * sy]
])
v_10 = sympy.Matrix([
  [sx1**2 + sl1**2 + 2 * pl1y * sl1 * sy, sx1**2 + pl1y * sl1 * sy],
  [sx1**2 + pl1y * sl1 * sy, sy**2]
])

v_all = sympy.diag(v_11, v_10)

mod_mat = sympy.Matrix([
  [1, 1, 0, 0],
  [1, 0, 1, 0],
  [1, 0, 0, 1],
  [1, 1, 0, 0],
  [1, 0, 0, 0]
])
y_vec = sympy.Matrix([
  [f11x1],
  [f11x2],
  [f11x12],
  [f10x1],
  [y01]
])

v_inv = sympy.diag(matinv_3x3(v_11), matinv_2x2(v_10))
xvinvx = mod_mat.T * v_inv * mod_mat

# xvinvx_inv = matinv_4x4(xvinvx)

# out = xvinvx_inv * mod_mat.T * v_inv * y_vec
# sympy.simplify(out[0, 0])
# vhatmuy = sympy.simplify(xvinvx_inv[0, 0])
vhatmuy = matdet_3x3(xvinvx[[1, 2, 3], [1, 2, 3]]) / matdet_4x4(xvinvx)
sympy.simplify(vhatmuy)
sympy.simplify(sympy.diff(vhatmuy, sl))

```