### 2024-03-06 src/explore/20240223-optvar.qmd

---
title: "Proof of Optimal Variance Estimators"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    grid:
      margin-width: 450px
bibliography: references.bib
comments:
  hypothesis: true
---

\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\argmin}{\text{argmin}}

# Summary

This report contains my current progress in establishing the optimal variance
estimators for the nonmonotone GLS estimation.

# General Notation

For now, we assume that we have a superpopulation model and a sampling
model. I would first like to establish optimality results using the simpliest
models, but it is not always clear what is simple in the nonmonotone case.
If we also consider a monotone model, we can have simple models based on if we
want to have complete cases (observations in which every variable is observed),
which I denote `CC`,
or if we want to have complete variables (variables which are never missing),
which I denote `CV`.
This leads to the following set of simple cases:

::: {#tbl-panel layout-ncol=4}
| $X$ | $Y$ |
|---|---|
| ✔️ | ✔️ |
| :heavy_check_mark: |  |
: Monotone {#tbl-one}

| $X$ | $Y$ |
|---|---|
| ✔️ | ✔️ |
| :heavy_check_mark: |  |
|  | ✔️ |
: Nonmonotone: CC {#tbl-two}

| $X_1$ | $X_2$ | $Y$ |
|-------|-------|---|
| ✔️     | ✔️     |   |
| ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: CV {#tbl-three}

| $X_1$ | $X_2$ | $Y$ |
|-------|-------|---|
| ✔️     | ✔️     | :heavy_check_mark: |
| ✔️     | ✔️     |   |
| ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: CC + CV {#tbl-four}

Set of Simple Monotone and Nonmonotone Cases
:::

For each of these cases, we can construct a GLS estimator from the corresponding
missingness model (I am giving the
monotone model in @tbl-one) like the following:

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

# Goals

I would like to establish the following:

1. For the corresponding missingness model, I would like to find all of the 
   $f(X)$ with $E[f(X)] = \mu_f$ such that the final estimator of $\hat \mu_y$
   has minimal variance.
2. Find the conditions under which the GLS estimator based on a superpopulation
   model is design consistent.

The second task is proving a result that is very similar to 
[@fuller2009sampling, Theorem 2.3.1].

# Results

::: {#thm-mono}
Assume that we have the simple monotone case in @tbl-one with independent
segments $A_{11}$ and $A_{10}$ having a sample size of $n_{11}$ and $n_{10}$
respectively. We assume that $n_{11} = n_{10}$. Then for GLS estimators in the
form,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\end{bmatrix},
V
\right)
$$

where 
$\hat f_{11}(X) = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}f(x_i)}{\pi_{11i}}$,
$\hat f_{10}(X) = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}f(x_i)}{\pi_{10i}}$,
$\hat Y_{11} = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}y_i}{\pi_{11i}}$,
and the function $f(X) = E[Y \mid X]$ minimizes the variance of $\hat \mu_Y$.

:::

::: {.proof}
The goal is to show that

$$ E[Y \mid X] = \argmin_{f} \hat \mu_Y(f). $$

Our approach is the following. Consider $f(X) = E[Y \mid X] + \ell(X)$ where
$E[\ell(X)] = 0$. With this formulation of $f(X)$ we can compute the variance
matrix $V$. Define the following notation:

| Label | Symbol |
| :----- | ------: |
| $\Var(E[Y \mid X])$ | $\sigma_x^2$ |
| $\Var(Y)$ | $\sigma_y^2$ |
| $\Var(\ell(X))$ | $\sigma_\ell^2$ |
| $\Cov(E[Y \mid X], Y)$ | $\rho_{xy} \sigma_x \sigma_y$ |
| $\Cov(E[Y \mid X], \ell(X))$ | $\rho_{x\ell} \sigma_x \sigma_\ell$ |
| $\Cov(Y, \ell(X))$ | $\rho_{y\ell} \sigma_y \sigma_\ell$ |
: A table to define the notation used in the rest of the proof. {#tbl-lab1}

Notice that,

$$ \Cov(E[Y \mid X], Y) = E[E[Y \mid X]^2] - E[Y]^2 = \Var(E[Y \mid X]) $$

and

$$ \Cov(E[Y \mid X], \ell(X)) = E[E[Y \mid X]\ell(X)] = \Cov(Y, \ell(X)). $$

This means that $\sigma_x = \rho_{xy} \sigma_y$ and 
$\rho_{x\ell} \sigma_x = \rho_{y\ell} \sigma_y$. Hence, we have the following
covariance matrix $V$,

[The constant in front of the matrix is equal to $n^{-1}$ because $n_{11} = n_{10}$. Since each segment is chosen via simple random sampling, the overall variance of 
$\hat f_{11}(X)$ is $(n_{11}^{-1} - n^{-1})S_{f_{11}}^2 = (2/n - 1/n)S_{f_{11}}^2
= S^2_{f_{11}}/n$. Since $n_{11} = n_{10}$ the variance of $\hat f_{10}(X)$ is
also equal to $S_{f_{10}}^2 / n$. This holds for all of the variance and
covariance terms which allows us to factor it out of the matrix.]{.aside}

$$V =  \frac{1}{n}
\begin{bmatrix}
\rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0
& \rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell \\
0 & \rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0 \\
\rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0 & \sigma_y^2
\end{bmatrix}.
$$

The goal is now to show that minimizing the variance of $\hat \mu_Y$ yields
$\sigma_\ell^2 = 0$. Using GLS results in a variance estimate of $\hat \mu$ of

$$ V(\hat \mu) = (X'V^{-1}X)^{-1}$$

where

$$ X = 
\begin{bmatrix}
1 & 1 \\
1 & 1 \\
1 & 0
\end{bmatrix}
\text{ for }
\mu = 
\begin{bmatrix}
\mu_Y \\ \mu_\ell
\end{bmatrix}.
$$

Evaluating this matrix expression and then viewing the first entry yields an
expression of the variance of $\mu_Y$,

$$\Var(\mu_Y) = n^{-1} \sigma_y^2 \left(1 + 
\frac{\rho_{xy}^4 \sigma_y^2 + 2\rho_{y\ell}\sigma_\ell \sigma_y \rho_{xy}^2 +
\rho_{y\ell}^2\sigma_{\ell}^2}{2(\rho_{xy}^2\sigma_y^2 + 2\rho_{y\ell}\sigma_\ell
\sigma_y + \sigma_\ell^2)}\right).$$

Then we can solve the equation,

$$ \frac{\partial V(\hat \mu_Y)}{\partial \sigma_\ell} = 0. $$

This yields (after some cancelations), the equation

$$ \sigma_\ell \rho_{y\ell} \sigma_y (\rho_{xy}^2 (\sigma_y - \sigma_\ell) +
\rho_{y\ell}^2 (1 - \rho_{xy}\sigma_y)) = 0.$$

Hence, the candidates for the minimum value of $V(\hat \mu_Y)$ are ($\rho_{y\ell}
= 0$, $\sigma_\ell = \sigma_y$), $\rho_{y\ell} = 0$, and $\sigma_\ell = 0$.
Plugging each of these back into $V(\hat \mu_Y)$ shows that $\sigma_\ell = 0$ is
actually the minimum. Hence, $\sigma_\ell^2 = 0$ yields the minimum variance of
$V(\hat \mu_Y)$ which means that since $E[\ell(X)] = 0$ that $\ell(X) = 0$. This
means that the optimal function of $f(X)$ is $E[Y \mid X]$. :black_medium_square:

:::

We can use the same technique to get a result for the nonmonotone case where we
have complete cases.

::: {#thm-nmcc}
Assume that we have the simple monotone case in @tbl-two with independent
segments $A_{11}$, $A_{10}$, and $A_{01}$ having sample sizes of $n_{11}$,
$n_{10}$, and $n_{01}$ respectively. We assume that $n_{11} = n_{10} = n_{01}$.
Then for GLS estimators in the form,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11} \\
\hat Y_{10}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_X \\
\mu_X \\
\mu_Y \\
\mu_Y \\
\end{bmatrix},
V
\right)
$$

where 
$\hat f_{11}(X) = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}f(x_i)}{\pi_{11i}}$,
$\hat f_{10}(X) = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}f(x_i)}{\pi_{10i}}$,
$\hat Y_{11} = n_{11}^{-1} \sum_{i = 1}^n \frac{\delta_{11i}y_i}{\pi_{11i}}$, 
$\hat Y_{10} = n_{10}^{-1} \sum_{i = 1}^n \frac{\delta_{10i}y_i}{\pi_{10i}}$,
and the function $f(X) = E[Y \mid X]$ minimizes the variance of $\hat \mu_Y$.

:::

::: {.proof}
This proof follows the same approach as @thm-mono. Using the same notation, we
have the following GLS model,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11} \\
\hat Y_{10} \\
\end{bmatrix} \sim
\left(
\begin{bmatrix}
1 & 1 \\
1 & 1 \\
1 & 0 \\
1 & 0 \\
\end{bmatrix},
\begin{bmatrix}
\mu_Y \\
\mu_\ell
\end{bmatrix}, \frac{2}{n}
\begin{bmatrix}
\rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0
& \rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0\\
0 & \rho_{xy}^2\sigma_y^2 + \sigma_\ell^2 + 2 \rho_{y\ell} \sigma_y \sigma_\ell & 0 & 0 \\
\rho_{xy}^2\sigma_y^2 + \rho_{y\ell} \sigma_y \sigma_\ell & 0 & \sigma_y^2 & 0 \\
0 & 0 & 0 & \sigma_y^2 \\
\end{bmatrix}.
\right)
$$

[The constant in front of the matrix is equal to $2/n$ because $n_{11} = n_{10}
= n_{01}$. Since each segment is chosen via simple random sampling, the overall
variance of 
$\hat f_{11}(X)$ is $(n_{11}^{-1} - n^{-1})S_{f_{11}}^2 = (3/n - 1/n)S_{f_{11}}^2
= 2S^2_{f_{11}}/n$. This holds for all of the variance and
covariance terms which allows us to factor it out of the matrix.]{.aside}

The variance of $\hat \mu_Y$ is

$$V(\hat \mu_Y) = 2n^{-1} \sigma_y^2 \left(1 + 
\frac{2 \rho_{xy}^2 \sigma_y^2 + 4 \rho_{y\ell} \sigma_\ell \sigma_y +
2\sigma_\ell^2}{\rho_{xy}^4 \sigma_y^2 + 2 \rho_{xy}^2 \rho_{y\ell} \sigma_\ell
\sigma_y - 4\rho_{xy}^2 \sigma_y^2 + \rho_{y\ell}^2 \sigma_\ell^2 - 8
\rho_{y\ell}\sigma_{\ell} \sigma_{y} - 4 \sigma_{\ell}^2}
\right).$$

Solving $\frac{\partial V(\hat \mu_Y)}{\partial \sigma_y} = 0$ yields (after
simplification),

$$ \sigma_{\ell} \sigma_y (\rho_{xy}^4 \sigma_y + \rho_{xy}^2 \rho_{y\ell}(\sigma_\ell
- \rho_{y\ell}\sigma_y) - \rho_{y\ell}^3 \sigma_\ell) = 0. $$

This implies that the minimum variance occurs at $\sigma_{\ell} = 0$ which means
that $\ell(X) = 0$. Hence, the optimal value of $f(X) = E[Y \mid X]$.
:black_medium_square:

:::

# Discussion

This current approach seems to be working. However, I am using SymPy [@meurer2017sympy]
to assist with and check the computational algebra. I have started to find the
optimal functions of $f(X_1), f(X_2)$, and $f(X_1, X_2)$ from @tbl-three but
inverting a $4 \times 4$ matrix is difficult to approach computationally. Since
this is very similar to the semiparametric literature, I am going to try to see
if I can use some of their techniques to simplify my work.

<!---
# Current Progress

The first thing that I have tried to consider is the optimal function $f(X)$ in
the monotone case (@tbl-one). For this missingness set, we have the GLS model:

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_f \\
\mu_f \\
\mu_y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

::: {.column-margin}

I also compared this result to the case in which we have $E[f(X)] = \mu_y$ which
occurs if $f(X) = E[Y \mid X]$. In this case, we have the model

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_y \\
\mu_y \\
\mu_y \\
\end{bmatrix},
\begin{bmatrix}
\Sigma_{11} & 0 & \Sigma_{fy} \\ 
0 & \Sigma_{10} & 0 \\ 
\Sigma_{yf} & 0 & \Sigma_y \\
\end{bmatrix}
\right).
$$

Then using GLS (it is not too hard to invert the covariance matrix because it is
a block diagonal), we get

$$
\hat \mu_y =
\frac{\hat f_{10}(X)(\Sigma_{11}\Sigma_y - \Sigma_{fy}^2) - \hat f_{11}(X)
\Sigma_{10}(\Sigma_{fy} - \Sigma_y) + Y_{11} \Sigma_{10}(\Sigma_{11} - \Sigma_{fy})}{
\Sigma_{10}(\Sigma_{11} - 2\Sigma_{fy} + \Sigma_y) + \Sigma_{11}\Sigma_y - \Sigma_{fy}^2
}.
$$

The variance of this is

$$
\Var(\mu_y) = 
\left(\frac{\Sigma_{11} - 2\Sigma_{fy} + \Sigma_y}{\Sigma_{11}\Sigma_y -
\Sigma_{fy}^2} + \frac{1}{\Sigma_{10}}\right)^{-1}.
$$

We can notice that if we let all of the variances 
($\Sigma_{11}, \Sigma_{10}, \Sigma_y = 1$) then we can compare this variance
result to @eq-condvarymono. If this case this marginal variance is always lower
than @eq-condvarymono unless $\Sigma_{fy} = 1$. In this case they are the same.

However, I do not think that this is a problem because in this marginal model we
are assuming that $f(X)$ is correctly specified for $E[Y \mid X]$. 
**Is this correct Dr. Kim?**

:::

Using the technique of @zhou2012efficient, we can first solve for the optimal
estimator of $\mu_f$, which is

$$
\hat \mu_f = 
\frac{\frac{\hat f_11(X)}{\Sigma_{11}} + \frac{\hat f_10(X)}{\Sigma_{10}}}{
\Sigma_{11}^{-1} + \Sigma_{10}^{-1}}.
$$

Then the conditional distribution of $Y_{11}$ given $f(X)$ is

$$
Y_{11} \mid \hat f(X) \sim \left(
\mu_y + \Sigma_{yf}\Sigma_{11}^{-1}(\hat f_11(X) - \hat \mu_f),
\Sigma_y - \Sigma_{yf} \Sigma_{11}^{-1} \Sigma_{fy}\right).
$$

This means that

$$
\hat \mu_y = Y_{11} - 
\left(\frac{\Sigma_{yf}}{\Sigma_{11} + \Sigma_{10}}\right)
(\hat f_{11}(X) - \hat f_{10}(X))
$$

and this has a variance of

$$ \Sigma_{y} - \frac{\Sigma_{yf}^2}{\Sigma_{11} + \Sigma_{10}}. $$ {#eq-condvarymono}

From semiparametric theory, I think that this result should be expected because
$\Sigma_{11} + \Sigma_{10} = \Var(\hat f_{11}(X) + \hat f_{10}(X))$. This does
generate follow up questions for me:

1. It seems like we want to find the function $f(X)$ such that $|\Sigma_{fy}|$ 
   is maximized. Does this mean that $f(X)$ should be $E[Y \mid X]$?
2. In this analysis, I am not using the superpopulation model directly. Is the
   only purpose of a superpopulation model to assert that the GLS estimator is
   optimal for a proposed model?

--->

# References