### 2024-03-06 src/explore/20240301-sagescore.qmd

---
title: "SymPy Representation for Efficient Score Functions"
author: "Caleb Leedy"
date: "March 1, 2024"
format: html
jupyter:
  kernel: sagemath
---

# Summary: Part 1

This part determines the efficient score function for the simple monotone model.

```{python}
#| label: sage score 1

from sage.symbolic.integration.integral import definite_integral

f11, f10, y = var('f11 f10 y')
muy, muf, sy, sf, pfy = var('muy muf sy sf pfy')

loglikefun = -3/2 * log(2 * pi) - 1/2 * log(sy^2 * (1 - pfy^2)) - 1 / (2 * sy^2
* (1 - pfy^2)) * (y - muy - (pfy * sy / sf) * (f11 - muf))^2 - log(sf^2) - 1/2 * (
  (f11 - muf)^2 + (f10 - muf)^2) / sf^2

S_muy = loglikefun.derivative(muy)
S_muf = loglikefun.derivative(muf)
S_sy = loglikefun.derivative(sy)
S_sf = loglikefun.derivative(sf)
S_pfy = loglikefun.derivative(pfy)

v_mat = matrix([
  [sf^2, 0 , pfy * sf * sy],
  [0, sf^2, 0],
  [pfy * sf * sy, 0, sy^2]
])

mean_diff = matrix([
  [f11 - muf],
  [f10 - muf],
  [y - muy]
])

dens_fy = (2 * pi)^(3/2) * (v_mat.det())^(-1/2) * exp(-1 / 2 * mean_diff.transpose() * v_mat.inverse() * mean_diff)[0, 0]

dens_y = (2 * pi * sy^2)^(-1/2) * exp(-1/(2 * sy^2) * (y - muy)^2)

res = definite_integral(dens_y, y, -infinity, infinity)
res.full_simplify()
```