### 2024-03-06 src/explore/20240304-altoptvar.qmd

---
title: "Alternate: Proof of Optimal Variance Estimators"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

\newcommand{\Var}{\text{Var}}
\newcommand{\argmin}{\text{argmin}}

# Summary

This report contains my current progress in establishing the optimal variance
estimators for the nonmonotone GLS estimation. I think that I got it.

# Report (Monday, March 3)

Previously^[See 20240223-optvar.html.] we argued that the conditional
expectation of $Y$ on the observed variables is the optimal form of $f(X)$. We
showed this by allowing $f(X) = E[Y \mid X] + \ell(X)$ where $E[\ell(X)] = 0$.
Then the minimal variance estimator of $E[Y]$ yielded $\Var(\ell(X)) = 0$, which
indicated that the optimal function of $f(X) = E[Y \mid X]$. Unfortunately, the
details in this proof made it difficult to scale beyond the case of having two
variables, and so we sought an alternative method which could scale better.

One approach is to use semiparametric theory for parametric models. In the
monotone case, the problem that we want to solve is to find the functional form
of $f(X)$ such that we have the minimum variance of $\hat \beta = E[Y]$ for the
following model, which holds if $n_{11} = n_{10}$,

$$
\begin{bmatrix}
\hat f_{11}(X) \\
\hat f_{10}(X) \\
\hat Y_{11}
\end{bmatrix} \sim
\left(
\begin{bmatrix}
\mu_f \\
\mu_f \\
\mu_y \\
\end{bmatrix},
n^{-1}
\begin{bmatrix}
\sigma_f^2 & 0 & \rho_{fy} \sigma_f \sigma_y \\ 
0 & \sigma_f^2 & 0 \\ 
\rho_{fy} \sigma_f \sigma_y & 0 & \sigma_y^2 \\
\end{bmatrix}
\right).
$$

We now have a parametric model with $\beta = \mu_y$ and $\eta = (\mu_f,
\sigma_y, \sigma_f, \rho_{fy})^T$. Previously, we were thinking about how the
choice of $f(X)$ determines $\mu_f$, $\sigma_f$, and $\rho_{fy}$, which is true.
However, since these parameters are the only things that we need for the model,
we can find the efficient influence function by letting $\theta = (\beta,
\eta^T)^T$ and then we have,

$$ \varphi_{eff} = \left(\frac{\partial \theta}{\partial \beta}\right)^{-1}
E[S_\theta S_\theta^T]^{-1} S_\theta.$$

This looks promising. However, the inverse information matrix, $E[S_\theta
S_\theta^T]^{-1}$ is very difficult to compute. Even for the monotone case that
we currently consider it is a 5x5 matrix.^[Technically, it is a block diagonal
matrix with blocks of size 1 and size 4. The 4x4 block is very challenging
because it requires the actual expectations of difference squared terms.] This
means that this technique also does not generalize well.

We have learned that we can work with the model directly, but this will require
slightly more advanced techniques like something similar to the partial linear
model (PLM).

# Update (Tuesday, March 4)

I think that I have a good method of proof. However, I want to check to make
sure that the details are correct. I argue that the optimal function for $f(X)$
in the GLS estimator is $f(X) = E[Y \mid X]$. It takes three steps:

1. We argue that $\alpha(X) = E[Y \mid X]$ is optimal in the PLM,
2. Then we show that the PLM holds for each independent segment, and
3. Finally, we combine the segment estimators using the variance weighted
   average.

::: {#lem-plm}
Let $Y$ and $X$ be random variables that are possibly vector valued with finite
variance, and denote $\mu_Y = E[Y]$. Then, for

$$ \alpha^*(X) = \argmin_{\alpha(X)} E[(Y - \mu_Y - \alpha(X))^2] \text{ such
that } E[\alpha(X)] = 0$$

we have $\alpha^*(X) = E[Y \mid X] - \mu_Y$.
:::

::: {.proof}
Let $\hat \alpha(X) = E[Y \mid X] - \mu_Y$. Notice, that

$$ E[\hat \alpha(X)] = E[E[Y \mid X]] - \mu_Y = 0. $$

Furthermore, let $\ell(X)$ satisfy $\alpha(X) = \hat \alpha(X) + \ell(X)$. Then,

\begin{align*}
E[(Y - \mu_y - \alpha(X))^2] 
&= E[(Y - \mu_y - \hat \alpha(X) - \ell(X))^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[(Y - \mu_Y - \hat \alpha(X))\ell(X)]
+ E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[E[(Y - E[Y \mid X])\ell(X)
\mid X]] + E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] - 2 E[(E[Y \mid X] - E[Y \mid X])\ell(X)] +
E[\ell(X)^2] \\
&= E[(Y - \mu_y - \hat \alpha(X))^2] + E[\ell(X)^2].
\end{align*}

Hence, $\alpha^*(X) = \hat \alpha(X) = E[Y \mid X] - \mu_Y$.
:::

## Step 1

Now consider the following partial linear model

$$Y = \mu_Y + \alpha(X) + \varepsilon$$

where $E[\alpha(X)] = 0$.^[Note that this partial linear model is different than
the one used by @robinson1988root.] By @lem-plm, the function of $\alpha$ that
minimizes the variance of $\varepsilon$ is $\alpha(X) = E[Y \mid X] - \mu_Y$.

## Step 2

Consider a single segment where we observe $Y$ and some variables $X$ where $X$
could be a vector. We consider the sample superpopulation model,

$$
\begin{bmatrix} X \\ Y \end{bmatrix} \sim \left(
\begin{bmatrix} \mu_X \\ \mu_Y \end{bmatrix},
\begin{bmatrix} 
\Sigma_{XX} & \Sigma_{XY} \\ \Sigma_{YX} & \sigma_Y^2
\end{bmatrix}
\right).
$$

The goal is to find the function of $X$, $f(X)$, such that we have the minimum
variance. Mathematically, this is

$$f^*(X) = \argmin_{f(X)} \Var(Y - \mu_Y - f(X)).$$

For identification purposes we also require that $E[f(X)] = 0$. This makes the
problem the same as Step 1, in which case the optimal form of $f(X)$ is $E[Y
\mid X] - \mu_Y$.^[If our model is assumed to be normal, then we can see what
the form of $f(X)$ is from the conditional distribution: $f(X) =
\Sigma_{YX}\Sigma_{XX}^{-1}(X - \mu_X)$.] Note that we define the
optimal form of $f(X)$ in the sample *superpopulation* model. We will worry
about estimating $f(X)$ later.

## Step 3

The overall goal is to find the optimal functional form of $f(G_r(X))$ where
$G_r(X)$ is the set of $X$ variables observed in the $r$-th segment. However,
since we assume that each segment is independent of the rest, and that $X$ and
$Y$ follow a normal model, by the Gauss-Markov theorem, the inverse-variance
weighted average of each segment contains best linear unbiased estimator of
$\theta = E[Y]$. This means that if we have the problem,

$$
\begin{bmatrix} f(G_0(X)) \\ f(G_1(X)) \\ \vdots \\ Y_0 \\ \vdots \end{bmatrix} \sim \left(
\begin{bmatrix} \mu_{G_0(X)} \\ \mu_{G_1(X)} \\ \vdots \\ \mu_Y \\ \vdots \end{bmatrix},
V
\right).
$$

The optimal form of $f(G_r(X))$ is $E[Y \mid G_r(X)] - \mu_Y$.

# Questions

1. I am not totally sure how to cover segments that do not contain $Y$. Should I
   say that the best estimator for $Y - \mu_Y$ is $E[Y \mid X] - \mu_Y$?

# Next Steps

* Try estimating the variance with REML so that we have unbiased estimation.
* Extend proof to semiparametric case where we do not necessarily assume a
normal model.
* Find the optimal estimator under MAR instead of missing by design.