### 2024-04-11 src/explore/20240313-eststeps.qmd

---
title: "Steps in Estimation Process of Optimal GLS Estimator"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

```{r}
#| label: global variables
B <- 1000

obs_seg <- 250
true_mu <- 5
rho <- 0.5

```

```{r}
#| label: setup

library(dplyr)

```

\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}

# Summary

For the nonmonotone GLS estimation we need to estimate the following:

1. The functional form of $f(G_r(X)) = E[Y \mid G_r(X)]$,
2. The variance, $\hat{V}(E[Y \mid G_r(X)])$,
3. The function, $\hat E[Y \mid G_r(X)]$, and 
4. The goal, $\hat \theta = E[Y]$.

In [Optimal Estimation with Nuisance Parameters](20240216-optest.qmd), we tested
to see if we could
get an unbiased estimate of $\hat \theta$ when we estimated the covariance
matrix. This report is similar. We compare the following:

* The Oracle estimator of $\hat \theta$ $ = E[Y] = n^{-1} \sum_{i = 1}^n y_i$,
* The Complete case estimator of $\hat \theta$,
* The IPW estimator of $\hat \theta$,
* The GLS estimate with correctly specified $E[Y \mid G_r(X)]$, and correctly
specified $V(E[Y \mid G_r(X)])$,
* The GLS estimate with estimated $\hat E[Y \mid G_r(X)]$, and correctly
specified $V(E[Y \mid G_r(X)])$, and 
* The GLS estimate with estimated $\hat E[Y \mid G_r(X)]$, and estimated
$\hat V(E[Y \mid G_r(X)])$.

These estimators are compared to each other under the nonmonotone complete case
and variable setting.

# Simulation Setup

## Nonmonotone Complete Case and Variable

For this nonmonotone case, we have three segments $A_1$, $A_2$, and $A_3$ with
missing patterns shown in @tbl-four.

| Segment | $X_1$ | $X_2$ | $Y$ |
|----|-------|-------|---|
| $A_1$ | ✔️     | ✔️     | :heavy_check_mark: |
| $A_2$ | ✔️     | ✔️     |   |
| $A_3$ | ✔️     |       | :heavy_check_mark: | 
: Nonmonotone: Complete Case + Complete Variable {#tbl-four}

The estimating vector is the following

$$
\begin{bmatrix} 
\hat f_1(X_1) \\ 
\hat f_1(X_2) \\ 
\hat f_1(X_1, X_2) \\ 
\hat Y_1 \\ 
\hat f_2(X_1) \\ 
\hat f_2(X_2) \\ 
\hat f_2(X_1, X_2) \\ 
\hat f_3(X_1) \\ 
\hat Y_3 
\end{bmatrix} =
\begin{bmatrix}
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{1i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{2i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}E[Y \mid x_{1i}, x_{2i}]}{\pi_{1i}} \\
n_1^{-1} \sum_{i = 1}^n \frac{\delta_{1i}y_i}{\pi_{1i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{1i}]}{\pi_{2i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{2i}]}{\pi_{2i}} \\
n_2^{-1} \sum_{i = 1}^n \frac{\delta_{2i}E[Y \mid x_{1i}, x_{2i}]}{\pi_{2i}} \\
n_3^{-1} \sum_{i = 1}^n \frac{\delta_{3i}E[Y \mid x_{1i}]}{\pi_{3i}} \\
n_3^{-1} \sum_{i = 1}^n \frac{\delta_{3i}y_i}{\pi_{3i}} \\
\end{bmatrix}.
$$

## Implementation Details

For this setup, we consider segments of equal size 
$n_1 = n_2 = n_3 =$ `{r} obs_seg` and a
data generating process of

$$
\begin{bmatrix} X_1 \\ \varepsilon_1 \\ \varepsilon_2 \end{bmatrix} \sim 
\left(
\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix},
\begin{bmatrix} 
1 & 0 & 0 \\ 
0 & 1 & \rho \\ 
0 & \rho & 1 \\
\end{bmatrix}
\right)
$$

in which $X_2 = X_1 + \varepsilon_1$ and $Y = \mu_Y + X_1 + \varepsilon_2$.
While we can vary the parameters of $\mu_Y$ and $\rho$, for this study we have,
$\mu_Y =$ `{r} true_mu`, and $\rho =$ `{r} rho`.

```{r}
#| label: data generating process

#' A data generation function to create a simple random sample of equally sized
#' segments of X_1, X_2, and Y. There variables are correlated but all normal.
#'
#' @param n_obs_seg - An integer greater than zero. This is the number of
#' observations in each segment
#' @param mu - The true mean of Y
#' @param rho - The correlation between (X_1 - X_2) and (Y - X_1).
#'
#' @details We always observe X1 but X2 and Y are only observed depending on the
#' value of \delta_{ij}. If i = 1 then X2 is observed. If j = 1 then Y is
#' observed.
gen_simple_data <- function(n_obs_seg, mu, rho) {

  # This code is modified from the case where we had X, Y_1, and Y_2. It only
  # changes at the end to X_1, X_2, and Y.
  # A_11
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_11 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 0, delta_11 = 1,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_10
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_10 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 1, delta_01 = 0, delta_11 = 0,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_01
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  df_01 <- 
    tibble(X1 = x, X2 = y1, Y = y2,
           delta_00 = 0, delta_10 = 0, delta_01 = 1, delta_11 = 0,
           pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)
  # A_00
  x <- rnorm(n_obs_seg)
  e_mat <- MASS::mvrnorm(n_obs_seg, c(0, 0), matrix(c(1, rho, rho, 1), nrow = 2))
  e1 <- e_mat[, 1]
  e2 <- e_mat[, 2]
  y1 <- x + e1
  y2 <- mu + x + e2

  # No empty segment
  # df_00 <- 
  #   tibble(X1 = x, X2 = y1, Y = y2,
  #          delta_00 = 1, delta_10 = 0, delta_01 = 0, delta_11 = 0,
  #          pi_00 = 0.25, pi_10 = 0.25, pi_01 = 0.25, pi_11 = 0.25)

  df <- bind_rows(df_11, df_10, df_01)

  # Return
  # X, Y_1, Y_2,
  # delta_00, delta_10, delta_01, delta_11,
  # pi_00, pi_10, pi_01, pi_11
  return(df)

}

```

## Estimation Details

We assume that we always use the correct functional form of $E[Y \mid G_r(X)]$.
The correctly specified and estimated forms of $f(G_r(X))$ are given in
@tbl-meanform

| Function | Correctly Specified Form | Estimated Form |
| -------- | :----------------------: | :------------: |
| $E[Y \mid X_1]$ | $\hat \mu + x_1$  | $\hat \beta_0 + \hat \beta_1 x_1$ |
| $E[Y \mid X_2]$ | $\hat \mu + \left(\frac{1 + \rho}{2}\right)x_2$ | $\hat \beta_0 + \hat \beta_2 x_2$ |
| $E[Y \mid X_1, X_2]$ | $\hat \mu + x_1 + \rho (x_2 - x_1)$ | $\hat \beta_0 +
\hat \beta_1 x_1 + \hat \beta_2 x_2$ |
: {#tbl-meanform}

In @tbl-meanform, the estimation of $\hat \mu$ for the correctly specified form
is always carried out as the mean of a difference. For example, to estimate 
the $\hat \mu$ in $E[Y \mid X_1]$, we use

$$ \hat \mu = \frac{\sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2i}}(y_i -
x_{1i})}{\sum_{i = 1}^n \frac{\delta_{2i}}{\pi_{2i}}}. $$

The estimation of $\beta_i$ in the `Estimated Form` column of @tbl-meanform, is
computed using least squares on all of the observations containing all of the
variables $G_r(X)$ and $Y$.

Likewise, we have a similar result for the covariance matrix in @tbl-covform.

| Covariance | Correctly Specified Form | Estimated Form |
| ---------- | :----------------------: | :------------: |
| $\Var(E[Y \mid X_1])$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Var(E[Y \mid X_2])$ | $\frac{(1 + \rho)^2}{2}$ | $\left(\frac{1 + \rho}{2}\right)^2
\hat \sigma_{X_2}^2$ |
| $\Var(E[Y \mid X_1, X_2])$ | $1 + \rho^2$ | $\sigma_{X_1}^2 + \rho^2$ |
| $\Var(Y)$ | $2$ | $\hat \sigma_y^2$ |
| $\Cov(E[Y \mid X_1], E[Y \mid X_2])$ | $\frac{1 + \rho}{2}$ | $\left(\frac{1 +
\rho}{2}\right)\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_1], E[Y \mid X_1, X_2])$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_2], E[Y \mid X_1, X_2])$ | $\frac{(1 + \rho)^2}{2}$ | $\frac{(1 + \rho)^2}{2}$ |
| $\Cov(E[Y \mid X_1], Y)$ | $1$ | $\hat \sigma_{X_1}^2$ |
| $\Cov(E[Y \mid X_2], Y)$ | $\frac{(1 + \rho)^2}{2}$ | $\frac{(1 + \rho)^2}{2}$ |
| $\Cov(E[Y \mid X_1, X_2], Y)$ | $1 + \rho^2$ | $\hat \sigma_{X_1}^2 + \rho^2$ |
: {#tbl-covform}

The estimated variances $\hat \sigma_{X_1}^2$, $\hat \sigma_{X_2}^2$, and $\hat
\sigma_Y^2$ are estimated from all of the observations in which that value
occurs.

In @tbl-covform, I don't want to have to estimate $\rho$ because I am not sure
what the REML estimator of $\rho$ is at the moment. If the REML estimated form
for the variances works, then I will consider how to estimate the covariance.

```{r}
#| label: cor mean, cor var

cormeancorvar <- function(df, rho) {

  # Steps:
  # 1. Separate into segments
  # 2. Construct estimates for each segment
  # 3. Combine estimates

  # 1. Separate into segments
  df_11 <- filter(df, delta_11 == 1)
  df_10 <- filter(df, delta_10 == 1)
  df_01 <- filter(df, delta_01 == 1)

  # 2. Construct estimates for each segment
  # 3. Combine estimates

}

```

```{r}
#| label: est mean, cor var

estmeancorvar <- function(df, rho) {

}

```

```{r}
#| label: est mean, est var

estmeanestvar <- function(df, rho) {

}

```

```{r}
#| label: MCMC

df <- gen_simple_data(obs_seg, true_mu, rho)

```