### 2024-04-11 src/explore/20240405-glsdcsim.qmd

---
title: "GLS-Debiased Calibration with Estimated Population Totals Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates the GLS-DC setup. We check for the following:

1. Bias of $\hat Y_1$ and $\hat Y_2$, and 
2. Bias of $\hat V(\hat Y_1)$ and $\hat V(\hat Y_2)$.

For each estimator $\hat Y_1$ and $\hat Y_1$, we construct

1. The two-phase regression estimator,
2. The Deville and Sarndal (1992) estimator,
3. The debiased calibration estimator, and
4. The debiased calibration estimator with estimated population totals.

# Simulation Setup

```{r}
#| label: setup

N_obs <- 10000
n1_obs <- 500
n2_obs <- 500

library(dplyr)
library(GECal)
library(doParallel)

```

We have the following superpopulation model:

$$
\begin{aligned}
X_{1i} &= 1 + e_{1i} \\
X_{2i} &= 2 + e_{2i} \\
X_{3i} &= 3 + e_{3i} \\
Y_{1i} &= 2 + 3x_{1i} + x_{2i} + e_{y_1, i}} \\
Y_{2i} &= 1 + 2x_{1i} + 2 x_{2i} + x_{3i} + e_{y_2, i}\\
\end{aligned}
$$

where $e_{m, i} \stackrel{ind}{\sim} N(0, 1)$. Then we generate a finite
population of size $N = $`{r} N_obs`. Sample $A_1$ has size $n_1 =$`{r} n1_obs`
and sample $A_2$ has size $n_2 =$`{r} n2_obs`. Each sample is an independent
Poisson from the finite population.

```{r}
#| label: data generating process

gen_data <- function(obs) {
  tibble(X1 = 1 + rnorm(obs),
         X2 = 2 + rnorm(obs),
         X3 = 3 + rnorm(obs)) %>%
    mutate(Y1 = 2 + 3 * X1 + X2 + rnorm(obs)) %>%
    mutate(Y2 = 1 + 2 * X1 + 2 * X2 + X3 + rnorm(obs))

}

gen_samples <- function(pop_df, n1, n2) {

  a1_ind <- sample(1:nrow(pop_df), n1)
  a2_ind <- sample(1:nrow(pop_df), n2)

  a1_df <- filter(pop_df, row_number() %in% a1_ind)
  a2_df <- filter(pop_df, row_number() %in% a2_ind)

  a1_df <- select(a1_df, X1, X2, Y1)
  a2_df <- select(a2_df, X1, X3, Y2)

  return(list(A1 = a1_df, A2 = a2_df))

}

```

# Estimation

```{r}
#| label: MCMC

set.seed(1)
pop_df <- gen_data(N_obs)
samps <- gen_samples(pop_df, n1_obs, n2_obs)
s1_df <- samps[[1]]
s2_df <- samps[[2]]

```