### 2024-04-16 src/explore/20240408-redokwongsim.qmd

---
title: "Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates part of the debiased calibration setup. I need to do this
to ensure that I understand how to use their package. I will compare their
method with the @deville1992calibration as done in the simulation section of 
@kwon2024debiased.

# Simulation Setup

```{r}
#| label: setup

N_obs <- 10000
sample_draws <- 1000

library(dplyr)
library(sampling)
library(GECal)
library(doParallel)
library(doRNG)

```

```{r}
#| label: data generation

gen_pop <- function(obs, method = 1) {
  x1 <- rnorm(obs, 2, 1)
  x2 <- runif(obs, 0, 4)
  mu_x <- x1 + x2
  z <- rnorm(obs, 0, 1)
  if (method == 1) {
    y <- mu_x + z + rnorm(obs, 0, 1)
  } else if (method == 2) {
    y <- mu_x + z^2 + rnorm(obs, 0, 1)
  } else {
    stop("method must be 1 or 2")
  }

  pi_vec <- ifelse(pt(-z - 2, 3) > 0.7, 0.7, pt(-z - 2, 3))

  return(tibble(X1 = x1, X2 = x2, Z = z, Y = y, Prob_i = pi_vec))
}

gen_samp <- function(pop_df) {
  # delta <- sampling::UPpoisson(pop_df$Prob_i)
  delta <- rbinom(nrow(pop_df), 1, pop_df$Prob_i)
  return(filter(pop_df, row_number() %in% which(delta == 1)))
}

```

```{r}
#| label: temp

#f = function(lambda, d, Xs, total, entropy, del, ..., returnw = F){
#  # w = d * ginv(drop(Xs %*% lambda), entropy = entropy)
#  if(entropy == "HD" & any(Xs %*% lambda >= 0)) return(rep(Inf, length(lambda)))
#  if(entropy == "PH" & any(abs(Xs %*% lambda) >= del)) return(rep(Inf, length(lambda)))
#  
#  if(entropy == "SL"){
#    w = d * drop(Xs %*% lambda)
#  }else if(entropy == "EL"){
#    w = -d / drop(Xs %*% lambda)
#  }else if(entropy == "ET"){
#    w = d * exp(drop(Xs %*% lambda))
#  }else if(entropy == "CE"){
#    w = d / (1 - exp(drop(Xs %*% lambda)))
#  }else if(entropy == "HD"){
#    w = d / drop(Xs %*% lambda)^2
#  }else if(entropy == "PH"){
#    w = d / sqrt(1 / drop(Xs %*% lambda)^2 - 1 / del^2)
#  }
#  if(entropy != "SL" & any(w <= 0)) return(rep(Inf, length(lambda)))
#  if(entropy == "CE" & any(w <= 1)) return(rep(Inf, length(lambda)))
#  if(returnw == T){
#    return(w)
#  }else{
#    return(colSums(Xs * w) - total)
#  }
#}
#
#h = function(lambda, d, Xs, total, entropy, del){
#  # return(t(Xs) %*% (Xs * d * ginvprime(drop(Xs %*% lambda), entropy = entropy)))
#  if(entropy == "SL"){
#    return(t(Xs) %*% (Xs * d))
#  }else if(entropy == "EL"){
#    w = -d / drop(Xs %*% lambda)
#    return(t(Xs) %*% (Xs * (w^2 / d)))
#  }else if(entropy == "ET"){
#    w = d * exp(drop(Xs %*% lambda))
#    return(t(Xs) %*% (Xs * w))
#  }else if(entropy == "CE"){
#    p_Stmp = 1 / (1 - exp(drop(Xs %*% lambda)))
#    return(t(Xs) %*% (Xs * d * p_Stmp * (p_Stmp - 1)))
#  }else if(entropy == "HD"){
#    return(t(Xs) %*% (Xs * (-2 * d / drop(Xs %*% lambda)^3)))
#  }else if(entropy == "PH"){
#    return(t(Xs) %*% (Xs * (d * (1 - (drop(Xs %*% lambda) / del)^2)^(-1.5))))
#  }
#}
##' @export
#GEcalib = function(Xs, d, total, entropy = c("SL", "EL", "ET", "CE", "HD", "PH"), ...,
#       DS = T, method = "Newton", control = list(maxit = 1e5, allowSingular = T)){
#  
#  del = quantile(d, 0.80) 
#  
#  init = rep(0, length(total))
#  if(!DS){
#    init[length(init)] = 1
#    # init = c(0, 0, 1)
#  }else{
#    if(entropy == "EL"){
#      init[1] = -1
#    }else if(entropy == "ET"){
#      # init[1] = 0
#    }else if(entropy == "SL"){
#      init[1] = 1
#    }else if(entropy == "CE"){
#      init[1] = -1
#    }else if(entropy == "HD"){
#      init[1] = -1
#    }else if(entropy == "PH"){
#      init[1] = 1 / sqrt(1 + 1 / del^2)
#    }
#  }
#  
#  nleqslv_res = nleqslv(init, f, jac = h, d = d, Xs = Xs, 
#                        total = total, entropy = entropy, del = del,
#                        method = method, control = control)
#  if(nleqslv_res$termcd != 1){
#    if(max(abs(f(nleqslv_res$x, d = d, Xs = Xs, total = total,
#                 entropy = entropy, del = del))) > 1e-5)
#      print(nleqslv_res)
#      w = NA
#  }else{
#    w = f(nleqslv_res$x, d = d, Xs = Xs, total = total,
#          entropy = entropy, del = del, returnw = T)             
#  }
#  return(w)
#}

```

```{r}
#| label: estimation

set.seed(1)
pop_df <- gen_pop(N_obs)
pop_Xtot <- c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2))
pop_d <- 1 / pop_df$Prob_i

clust <- makeCluster(min(detectCores() - 2, 100))
registerDoParallel(clust)

mc_res <- foreach(ind = 1:sample_draws, .packages = c("dplyr", "GECal")) %dorng% {

# Sample
set.seed(2)
samp_df <- gen_samp(pop_df)
samp_X <- samp_df %>% select(X1, X2) %>% as.matrix()
samp_X <- cbind(rep(1, nrow(samp_X)), samp_X)
d_weights <- 1 / samp_df$Prob_i

## Method: EL
method = "EL"
if (method == "EL") {
  dtot <- sum(-1 / pop_d)
  dsamp <- -1 / d_weights
  samp_df <- samp_df %>%
    mutate(gdi = dsamp) %>%
    mutate(gpinv = 1 / Prob_i^2)
  ds_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "EL", DS = TRUE)
  dc_w <- GEcalib(cbind(samp_X, dsamp), d = rep(1, nrow(samp_df)), c(pop_Xtot, dtot),
   entropy = "EL", DS = FALSE)
} else if (method == "ET") {
  dtot <- sum(log(d_weights))
  dsamp <- log(d_weights)
  samp_df <- samp_df %>%
    mutate(gdi = dsamp) %>%
    mutate(gpinv = 1 / Prob_i)
  ds_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "ET", DS = TRUE)
  dc_w <- GEcalib(cbind(samp_X, dsamp), d = d_weights, c(pop_Xtot, dtot),
    entropy = "ET", DS = FALSE)
}

# Mean Estimation
y_ds <- sum(samp_df$Y * ds_w) / N_obs
y_dc <- sum(samp_df$Y * dc_w) / N_obs
y_hajek <- sum(samp_df$Y / samp_df$Prob_i) / sum(1 / samp_df$Prob_i)

## Variance Estimation
#ds_mod <- lm(Y ~ X1 + X2, weights = gpinv, data = samp_df)
#dc_mod <- lm(Y ~ X1 + X2 + gdi, weights = gpinv, data = samp_df)
#
#v_ds <- sum((1 - samp_df$Prob_i) * (ds_w / N_obs)^2 *
#    (samp_df$Y - predict(ds_mod, samp_df))^2)
#v_dc <- sum((1 - samp_df$Prob_i) * (dc_w / N_obs)^2 *
#    (samp_df$Y - predict(dc_mod, samp_df))^2)
v_hajek <- 1 / N_obs^2 *
sum((samp_df$Y - y_hajek)^2 * (1 / samp_df$Prob_i^2 - 1 / samp_df$Prob_i))

Pi_ds <- diag(1 - samp_df$Prob_i) * (ds_w / N_obs)^2
z_ds <- model.matrix(~X1 + X2 + gdi, data = samp_df)
gam_ds <- solve(t(z_ds) %*% diag(1 / samp_df$Prob_i) %*% z_ds,
                t(z_ds) %*% diag(1 / samp_df$Prob_i) %*% samp_df$Y)
v_ds <- (samp_df$Y - t(gam_ds) %*% t(z_ds)) %*% 
             Pi_ds %*% t(samp_df$Y - t(gam_ds) %*% t(z_ds))

Pi_dc <- diag(1 - samp_df$Prob_i) * (dc_w / N_obs)^2
z_dc <- model.matrix(~X1 + X2 + gdi, data = samp_df)
gam_dc <- solve(t(z_dc) %*% diag(samp_df$gpinv) %*% z_dc,
                t(z_dc) %*% diag(samp_df$gpinv) %*% samp_df$Y)
v_dc <- 
  (samp_df$Y - t(gam_dc) %*% t(z_dc)) %*% Pi_dc %*% t(samp_df$Y - t(gam_dc) %*% t(z_dc))

return(tibble(y_ds = y_ds, y_dc = y_dc, y_h = y_hajek,
              v_ds = as.numeric(v_ds), v_dc = as.numeric(v_dc), v_h = v_hajek))

} %>% bind_rows()

stopCluster(clust)

## Assessment
bias_ds <- mean(mc_res$y_ds - mean(pop_df$Y))
bias_dc <- mean(mc_res$y_dc - mean(pop_df$Y))
bias_h <- mean(mc_res$y_h - mean(pop_df$Y))

rmse_ds <- sqrt(mean((mc_res$y_ds - mean(pop_df$Y))^2))
rmse_dc <- sqrt(mean((mc_res$y_dc - mean(pop_df$Y))^2))
rmse_h <- sqrt(mean((mc_res$y_h - mean(pop_df$Y))^2))

cr_ds <- mean(abs(mc_res$y_ds - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_ds))
cr_dc <- mean(abs(mc_res$y_dc - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_dc))
cr_h <- mean(abs(mc_res$y_h - mean(pop_df$Y)) < qnorm(0.975) * sqrt(mc_res$v_h))

```