### 2024-05-08 final/reports/twophaseridge_ov.tex

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, amsthm, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
%\usepackage[style=authoryear]{biblatex}
\usepackage{booktabs}
\usepackage{float}

\usepackage{amsmath, amsthm, mathtools,commath}
\usepackage{graphics, color}
\usepackage{latexsym}
\usepackage{amssymb, amsfonts, bm}
\usepackage{mathrsfs}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{makeidx}
\usepackage{fullpage}
\usepackage{booktabs, arydshln}
\usepackage{comment} 
%\makeindex

\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
in-ven-tory e-num-er-ate char-ac-ter-is-tic}

\usepackage[round]{natbib}
%\bibliographystyle{apalike2}
% \bibliographystyle{jmr}

\newcommand{\biblist}{\begin{list}{}
{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
\labelwidth 0 cm \labelsep 0.50 cm
\usecounter{list}}\clubpenalty4000\widowpenalty4000}
\newcommand{\ebiblist}{\end{list}}

\newcounter{list}

%\usepackage{setspace}

%\usepackage{hangpar}
\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
% remove % from next line for final copy
\renewcommand{\lbl}[1]{\label{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}

\newtheorem{lemma}{Lemma}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\bf}[1]{\mathbf{#1}}

\begin{document}

\title{Similarities between Nonnested Regression Estimation and Ridge Regression}
\author{Caleb Leedy}
\maketitle

\baselineskip .3in

\section{Nonnested Regression}

In the case of non-nested two phase sampling, we have $A_1 = (X_i)_{i = 1}^{n_1}$
and $A_2 = (X_1, Y_i)_{i = 1}^{n_2}$ with $A_1$ and $A_2$ being selected
independently from the same sampling frame. The regression estimator is then

\begin{align*}
\hat{\bar{Y}}_{reg} 
  &= \hat{\bar{Y}}_{HT} + (\hat{\bar{X}}_c - \hat{\bar{X}}_2)' \hat \beta_2 
  & \text{for } \hat \beta_2 = \left(\sum_{i \in A_2} x_i x_i'\right)^{-1} 
    \sum_{i \in A2} x_i y_i \text{ with } x_{1i} = 1 \\
  &= \hat{\bar{Y}}_{HT} + (\hat{\bar{X}}_1 - \hat{\bar{X}}_2)'W'\hat \beta_2 
  & \text{ if } \hat{\bar{X}}_c = W \hat{\bar{X}}_1 + (I - W) \hat{\bar{X}}_2\\
  &= \sum_{i \in A_1} x_i' W' \hat \beta_2 + \sum_{i \in A_2} (y_i - x_i' W'
  \hat \beta_2).
\end{align*}

While the matrix $W$ controls the interaction between $A_1$ and $A_2$ it also
plays the role of shrinkage on $\hat \beta_2$.

\textcolor{red}{If the population total of $X$ is known, then the regression estimator $$ \hat{\bar{Y}}_{reg} 
  = \hat{\bar{Y}}_{HT} + ({\bar{X}} - \hat{\bar{X}}_2)' \hat \beta_2$$
  can be used. In this case, $\hat{\beta}_2$ does not have the shrinkage interpretation. Now, if $\bar{X}$ is unknown, we need to use an estimator. If we use the best estimator $\hat{\bar{X}}_c$, then the residual is orthogonal. Otherwise, if we use an inefficient estimator $\hat{\bar{X}}_1$, we pay its price. It is more related to attenuation effect in the measurement error model problem. 
  }
\section{Ridge Regression}

Given a sample $A$, ridge regression solves the optimization problem,

$$\hat \beta = \argmin_{\beta \in \R^p} \sum_{i \in A} (y_i - x_i'\beta)^2 + \beta^T
(\lambda I_p) \beta.$$

Differentiating with respect to $\beta$ and setting this equal to zero yields a
solution,

$$\hat \beta = \left(\sum_{i \in A} x_i x_i' + \lambda I_p\right)^{-1} \sum_{i
\in A} x_i' y_i = (X'X + \lambda I_p)^{-1} X'Y.$$

Let $\hat \beta_{OLS} = \hat \beta_2 = (X'X)^{-1} X'Y$, then using the
Sherman-Morrison-Woodbury inverse formula,

$$\hat \beta = (I_p - (X'X)^{-1}(\lambda^{-1}I_p + (X'X)^{-1}))^{-1} \hat \beta_{OLS}.$$

\section{Discussion}

The previous two sections suggest that if we let

$$
\begin{aligned}
  W &= (I_p - (X'X)^{-1}(\lambda^{-1}I_p + (X'X)^{-1}))^{-1} \\
  \iff \lambda W &= (X'X)(I_p - W) \\ 
  \iff \lambda &= (X'X)(W^{-1} - I_p)
\end{aligned}
$$

then we would have equivalent results. In this way, non-nested two phase
sampling is like ridge regression.

\begin{itemize}
  \item We could connect this to LASSO?
  \item This could be a connection to Bayesian statistics and choosing priors on
    our estimates?
  \item Maybe this can help us choose priors for ridge regression better?
\end{itemize}

\section{Alternative idea for non-parametric regression estimation}

One possible idea is to develop a debiased estimation under two-phase sampling. If the covariates are high dimensional or the regression employs nonparametric regression (such as spline or random forest, etc), then the resulting two-phase regression estimator can be biased. To correct for the bias, we can use the following approach.

\begin{enumerate}
\item Split the sample $A_2$ into two parts: $A_2= A_2^{(a)} \cup A_2^{(b)}$.
We can use SRS to split the sample, but other sampling designs can be used. 
\item Use the observations in $A_2^{(a)}$ only to obtain a predictor of $y_i$, $\hat{f}^{(a)} (\bx_i)$. Also, use the observations in $A_2^{(b)}$ only to obtain a predictor of $y_i$, $\hat{f}^{(b)} (\bx_i)$. 
\item Let 
$$ \hat{f}(\bx_i) = \left( \hat{f}^{(a)} (\bx_i) + \hat{f}^{(b)} (\bx_i) \right)/ 2$$
be the predictor combining two samples. 
\item The final debiased two-phase regression estimator is given by 
\begin{equation}
\hat{Y}_{\rm d, reg} = \sum_{i \in A_1} w_{1i}  \hat{f}(\bx_i) +  \sum_{i \in A_2^{(a)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{f}^{(b)} (\bx_i) \right\} +\sum_{i \in A_2^{(b)} } w_{1i} \pi_{2i \mid 1}^{-1}   \left\{y_i - \hat{f}^{(a)} (\bx_i) \right\}
\label{debiased}
\end{equation}
\end{enumerate}
 If I am not mistaken, there are several advantages of the debiased two-phase regression estimator in (\ref{debiased}). 
\begin{enumerate}
\item Unlike the classical two-phase regression estimator using nonparametric regression, we can establish asymptotic unbiasedness and  $\sqrt{n}$-consistency. 
\item Even if we use the sample split, there is no efficiency loss. That is, the asymptotic variance is equal to 
$$ V \left( \hat{Y}_{\rm d, reg} \right) = V \left( \hat{Y}_1 \right) + E\left[ V \left\{ \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} \left( y_i - f(\bx_i) \right) \mid A_1 \right\} \right] 
$$
where $f(\bx_i)$ is the probability limit of $\hat{f}(\bx_i)$. 
\item Variance estimation is also straightforward. We can compute 
$$ \hat{\eta}_i = \hat{f}(\bx_i) +\delta_i  \pi_{2i \mid 1}^{-1} I_i^{(a)} \left\{ y_i - \hat{f}^{(b)}(\bx_i) \right\} +\delta_i \pi_{2i \mid 1}^{-1} I_i^{(b)} \left\{ y_i - \hat{f}^{(a)} (\bx_i) \right\}$$
and apply to the variance estimation formula for the first-phase sample, 
where $I_i^{(a)}$ is the indicator function for $A_2^{(a)}$ such that $I_i^{(a)} = 1$ if $i \in A_2^{(a)}$ and $I_i^{(a)}=0$ otherwise. Also, $I_i^{(b)}= 1- I_i^{(a)}$. 
\end{enumerate}

I also have an idea on how to implement the above debiased regression estimator using calibration. I will give more details once we are confident in the proposed idea. 
\end{document}