### 2024-05-08 src/explore/20240506-msdcsim.qmd

---
title: "Multi-Source Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates multi-source sampling with the debiased
calibration of @kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

For the estimator $\hat Y$ we construct

1. The $\pi^*$-estimator,
2. The two-phase regression estimator,
3. The debiased calibration estimator (with known finite population total), and
4. The debiased calibration estimator with estimated population totals from the
   Phase 1 sample.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 1000
n2_obs <- 2000

library(dplyr)
library(GECal)
library(nleqslv)
library(doParallel)
library(doRNG)
library(CVXR)
library(rlang)

```

We have the following superpopulation model with $N=$`{r} N_obs` elements:

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{0i} &= \Phi(-x_{1i} - 2)\\
\pi_{1i} &= |x_{1i}| / \sum_{i \in U} |x_{i1}|\\
\pi_{2i} &= n_2 / N.\\
\end{aligned}
$$

We observe the following columns in each sample

| Sample | $X_1$ | $X_2$ | $X_3$ |  $Y$ |
|:-------|------:|------:|------:|-----:|
| $A_0$  | ✓     | ✓     | ✓     |  ✓   |  
| $A_1$  | ✓     | ✓     |       |      |  
| $A_2$  | ✓     |       | ✓     |      |

For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
sample $A_2$ is a simple random sample.

```{r}
#| label: generate population and samples

gen_pop <- function(obs) {

  x1 <- rnorm(obs, 2, 1)
  x2 <- runif(obs, 0, 4)
  x3 <- rnorm(obs, 0, 1)
  x4 <- runif(obs, 0.1, 0.9)
  eps <- rnorm(obs)

  y <- 3 * x1 + 2 * x2 + eps

  return(tibble(X1 = x1, X2 = x2, X3 = x3, X4 = x4, Y = y))
}

#' This function adds sampling probabilites to a population data frame.
#'
#' @details The phase 2 formula will be included for every element. There is no
#' information about if an element is part of the Phase 1 or Phase 2 sample.
update_pop <- function(pop_df, p0_formula, p1_formula, p2_formula) {

  pi0 <- eval(parse_expr(p0_formula))
  pi1 <- eval(parse_expr(p1_formula))
  pi2 <- eval(parse_expr(p2_formula))

  pop_df %>%
    mutate(pi0 = pi0, pi1 = pi1, pi2 = pi2)
}

#' This function generates Phase 1 and Phase 2 samples from an updated
#' population with selection probabilites.
gen_samps <- function(upop_df, p0_type, p1_type, p2_type) {

  if (p0_type == "poisson") {
    del0 <- rbinom(nrow(upop_df), 1, upop_df$pi0)
  } else if (p1_type == "srs") {
    ind <- sample(1:nrow(upop_df), size = upop_df$pi0[1] * nrow(upop_df))
    del0 <- as.numeric(1:nrow(upop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 0.")
  }

  p0_df <- mutate(upop_df, del0 = del0) %>% filter(del0 == 1)

  if (p1_type == "poisson") {
    del1 <- rbinom(nrow(upop_df), 1, upop_df$pi1)
  } else if (p1_type == "srs") {
    ind <- sample(1:nrow(upop_df), size = upop_df$pi1[1] * nrow(upop_df))
    del1 <- as.numeric(1:nrow(upop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 1.")
  }

  p1_df <- mutate(upop_df, del1 = del1) %>% filter(del1 == 1)

  if (p2_type == "poisson") {
    del2 <- rbinom(nrow(upop_df), 1, upop_df$pi2)
  } else if (p2_type == "srs") {
    ind <- sample(1:nrow(upop_df), size = round(upop_df$pi2[1] * nrow(upop_df)))
    del2 <- as.numeric(1:nrow(upop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 2.")
  }

  p2_df <- mutate(upop_df, del2 = del2) %>% filter(del2 == 1)

  return(list(p0_df, p1_df, p2_df))
}

```

```{r}
#| label: estimators

```