### 2024-05-13 src/explore/20240506-msdcsim.qmd

---
title: "Multi-Source Debiased Calibration Simulation"
author: "Caleb Leedy"
from: markdown+emoji
format: 
  html:
    embed-resources: true
    code-fold: true
    grid:
      margin-width: 450px
bibliography: references.bib
reference-location: margin
comments:
  hypothesis: true
---

# Summary

This report simulates multi-source sampling with the debiased
calibration of @kwon2024debiased. We check for the following:

1. Bias of $\hat Y$ and,
2. Bias of $\hat V(\hat Y)$.

For the estimator $\hat Y$ we construct

1. The a HT-estimator,
2. The non-nested two-phase regression estimator ($A_0$ and $A_2$),
3. The debiased calibration estimator (with known finite population total), and
4. The debiased calibration estimator with estimated population totals from the
   Phase 1 sample.

# Simulation Setup

```{r}
#| label: setup

B_sims <- 1000

N_obs <- 10000
n1_obs <- 2000

library(dplyr)
library(nleqslv)
library(doParallel)
library(doRNG)

```

We have the following superpopulation model with $N=$`{r} N_obs` elements:

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
Z_i &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{0i} &= \min(\max(\Phi(-x_{1i}), 0.02), 0.9)\\
\pi_{1i} &= n_1 / N\\
\pi_{2i} &= \Phi(x_{2i} - 2) \\
\end{aligned}
$$

We observe the following columns in each sample

| Sample | $X_1$ | $X_2$ | $X_3$ |  $Y$ |
|:-------|------:|------:|------:|-----:|
| $A_0$  | ✓     | ✓     | ✓     |  ✓   |  
| $A_1$  | ✓     |       | ✓     |      |  
| $A_2$  | ✓     | ✓     |       |      |

For the sampling mechanism both $A_0$ and $A_1$ are selected using a Poisson
sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
sample $A_2$ is a simple random sample.

```{r}
#| label: generate population and samples

gen_pop <- function(N_obs, n1_obs) {

  x1 <- rnorm(N_obs, 2, 1)
  x2 <- runif(N_obs, 0, 4)
  x3 <- rnorm(N_obs, 5, 1)
  z <- rnorm(N_obs, 0, 1)

  eps <- rnorm(N_obs)

  y <- 3 * x1 + 2 * x2 + eps

  pi0 <- pmin(pmax(pnorm(-x1), 0.02), 0.9)
  # pi1 <- pnorm(-z - 1) # Informative sampling. Do we want this?
  pi1 <- n1_obs / N_obs
  # pi2 <- pmin(pmax(pnorm(x2 - 4), 0.01), 0.9)
  pi2 <- pnorm(x2 - 2)

  return(tibble(X1 = x1, X2 = x2, X3 = x3, Y = y,
                pi0 = pi0, pi1 = pi1, pi2 = pi2))
}

#' This function generates Phase 1 and Phase 2 samples from an updated
#' population with selection probabilites.
gen_samps <- function(pop_df, p0_type, p1_type, p2_type) {

  if (p0_type == "poisson") {
    del0 <- rbinom(nrow(pop_df), 1, pop_df$pi0)
  } else if (p0_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$pi0[1] * nrow(pop_df))
    del0 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 0.")
  }

  p0_df <- mutate(pop_df, del0 = del0) %>% filter(del0 == 1)

  if (p1_type == "poisson") {
    del1 <- rbinom(nrow(pop_df), 1, pop_df$pi1)
  } else if (p1_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = pop_df$pi1[1] * nrow(pop_df))
    del1 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 1.")
  }

  p1_df <- mutate(pop_df, del1 = del1) %>% filter(del1 == 1)

  if (p2_type == "poisson") {
    del2 <- rbinom(nrow(pop_df), 1, pop_df$pi2)
  } else if (p2_type == "srs") {
    ind <- sample(1:nrow(pop_df), size = round(pop_df$pi2[1] * nrow(pop_df)))
    del2 <- as.numeric(1:nrow(pop_df) %in% ind)
  } else {
    stop("We have only implemented poisson and srs for Sample 2.")
  }

  p2_df <- mutate(pop_df, del2 = del2) %>% filter(del2 == 1)

  return(list(p0_df, p1_df, p2_df))
}

```

```{r}
#| label: estimators

ht_est <- function(p0_df, N_obs, p0_type = "poisson") {

  theta <- sum(p0_df$Y / p0_df$pi0) / N_obs

  if (p0_type == "poisson") {

    v_hat <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * p0_df$Y^2) / N_obs^2

  } else {

    v_hat <- NA

  }

  return(list(theta = theta, v_hat = v_hat))
}

reg_est <- function(p0_df, p1_df, N_obs, p0_samp = "poisson", p1_samp = "srs") {

  v_poisson <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * p0_df$Y^2) / N_obs^2
  v_srs <- (1 / nrow(p0_df) - 1 / N_obs) * var(p0_df$Y)
  d_eff <- v_poisson / v_srs
  n0_eff <- nrow(p0_df) / d_eff
  n1_eff <- nrow(p1_df)

  x1_hat <- 
    c(sum(p1_df$X1 / p1_df$pi1) / N_obs,
      sum(p1_df$X2 / p1_df$pi1) / N_obs)

  x0_hat <- 
    c(sum(p0_df$X1 / p0_df$pi0) / N_obs,
      sum(p0_df$X2 / p0_df$pi0) / N_obs)

  w <- n1_eff / (n1_eff + n0_eff)
  xc_hat <- (n1_eff * x1_hat + n0_eff * x0_hat) / (n1_eff + n0_eff)

  # Estimate beta
  mod <- lm(Y ~ X1 + X2, data = p0_df)

  # Regression Estimation
  y_ht <- sum(p0_df$Y / p0_df$pi0) / N_obs
  theta <- y_ht + sum((xc_hat - x0_hat) * mod$coefficients[2:3])

  # Variance Estimation
  coefs <- mod$coefficients
  eps <- (p0_df$Y - coefs[1] * w - coefs[2] * w * p0_df$X1 - coefs[3] * w * p0_df$X2)
  var_x <- 
  (1 / nrow(p1_df) - 1 / N_obs) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))

  v1 <- w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
  v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / N_obs^2

  cov_xy <- c(sum((1 - p0_df$pi0) / p0_df$pi0^2 * coefs[2] * 
                 (n0_eff) / (n1_eff + n0_eff) * p0_df$X1 * eps) / N_obs^2,
              sum((1 - p0_df$pi0) / p0_df$pi0^2 * coefs[3] * 
                 (n0_eff) / (n1_eff + n0_eff) * p0_df$X3 * eps) / N_obs^2)
  v3 <- sum(cov_xy)

  v_hat <- v1 + v2 + v3

  return(list(theta = theta, v_hat = as.numeric(v_hat)))
}

#' This  function does the debiased calibration
#'
#' @param zmat - A n x p matrix
#' @param w1i - A vector of length n
#' @param d2i - A vector of length n
#' @param T1 - A vector of length p
#' @param qi - A scaler or a vector of length n
#' @param entropy - One of "EL", "ET"
solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {

  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

    return(res)
  }

  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)

  if (!(res$termcd %in% c(1))) {
    return(NA)
  }

  return(resw)

}

dc_ms_ybar <- 
  function(p0_df, p1_df, p2_df, pop_df, estT1, qi = 1, entropy = "EL", reg = FALSE) {

  d_vec <- 1 / (p0_df$pi0)
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gdip <- -1 / (1 / pop_df$pi0)
    gpinv <- d_vec^2
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdip <- log(1 / pop_df$pi0)
    gpinv <- d_vec
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- cbind(rep(1, nrow(p0_df)), as.matrix(select(p0_df, X1, X2, X3)))
  zmat <- cbind(samp_X, gdi)
  w1i <- 1
  d0i <- 1 / p0_df$pi0

  if (estT1) {
    x1_0 <- sum(p0_df$X1 / p0_df$pi0) / N_obs
    x2_0 <- sum(p0_df$X2 / p0_df$pi0) / N_obs
    x3_0 <- sum(p0_df$X3 / p0_df$pi0) / N_obs

    x1_1 <- sum(p1_df$X1 / p1_df$pi1) / N_obs
    x3_1 <- sum(p1_df$X3 / p1_df$pi1) / N_obs

    x1_2 <- sum(p2_df$X1 / p2_df$pi2) / N_obs
    x2_2 <- sum(p2_df$X2 / p2_df$pi2) / N_obs

    xhat <- c(x1_0, x2_0, x3_0, x1_1, x3_1, x1_2, x2_2)
    x_mat <- 
    matrix(c(1, 0, 0,
             0, 1, 0,
             0, 0, 1,
             1, 0, 0,
             0, 0, 1,
             1, 0, 0,
             0, 1, 0), ncol = 3, byrow = TRUE)
    vmat <- matrix(0, nrow = 7, ncol = 7)
    vmat[1, 1] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1^2) / nrow(pop_df)^2
    vmat[1, 2] <- 
      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1 * p0_df$X2) / nrow(pop_df)^2
    vmat[1, 3] <-
      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X1 * p0_df$X3) / nrow(pop_df)^2
    vmat[2, 1] <- vmat[1, 2]
    vmat[3, 1] <- vmat[1, 3]
    vmat[2, 2] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X2^2) / nrow(pop_df)^2
    vmat[2, 3] <-
      sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X2 * p0_df$X3) / nrow(pop_df)^2
    vmat[3, 2] <- vmat[2, 3]
    vmat[3, 3] <- sum((1 / p0_df$pi0^2 - 1 / p0_df$pi0) * p0_df$X3^2) / nrow(pop_df)^2

    vmat[4, 4] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * var(p1_df$X1)
    vmat[5, 5] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * var(p1_df$X3)
    vmat[4, 5] <- (1 / nrow(p1_df) - 1 / nrow(pop_df)) * cov(p1_df$X1, p1_df$X3)
    vmat[5, 4] <- vmat[4, 5]

    vmat[6, 6] <- sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X1^2) / nrow(pop_df)^2
    vmat[7, 7] <- sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X2^2) / nrow(pop_df)^2
    vmat[6, 7] <- 
      sum((1 / p2_df$pi2^2 - 1 / p2_df$pi2) * p2_df$X1 * p2_df$X2) / nrow(pop_df)^2
    vmat[7, 6] <- vmat[6, 7]

    xgls <- drop(MASS::ginv(t(x_mat) %*% solve(vmat) %*% x_mat) %*% 
                (t(x_mat) %*% solve(vmat) %*% xhat))

    T1 <- c(nrow(pop_df),
            xgls[1] * nrow(pop_df),
            xgls[2] * nrow(pop_df),
            xgls[3] * nrow(pop_df),
            sum(gdip * qi))
  } else {
    T1 <- c(nrow(pop_df),
            sum(pop_df$X1),
            sum(pop_df$X2),
            sum(pop_df$X3),
            sum(gdip * qi))
  }

  dc_w <- solveGE(zmat, w1i, d0i, T1, qi, entropy)

  # Mean Estimation
  theta <- sum(p0_df$Y * dc_w * w1i) / nrow(pop_df)
  z_dc <- model.matrix(~1 + p0_df$X1 + p0_df$X2 + p0_df$X3 + gdi)
  gam_dc <- 
    solve(t(z_dc) %*% diag(d0i * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d0i * gpinv) %*% p0_df$Y)
  coefs <- as.numeric(gam_dc)
  u_mat <- 
  matrix(c(rep(1, nrow(p0_df)), p0_df$X1, p0_df$X2, p0_df$X3, gdi * qi), ncol = 5)
  eps <- (p0_df$Y - drop(u_mat %*% coefs))
  if (estT1) {
    # Variance Estimation
    # This is the variance of bar x gls not x gls total.
    var_x <- MASS::ginv(t(x_mat) %*% MASS::ginv(vmat) %*% x_mat)

    v1 <- matrix(coefs[2:4], nrow = 1) %*% var_x %*% matrix(coefs[2:4], ncol = 1)
    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
    v_hat <- v1 + v2 
  } else {
    v2 <- sum((1 - p0_df$pi0) / p0_df$pi0^2 * eps^2) / nrow(pop_df)^2
    v_hat <- v2
  }
  if (reg) {
    theta <- 
    (drop(T1 %*% gam_dc) + 
      sum(1 / p0_df$pi0 * (p0_df$Y - drop(z_dc %*% gam_dc)))) / nrow(pop_df) 
  }

  return(list(theta = theta, v_hat = as.numeric(v_hat)))
}

```

```{r}
#| label: run MCMC

set.seed(1)
pop_df <- gen_pop(N_obs, n1_obs)
true_theta <- mean(pop_df$Y)

clust <- makeCluster(min(detectCores() - 2, 100), outfile = "")
registerDoParallel(clust)

mc_res <- 
  foreach(iter=1:B_sims,
          .packages = c("dplyr", "nleqslv"),
          .options.RNG = 1) %dorng% {

    if (iter %% 100 == 0) {
      print(paste0("Iter: ", iter))
    }

    samps <- gen_samps(pop_df, "poisson", "srs", "poisson")
    p0_df <- samps[[1]]
    p1_df <- samps[[2]]
    p2_df <- samps[[3]]

    ht <- ht_est(p0_df, N_obs)
    reg <- reg_est(p0_df, p1_df, N_obs)
    ms_pop <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = FALSE)
    ms_est <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE)
    ms_reg <- dc_ms_ybar(p0_df, p1_df, p2_df, pop_df, estT1 = TRUE, reg = TRUE)

    return(tibble(Est = c("HT", "NNReg", "MSPop", "MSEst", "MSReg"),
                  Theta = c(ht[[1]], reg[[1]], ms_pop[[1]], ms_est[[1]], ms_reg[[1]]),
                  Var = c(ht[[2]], reg[[2]], ms_pop[[2]], ms_est[[2]], ms_reg[[2]]),
                  Iter = iter)
    )

  } %>% bind_rows()

stopCluster(clust)

```

```{r}
#| label: analyze results

# Check for failures
filter(mc_res, is.na(Theta))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "MSReg" ~ 5,
    Est == "MSEst" ~ 4,
    Est == "MSPop" ~ 3,
    Est == "NNReg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest) %>%
  knitr::kable("latex", booktabs = TRUE, digits = 4) %>%
  cat(file = "tables/msdcsim1.tex")

```