### 2025-01-23 final/paper/paper.tex
Updated paper with Dr. Kim's suggestions.

\documentclass[12pt]{article}

\usepackage{amsmath, amssymb, amsthm, mathrsfs, fancyhdr}
\usepackage{syntonly, lastpage, hyperref, enumitem, graphicx}
%\usepackage[style=authoryear]{biblatex}
\usepackage{booktabs}
\usepackage{float}

\usepackage{amsmath, amsthm, mathtools}
\usepackage{graphics, color}
\usepackage{latexsym}
\usepackage{amssymb, amsfonts, bm}
\usepackage{mathrsfs}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{makeidx}
\usepackage{fullpage}
\usepackage{booktabs, arydshln}
\usepackage{comment} 
%\makeindex

\hbadness=10000 \tolerance=10000 \hyphenation{en-vi-ron-ment
in-ven-tory e-num-er-ate char-ac-ter-is-tic}

\usepackage[round]{natbib}
%\bibliographystyle{apalike2}
% \bibliographystyle{jmr}

\newcommand{\biblist}{\begin{list}{}
{\listparindent 0.0cm \leftmargin 0.50cm \itemindent -0.50 cm
\labelwidth 0 cm \labelsep 0.50 cm
\usecounter{list}}\clubpenalty4000\widowpenalty4000}
\newcommand{\ebiblist}{\end{list}}

\newcounter{list}

%\usepackage{setspace}

%\usepackage{hangpar}
\newcommand{\lbl}[1]{\label{#1}{\ensuremath{^{\fbox{\tiny\upshape#1}}}}}
% remove % from next line for final copy
\renewcommand{\lbl}[1]{\label{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{result}{Result}

\newtheorem{lemma}{Lemma}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\MAP}{{\text{MAP}}}
\newcommand{\Cov}{{\text{Cov}}}
\newcommand{\Var}{{\text{Var}}}
\newcommand{\logistic}{{\text{logistic}}}

\newcommand{\bx}{\mathbf{x}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\bf}[1]{\mathbf{#1}}

\begin{document}

\title{Debiased Calibration for Generalized Two-Phase Sampling}
\author{Caleb Leedy \and Jae-Kwang Kim}
\maketitle

\baselineskip .3in

\begin{abstract}
Generalized entropy calibration \citep{gneiting2007strictly} is a nonparametric
method to estimate survey weights without using the design weights in the
specified loss function. \cite{kwon2024debiased} explore generalized
entropy calibration to estimate survey weights with known population totals.
By using a two-phase sampling framework, we extend this method to allow for
estimated population totals. We then generalize the two-phase sampling framework to
two additional cases: first, when we have two non-nested samples, and
second, when we want to combine three or more samples. 
We apply our method to the Medicare Current Beneficiary Survey for which we
get more efficient estimates by combining this data set with population level
controls obtained from the Center of Medicare and Medicaid Services as well as
survey data from the U.S. Census' American Community Survey.
\end{abstract}

\newpage

\section{Introduction}

Government surveys in the United States provide the public with diverse
information about a range of population characteristics. These include health
indicators from the National Health and Nutrition Examination Survey, population
and housing data from the American Community Survey, and labor force statistics
from the Current Population Survey. All of these surveys are probability
surveys. Probability sampling is considered the
gold-standard for survey sampling. However, obtaining a large probability
sample is getting more difficult because people are more likely to choose not to
respond. One solution to this problem is data integration. Instead of analyzing
a probability sample by itself, we can incorporate information from other
probability samples in order to have an analysis with higher statistical power.

The goal of statistical data integration is to combine information from one
survey with the variable of interest to another survey with
shared covariates. 
This integration usually involves one of two contexts: incorporating item
level responses or using summary statistics. These mirror the micro and macro
approaches described in \cite{yang2020statistical} respectively. The main
technique to incorporate item level responses is mass imputation. With access to
individual item level responses, the traditional approach is to build a
predictive model of the variable of interest from the covariates shared between
the two surveys and predict the missing outcome variable in the survey without
the variable of interest. An early use of mass imputation is found in
\cite{breidt1996two} where they use mass imputation to improve estimation in
the National Resources Inventory survey. Later, \cite{kim2012combining} use a
parametric model to develop a novel resampling technique to construct confidence
intervals of population level characteristics. \cite{chen2022nonparametric} use
a nonparametric model to estimate population characteristics. In the macro
approach, one uses summary statistics from other studies to adjust the current
estimation. This is commonly done using generalized least squares (GLS).
\cite{hidiroglou2001double} apply the GLS technique to non-nested two-phase
samples. \cite{merkouris2004combining} use GLS with multiple surveys and showed
an optimality result. This was later extended to area level domains in
\cite{merkouris2010combining}. The main idea of the GLS approach is to construct
a regression estimator for two-phase samples. In Section 2 of the paper, we
explain the basic approach of regression estimation and show how it can be
generalized to a calibration estimator.

This paper, however, generalizes the two-phase regression estimator using the
debiased calibration technique of \cite{kwon2024debiased}.
We incorporate summary statistics of other surveys by
using a generalized two-phase sample similar to the GLS method. The first phase
sample consists of data
from multiple sources, and the second phase sample contains the existing
data. Unlike previous data integration methods, however, we also include a
pre-chosen function of the second phase design weights, which allows us to
develop theory showing that our
estimates are asymptotically design unbiased and that we have valid confidence
intervals. Moreover, we can show that this technique outperforms traditional
regression methods in finite samples.

\textcolor{red}{(I am not sure whether the current introduction provides a good
  guide for the main contents of the paper. The introduction does not match
  well with the main contents of the paper.)
}

\textcolor{red}{CL: I have modified the last two paragraphs to help tie into the
main work of the paper.}

This paper is organized as follows. 
Section 2 describes the existing methods to combine probability samples using
regression estimators or calibration.
Section 3 details the generalized
two-phase sampling framework and provides theory for our estimator to be design
unbiased and the variance estimation to be valid. In Section 4, we extend this
framework first to non-nested two-phase sampling and second to multiple
samples. Section 5 provides several simulation studies which show that our
method works, and Section 6 applies the method to the Medicare Current
Beneficiary Survey.

\newpage

\section{Basic Setup}

Consider a finite population, $U = \{1, \dots, N\}$, of size $N$ 
with measurement  $(\bx_i, y_i)$, where $\bx_i$ is the realization of the
auxiliary variables and $y_i$ is the realization of the study variable of
interest for element  $i$ in the finite population. The auxiliary variables are
cheap to measure but the measurement for the study variable is expensive.
Two-phase sampling is a popular method for cost effective sampling.
In two-phase sampling, the first phase sample $A_1$  is selected from the
target population to obtain the measurement of $\bx_i$ in the sample. The
sampling design for the first-phase sample can be denoted by $P(A_1)$. After
that, based on the observed auxiliary information in the first-phase  sample,
the second-phase sample $A_2 \subseteq A_1$ is selected from the first-phase
sample to observe $y_i$. The sampling design for the second-phase sample is
denoted as $P( A_2 \mid A_1)$.  
The
goal of two-phase sampling is to construct an estimator of $Y_N =  \sum_{i=1}^N y_i$ 
that uses both the observed information in the Phase 2 sample and also the 
auxiliary information from $\bf X \in \R^p$ in the Phase 1 sample. How to
combine information efficiently is an important problem in two-phase sampling.

Let $\pi_{1i} = \sum_{A_1: i \in A_1} \Pr(A_1)$ be the first-order inclusion
probability for the first-phase sample and let $\pi_{2i | 1} = \sum_{A_2: i \in
A_2} \Pr(A_2 \mid A_1)$ be the first-order inclusion probability for the
second-phase sample. We write $d_{1i} = \pi_{1i}^{-1}$ and $d_{2i|1} =
\pi_{2i|1}^{-1}$ to denote the design weights.
From the first-phase sample, we can compute
$\widehat{\mathbf{X}}_1 = \sum_{i \in A_1} d_{1i}\bx_i$ as an unbiased
estimator of $\mathbf{X}_N = \sum_{i=1}^N  \bx_i$. Also, from the second-phase
sample, we can compute $\left( \widehat{\mathbf{X}}_2 , \hat{Y}_2 \right) =
\sum_{i \in A_2} d_{1i} d_{2i \mid 1} \left( \bx_i, y_i\right)$ as
an unbiased estimator of $(\mathbf{X}_N, Y_N)$.
We therefore have three estimators for
two parameters. The two-phase regression estimator can be  used to combine this
information efficiently and it can be expressed as  
\begin{equation} \label{eq:1}
\hat Y_{\rm reg, tp} 
= \sum_{i \in A_1} d_{1i} \bf x_i^\top  \hat{\bm \beta_q} + 
\sum_{i \in A_2} d_{1i} d_{2i|1} \left(y_i - \bf x_i^\top  \hat{\bm \beta_q}\right)
\end{equation} 
where $q_i = q(\bf x_i)$  is a function of $\bf x_i$ and  
$$
\hat{\bm \beta_q} = \left(\sum_{i \in A_2} 
d_{1i} \bf x_i \bf x_i^\top q_i \right)^{-1} 
 \sum_{i \in A_2} d_{1i}\bf{x}_i y_i q_i.
$$

Regarding the choice of $q_i$, one can consider a linear regression model 
\begin{equation}
y_i = \bx_i^\top \bm \beta + e_i 
\label{regression}
\end{equation}
where $e_i \sim \left(0, \sigma^2 c_i \right)$, where $c_i=c( \bx_i)$ is a known
function. In this case,  $q_i = c_i^{-1}$ is a reasonable choice. 
One way to understand the two-phase regression estimator in \eqref{eq:1}
is to use the following decomposition: 
$$ 
\hat Y_{\rm reg, tp} 
= \underbrace{\sum_{i \in A_1} d_{1i}\bf x_i^\top \hat{\bm \beta_q}}_{
  \text{ Prediction  }} + \underbrace{\sum_{i \in A_2}
d_{1i} d_{2i|1}(y_i - \bf x_i^\top \hat{\bm \beta_q})}_{
\text{Bias correction}}.
$$
The prediction term is obtained by finding the best linear unbiased predictor
under the working superpopulation model. The bias correction term 
makes the two-phase regression estimator asymptotically design unbiased
regardless if the underlying the superpopulation model is correct or not.
We can find an equivalent estimator in the form $\hat Y_{\rm reg, tp} =
\sum_{i \in A_2} d_{1i} \hat w_{2i} y_i$. This is useful because it shows that
the estimator is linear and it is interesting because the estimated weights
$\hat w_{2i}$ are the weights closest to the design weights that satisfy the
constraints from \eqref{eq:constr}.
\begin{equation}\label{eq:constr}
\sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
= \sum_{i \in A_1} d_{1i} \bf{x}_i.
\end{equation}
Closeness is in an $L^2$-sense. Overall, the weights $\hat w_{2i}$ are the
solution to the following optimization problem:
$$
\hat w_{2} = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i|1})^2 q_i
\text{ such that } \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i 
= \sum_{i \in A_1} d_{1i} \bf{x}_i.
$$

This means that $\hat Y_{\rm reg, tp}$ is also a calibration estimator. The idea
that regression estimation is a form of calibration was noted by
\cite{deville1992calibration} and extended by them to consider loss functions
other than squared loss. Their generalized loss function minimizes
$\sum_i G(w_i, d_i)q_i$ for weights $w_i$ and design-weights $d_i$ where
$G(\cdot)$ is a non-negative, strictly convex function with respect to $w$,
defined on an interval containing $d_i$, with $g(w_i, d_i) = \partial G /
\partial w$ being continuous with respect 
to $w_i$.\footnote{The \cite{deville1992calibration} paper
considers regression estimators for a single phase setup, which we apply to our
two-phase example.} This
generalization includes empirical likelihood estimation, and maximum entropy
estimation among others \citep{schennach2007point}.

The \cite{deville1992calibration} method incorporates the design weights into
the loss function, which is the part minimizing the variance. We would rather
have bias calibration separate from the minimizing the variance so that
we can control each in isolation. 
\cite{kwon2024debiased} shows that for a generalized entropy
function $G(w)$, including a term of $g(d_{2i|1})$ into the calibration
for $g = \partial G / \partial w$ not only creates a design consistent
estimator, but it also has better efficiency than the generalized regression
estimators of \cite{deville1992calibration}.

However, the method of \cite{kwon2024debiased} assumes known finite population 
calibration levels. It does not handle the
two-phase setup where we need to estimate the finite population total of $\bf x$
from the Phase 1 sample. In the rest of this paper, we extend their method to 
two-phase sampling so that we construct asymptotically valid confidence
intervals when using estimated finite population totals derived from the Phase 1
sample.

\section{Main Proposal}

We follow the approach of \cite{kwon2024debiased} for the debiased calibration
method.  Let $G: \mathcal{V} \to \R$ be the generalized entropy function that is
strictly convex and differentiable. We consider maximizing 
\begin{equation}\label{eq:primalloss}
  H(w_2) = - \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i
\end{equation}
subject to the following constraints:
\begin{equation}\label{eq:calconst1}
  \sum_{i \in A_2} d_{1i} w_{2i} \bf{x}_i = 
  \sum_{i \in A_1} d_{1i}\bf{x}_i
\end{equation}
and 
\begin{equation}\label{eq:calconst2}
  \sum_{i \in A_2} d_{1i} w_{2i} g(d_{2i|1})q_i = 
  \sum_{i \in A_1} d_{1i} g(d_{2i|1})q_i, 
\end{equation}
where $g(w) = \frac{\partial G}{\partial w}$.

Also known as covariate balancing \citep{imai2014covariate},
constraint~\eqref{eq:calconst1} ensures that the weighted mean of the
covariates, $\bf{x}_i$, are the same.
This is also the usual calibration constraint motivated from
the regression model in \eqref{regression}.
Constraint~\eqref{eq:calconst2} is how we incorporate the design weights. Using
$g(w) = \frac{\partial G}{\partial w}$, ultimately ensures that $\hat w_{2i}$
converges to $d_{2i}$ in probability. This 
ensures that design consistency is achieved. 
The original method of \cite{kwon2024debiased} only considers known finite
population quantities on the right hand side of \eqref{eq:calconst1}. We extend
their method within our two-phase framework to allow for an estimated quantity
on the right hand side of \eqref{eq:calconst1}.
To estimate the weights, we solve
\begin{equation}\label{eq:primal}
  \hat w_2  = \argmin_{w_{2}} \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
  \text{ such that}
  \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i = \sum_{i \in A_1} d_{1i} \bf z_i 
\end{equation}
where  $\bf z_i = (\bf x_i , g(d_{2i|1})q_i)$. The resulting 
estimator of $\bar Y_N$ is 
\begin{equation}\label{eq:tpest}
  \hat Y_{\rm DCE} = N^{-1} \sum_{i \in A_2} d_{1i} \hat w_{2i} y_i.
\end{equation}

\subsection{Theoretical Results}

We compute the estimated weights, $\hat w_{2i}$ from
\eqref{eq:primal} using the method of Lagrange multipliers. We 
minimize the Lagrangian function 
\begin{equation}\label{eq:legragedc1}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} d_{1i} G(w_{2i}) q_i 
  + \bm \lambda^\top \left( \sum_{i \in A_1} d_{1i} \bf z_i -
    \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i\right)
\end{equation}
where $\bm \lambda \in \R^{p+1}$ is a vector of Lagrange multipliers and
$\bf{z}_i = (\bf{x}_i^\top, g(d_{2i|1})q_i)^\top$.
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero, we can obtain 
$$ \hat w_{2i} = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
where $g(w) = \frac{\partial G}{\partial w}$ and 
$\hat{\bm \lambda}$ is the solution to
\begin{equation}\label{eq:lamdc1}
  \sum_{i \in A_1} d_{1i} \bf{z}_i  -
  \sum_{i \in A_2} d_{1i} w_{2i}(\hat{\bm \lambda}) \bf{z}_i = 0.
\end{equation}
Under regularity conditions presented in the appendix, this solution is unique.
The following theorem presents a linearization of
\eqref{eq:primal} which leads to a design consistent estimator 
$\hat Y_{\rm DCE}$ in \eqref{eq:tpest}.

\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$. 
  Under the regularity conditions stated in the appendix
  $$
  \hat Y_{\rm DCE} = \hat Y_\ell(\bm{\lambda}^*, \bm \phi^*) + O_p(N / n_2)
  $$
  where
  $$
  \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
  + \left(\sum_{i \in A_1} d_{1i} \bf z_i - 
  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\bm \phi^*,
  $$
  and
  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}\bf z_i y_i}{g'(d_{2i|1})q_i}.
  $$

%   \begin{align*}
%   \hat Y_{\ell}(\bm \lambda^*, \bm \phi^*) 
%   &= \hat{Y}_{\rm DCE} + \left(\sum_{i \in A_1} d_{1i} \bf x_i  -
%     \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1} \bf x_i  \right)^\top \bm \phi_1^* \\
%   &\quad+ \left(\sum_{i \in A_1} d_{1i} g_i  -
%  \sum_{i \in A_2} d_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^\top \phi_2^*  
%  \end{align*} 
%\textcolor{red}{(You introduced $\hat{Y}_{\ell}$ twice!)}
%  and  
%
%  $$
%  \begin{pmatrix}
%  \bm \phi_1^* \\
%  \phi_2^* 
%  \end{pmatrix}
%  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
%  \begin{pmatrix}
%  \bx_i \bx_i^\top &   \bx_i g_i   \\
%  g_i  \bx_i^\top   & g_i^2     
%  \end{pmatrix} \right]^{-1}
%  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} 
%  \begin{pmatrix}
%    \bx_i \\ g_i 
%  \end{pmatrix}y_i 
%  $$

% with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
\end{theorem}

See the appendix for the proof of Theorem~\ref{thm:dc1}. Notice that this
theorem implies
that $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ and $\hat Y_{\rm DCE}$ have
the same asymptotic expectation and variance, 
which means that we can estimate
the variance of $\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ to get an
estimate of the variance of $\hat Y_{\rm DCE}$. The basic idea of 
Theorem~\ref{thm:dc1} is that asymptotically the two-phase debiased calibration
method is equivalent to a two-phase regression estimator. By Theorem~\ref{thm:dc1}, 
$\hat Y_{\ell}(\bm \lambda^*, \bm \phi^*)$ is design consistent,
and the proposed calibration estimator in \eqref{eq:primal} is
asymptotically equivalent to the two-phase regression estimator given by 
  $$
  \hat Y_{\rm tp, reg}  = 
  \sum_{i \in A_2} d_{1i} d_{2i|1} y_i
  + \left(\sum_{i \in A_1} d_{1i} \bf z_i  - 
  \sum_{i \in A_2} d_{1i} d_{2i|1} \bf{z}_i \right)\hat{\bm \phi},
  $$
where $\hat{\bm \phi}$ is defined in \eqref{eq:8}
\begin{equation}
\hat{\bm \phi} = 
\left[\sum_{i \in A_2} 
d_{1i}\frac{\bf z_i \bf z_i^\top}{g'(d_{2i|1})q_i}\right]^{-1}
\sum_{i \in A_2} d_{1i}\frac{\bf z_i y_i}{g'(d_{2i|1})q_i}.
\label{eq:8}
\end{equation}

For specific sample designs, we can construct the optimal entropy function
$G(\cdot)$. From Theorem~\ref{thm:dc1}, the choice of entropy function affects
the variance of the debiased calibration estimator through the form of $g'(d_{2i|1})$
in $\bm \phi^*$. Consider a class of generalized regression estimators,
$$
\mathbf{B} = 
\left\{\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \gamma + 
\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \gamma):
\bm \gamma \in \R^{p+1}\right\}
$$
The design optimal estimator is 
$$
\hat \theta_{\ell, \rm{opt}} = 
\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \phi_{\rm opt} + 
\sum_{i \in A_2} d_{1i}d_{2i|1}(y_i - \bf{z}_i^\top \bm \phi_{\rm opt}):
$$
where 
$$
\bm \phi_{\rm opt} = 
\Var\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i\right)^{-1} 
\Cov\left(\sum_{i \in A_2} d_{2i|1} \bf{z}_i, \sum_{i \in A_2}, d_{2i|1} y_i \right).
$$
We assume that both $A_1$ and $A_2$ are Poisson samples. Then,
$$
\bm \phi_{\rm opt} = 
\left\{
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i}\bf{z_i}^\top
\right\}^{-1}
\sum_{i \in U} \left(d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1)\right) \bf{z_i} y_i.
$$
To find the optimal entropy function, we solve the differential equation
$$
\frac{1}{d_{2i|1} g'(d_{2i|1})} = d_{2i|1} - 1 + d_{2i|1}(d_{1i} - 1).
$$
Using the fact that $(y(xy-1))^{-1} = -y^{-1} + x(xy-1)^{-1}$, the optimal
entropy function is 
$$
G(w_{2i|1}) = - \log(w_{2i|1}) + \log(w_{2i|1}d_{1i} - 1).
$$
To construct an asymptotically unbiased estimate of the variance we assume that
the sampling designs for $A_1$ and $A_2$ are measurable and we define the
following joint inclusion probabilities:
$$
\Pr(i \in A_1, j \in A_1) = \pi_{1ij}, \text{ and }
\Pr(i \in A_2, j \in A_2 \mid i \in A_1, j \in A_1) = \pi_{2ij|1}.
$$
The next theorem discusses how to compute an estimate for the variance of 
$\hat Y_{\rm DCE}$. For the proof see the appendix. The idea behind the proof is
to use the two-phase sampling approach to estimate the variance. If the sampling
fraction $n_2 / n_1 \to 0$ as $N$ increases, then the reverse framework of
\cite{fay1992inferences} allows us to make a good approximation.

\begin{theorem}[Variance Estimation]\label{thm:mainvar}

  Define $\hat \eta_i = \bf z_i \hat{\bm \phi} + 
  \delta_i d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})$.
  We can construct an unbiased estimate the variance of $\hat Y_{\rm DCE}$ with
  $$
  \hat V_{\rm DCE} 
  = \sum_{i \in A_1} \sum_{j \in A_1} d_{1ij} \Delta_{1ij}
  \hat \eta_i \hat \eta_j
  + \sum_{i \in A_2} \sum_{j \in A_2} d_{1i} d_{2ij|1}\Delta_{2ij|1}
  d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})
  d_{2j|1}(y_j - \bf z_j \hat{\bm \phi})
  $$
  where $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$,
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
  $d_{1ij} = \pi_{1ij}^{-1}$, $d_{2ij|1} = \pi_{2ij|1}^{-1}$,
  $\pi_{1ij} = \Pr(i \in A_1, j \in A_1)$, and 
  $\pi_{1ij} = \Pr(i \in A_2, j \in A_2 | i \in A_1, j \in A_1)$.
\end{theorem}

\section{Extensions}

\subsection{Non-nested Two-Phase Sampling}

We now consider the sampling mechanism known as non-nested two-phase sampling 
\citep{hidiroglou2001double}. In the last section, we considered two-phase sampling
in which the Phase 2 sample was a subset of the Phase 1 sample. With non-nested
two-phase sampling the Phase 2 sample is independent of the Phase 1 sample. 
Like traditional two-phase sampling, we consider the Phase 1 sample, $A_1$, to
consist of observations of $(\bf X_i)$ and the Phase 2 sample,
$A_2$, to consist of observations of $(\bf X_i, Y_i)$. We denote the sample size
of $A_1$ as $n_1$ and the sample size of $A_2$ as $n_2$.

Whereas the classical two-phase estimator uses a single Horvitz-Thompson
estimator of the Phase 1 sample to construct estimates for calibration totals,
in the non-nested two-phase sample we have two independent Horvitz-Thompson
estimators of the total of $\bf X_N$,
$$\hat{\bf X}_1 = \sum_{i \in A_1}^{n_1} d_{1i} \bf x_i \text{ and } 
\hat{\bf X}_2 = \sum_{i \in A_2}^{n_2} d_{2i} \bf x_i $$
where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, 
$\pi_{1i} = \Pr(i \in A_1)$ and $\pi_{2i} = \Pr(i \in A_2)$. 
We can combine these estimates using the effective sample size 
\citep{kish1965survey} to get
$\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
(n_{1, \rm{eff}} + n_{2, \rm{eff}})$ where $n_{1, \rm{eff}}$ and 
$n_{2, \rm{eff}}$ are the effective sample size for $A_1$ and $A_2$ respectively.
Then we can define a regression estimator as
\begin{equation}\label{eq:nntpreg}
\hat Y_{\rm NN, reg} = 
\hat Y_2 + (\hat{\bf X}_c - \hat{\bf X}_2)^\top \hat{\bm \beta_q} = 
\hat Y_2 + (\hat{\bf X}_1 - \hat{\bf X}_2)^\top W\hat{\bm \beta_q}
\end{equation}
where, $W = n_{1, \rm{eff}} / (n_{1, \rm{eff}} + n_{2, \rm{eff}})$, and
$$
\hat{\bm \beta_q} = 
\left(\sum_{i \in A_2} \bf x_i \bf x_i^\top q_i\right)^{-1} 
\sum_{i \in A_2} \bf x_i y_i q_i \text{ and }
\hat Y_2 = \sum_{i \in A_2} d_{2i} y_i.
$$

From an optimality perspective the choice of using the effective sample size to
weight the estimates from $A_1$ and $A_2$ is reasonable because the
effective sample sizes are often proportional to the variance of the estimates 
of $\bf X_N$. While the inverse variance weighted estimate is optimal
to combine independent samples for a linear estimate, using the effective sample
size approximates this procedure without requiring the actual variance of an
estimator to be known. Since the samples $A_1$ and $A_2$ are independent, 
$$
V(\hat Y_{\rm NN, reg}) = 
V\left(\sum_{i \in A_2} d_{2i}(y_i - \bf x_i^\top W\bm \beta^*_q)\right)
+ (\bm \beta^*_q)^\top W^\top V(\hat{\bf X}_1) W \bm \beta_q^*
$$
where $\bm \beta_q^*$ is the probability limit of $\hat{\bm \beta_q}$. Like the
two-phase sample, the regression estimator in \eqref{eq:nntpreg} can be viewed
as a linear estimator given a particular choice of weights $\hat w_{2i}$. These
weights are the solution to the calibration equation in \eqref{eq:nncal}.
\begin{equation}\label{eq:nncal}
  \hat w_2 = \argmin_{w_2} \sum_{i \in A_2} (w_{2i} - d_{2i})^2 q_i 
  \text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c
\end{equation}
These estimated weights yield an estimator of $\theta = E[Y]$ of 
$\hat Y_{\rm NN, reg} = N^{-1} \sum_{i \in A_2} \hat w_{2i} y_i$. 
We can extend the debiased calibration estimator of \cite{kwon2024debiased} to
the non-nested two-phase sampling case where we use a combined estimate
$\hat{\bm X_c}$ as the calibration totals instead of using the true totals from
the finite population.

%\subsubsection{Theoretical Results}

The methodology for the non-nested two-phase sample is very similar to the setup
described as part of Topic 1. Given a strictly convex differentiable function,
$G: \mathcal{V} \to \mathbb{R}$, the goal is to minimize the variance of the
estimator while ensuring that the covariate balancing condition is met. This
results in $\hat w_2$ being the solution to the optimization problem in 
\eqref{eq:nnopt}
\begin{equation}\label{eq:nnopt}
\hat w_2 = \argmin_{w_2} \sum_{i \in A_2} G\left(w_{2i}\right) q_i 
\text{ where } 
\sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_{c} \text{ and } 
\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
\end{equation}
for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
The difference
between solving \eqref{eq:nnopt} and \eqref{eq:primal} is that the
estimator $\hat{\bf X}_c$ is estimated from the combined sample $A_c = A_1 \cup
A_2$. Before using $\hat{\bf X}_c$ in the debiased calibration estimator, we
need to estimate it from the non-nested samples. We can get multiple estimates
of $\bf X_N$, 
$$
\hat{\bf X}_1 = \sum_{i \in A_1} d_{1i} \bf x_i \text{ and }
\hat{\bf X}_2 = \sum_{i \in A_2} d_{2i} \bf x_i.
$$
If $n_{1, \rm{eff}}$ and $n_{2, \rm{eff}}$ are the effective samples sizes for
$A_1$ and $A_2$ respectively, then the optimal combined estimate is 
$$
\hat{\bf X}_c = (n_{1, \rm{eff}} \hat{\bf X}_1 + n_{2, \rm{eff}}\hat{\bf X}_2) / 
(n_{1, \rm{eff}} + n_{2, \rm{eff}})
$$
We can construct a non-nested two-phase estimator $\hat Y_{\rm NNE}$ for 
$\bar Y_N$ where
$\hat Y_{\rm NNE} = N^{-1} \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$
solves \eqref{eq:nnopt}. Like the classical two-phase approach, to solve this
setup we minimize the Lagrangian,
\begin{equation}\label{eq:legragedc2}
  L(w_{2i}, \bm \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + 
  \bm \lambda^\top \left( \hat{\bf T} - \sum_{i \in A_2} w_{2i} \bf z_i \right).
\end{equation}
where $\bm \lambda$ are the Lagrange multipliers, $\bf{z}_i = (\bf{x}_i^\top, g(d_{2i})q_i)^\top$,
and
$$\hat{\bf T} = 
\begin{bmatrix}
\hat{\bf X}_c \\ \sum_{i \in U} g(d_{2i}) q_i
\end{bmatrix}.
$$
Differentiating with respect to $w_{2i}$ and setting this expression equal to
zero yields the fact that $\hat w_{2i}$ satisfies 
$$ \hat w_{2i}(\hat{\bm \lambda}) = g^{-1}(\hat{\bm \lambda}^\top \bf z_i) $$
where $\hat{\bm \lambda}$ is the solution to
\begin{equation}\label{eq:lamdc2}
  \hat{\bf T} - \sum_{i \in A_2} w_{2i}(\hat{\bm \lambda}) \bf z_i = 0.
\end{equation}
Like the previous section we first prove that the estimator is 
asymptotically equivalent to a regression estimator. This ensures that it is
design consistent. The asymptotic equivalence is useful because our estimator
now has the same variance as a regression estimator. Using this fact,
we provide a result showing how to estimate the variance.

\begin{theorem}[Design Consistency]\label{thm:dc2}
  Let $\bm \lambda^*$ to be the probability limit of $\hat{\bm \lambda}$,
  under some regularity conditions,
  $\hat Y_{\rm NNE} = \hat Y_{\ell, \rm NNE}(\bm \lambda^*,
  \bm \phi^*) + O_p(Nn_2^{-1})$ where
  $$
  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_2} d_{2i} y_i + 
  \left(\hat{\bf T} - \sum_{i \in A_2} d_{2i} \bf z_i \right)^\top \bm \phi^*
  $$
  and 
  $$
  \bm \phi^* =
  \left(\sum_{i \in U} \frac{\pi_{2i}}{g'(d_{2i})q_i} 
  \begin{bmatrix}
    \bf x_i^2 & \bf x_i g(d_{2i}) q_i\\
    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
  \end{bmatrix}
  \right)^{-1}
  \sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})q_i} 
  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
  $$
\end{theorem}

The proof of this result is very similar to the proof the
Theorem~\ref{thm:dc1}. The biggest difference is that the total for $\bf X_N$ is
estimated from both samples using $\hat{\bf X}_c$ instead of 
$\hat{\bf X}_{\rm HT}$ from the Phase 1 sample. 
By the dual formulation of \eqref{eq:legragedc2}, for any $\bm \phi$
$$\hat Y_{\rm NNE} (\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} \hat
w_{2i}(\hat{\bm \lambda})y_i + \left(\bf T - \sum_{i \in A_2} \hat
w_{2i}(\hat{\bm \lambda}) \bf z_i \right)^\top \bm \phi.$$
If we choose $\bm \phi^*$ such that $E\left[\frac{\partial}{\partial \bm \lambda} 
  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)\right] = 0$, then
$$\bm \phi^* =
\begin{bmatrix}
  \bm \phi^*_1 \\ \bm \phi^*_2
\end{bmatrix} = 
\left(\sum_{i \in U} \frac{\pi_{2i}}{g'(d_{2i})q_i} 
\begin{bmatrix}
  \bf x_i^2  & \bf x_i g(d_{2i}) q_i \\
  \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
\end{bmatrix}
\right)^{-1}
\sum_{i \in U} \frac{\pi_{2i} y_i}{g'(d_{2i})q_i} 
\begin{bmatrix} \bf x_i \\ g(d_i) q_i\end{bmatrix}.
$$
Hence, by a Taylor expansion around $\hat{\bm \lambda}$,
$$\hat Y_{\rm NNE}(\hat{\bm \lambda}) = \hat Y_{\ell, \rm NNE}(\bm \lambda^*, 
\bm \phi^*) + O_p(Nn_2^{-1}).$$
The next theorem shows how to estimate the variance of $\hat Y_{\rm NNE}$.
Since we have already proves that the estimator is asymptotically equivalent to
a type of regression estimator, we only need to compute the variance of the
regression estimator to get the variance of $\hat Y_{\rm NNE}$.
The proof of this result is in the appendix.

\begin{theorem}[Variance Estimation]\label{thm:nnve}
  The variance of $\hat Y_{\rm NNE}$ is 
  \begin{align*}
    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda}))
    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i}
    (y_i - \bf z_i \bm \phi^* q_i)d_{2j}(y_j - \bf z_j \bm \phi^* q_j)\\
    &\qquad + (1 - W)\bm \phi_1^* \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}
    d_{2i} \bf x_i d_{2j}(y_j - \bf z_j \bm \phi^*_1 q_j)\\
  \end{align*}
  We can estimate the variance using
  \begin{align*}
    \hat V_{\rm NNE} 
    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X}_c) \hat{\bm \phi_1} + 
    \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij} \Delta_{2ij} d_{2i}
    (y_i - \bf z_i \hat{\bm \phi} q_i) d_{2j} (y_j - \bf z_j \hat{\bm \phi} q_j) \\
    &\qquad + (1 - W)\hat{\bm \phi_1} \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij} \Delta_{2ij}
    d_{2i}\bf x_i d_{2j}(y_j - \bf z_j \hat{\bm \phi_1} q_j)
  \end{align*}
  where $\Delta_{2ij} = \pi_{2ij} - \pi_{2i}\pi_{2j}$, $d_{2ij} = \pi_{2ij}^{-1}$, 
  $\pi_{2ij} = \Pr(i \in A_2, j \in A_2)$, and 
  $$\hat{\bm \phi} =
  \begin{bmatrix}
    \hat{\bm \phi_1} \\ \hat \phi_2
  \end{bmatrix} = 
  \left(\sum_{i \in A_2} \frac{1}{g'(d_{2i})q_i} 
  \begin{bmatrix}
    \bf x_i^2 & \bf x_i g(d_{2i}) q_i \\
    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
  \end{bmatrix}
  \right)^{-1}
  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i}) q_i} 
  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
  $$
\end{theorem}

\subsection{Multi-Source Two-Phase Sampling}

When considering non-nested two-phase sampling, we focused on the case of having
two samples. Now, we incorporate more than two independent samples
together with a debiasing constraint. We examine the case in which we
want to estimate $\theta = E[Y]$ where $Y$ is only observed in one sample.

Consider the setup in which we have independent samples $A_0, A_1, \dots, A_M$
where $Y$ is observed only in $A_0$ but $\bf X$ is observed in all of the
samples with variables $\bf X^{(0)}$ observed in $A_0$ and variables $\bf X^{(m)}$
observed in $A_m$ for each $m = 1, \dots, M$. Like the non-nested case, we
assume that each survey is sampled independently from the same population
sampling frame. Some basic examples of data structures that would benefit from
the multi-source two-phase sampling can be found in Table~\ref{tab:msexample}.

\begin{table}[ht!]
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule 
           & \multicolumn{3}{c}{Example 1} & \multicolumn{3}{c}{Example 2} \\
           \cmidrule(r){2-4}  
           \cmidrule(r){5-7}
    Sample & $X_1$ & $X_2$ & $Y$ & $X_1$ & $X_2$ & $Y$ \\
    \midrule
    $A_0$ & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark \\
    $A_1$ & \checkmark & &  & \checkmark & \checkmark & \\
    $A_2$ & & \checkmark &  & & \checkmark & \\
    \bottomrule
  \end{tabular}
  \caption{This tables shows two examples configuations of multi-source sampling.
  In each example, $A_0$ contains the variable of interest $Y$. For Example 1,
  the covariates in $A_1$ and $A_2$ are both contained in $A_0$. In Example 2,
  sample $A_2$ improves the estimation of $Y$ via improving the estimation of
  $X_1$ in $A_1$.}
  \label{tab:msexample}
\end{table}

The traditional multi-source approach \citep{kim2024statistics} is to use
GLS to obtain an optimal estimator of $\bf X_N$ from the
samples $A_1, \dots, A_M$. Then we can incorporate this information in the
estimation of $\theta$ by using the following estimate 
$$\hat \theta = \sum_{i \in A_0} \hat w_i y_i$$
where 
\begin{equation}\label{eq:mstps}
\hat w = \argmin_w \sum_{i \in A_0} G(w_i)q_i \text{ such that } 
\sum_{i \in A_0} w_i \bf x_i = \hat{\bf X}_{\rm{GLS}}^{(0)}
\end{equation}
where $\hat{\bf X}_{\rm{GLS}}^{(0)}$ is the GLS estimate of $\bf X$ for all of
the $X$-variables measured in sample $A_0$, and $G(\cdot)$ is a generalized entropy
function. In this setup, all of the data integration occurs when estimating
$\hat{\bf{X}}_{\rm GLS}$, which is used as a calibration constraint.
In order to have a design consistent estimator that incorporates information from
samples $A_0, A_1, \dots, A_M$, we want to add the debiasing constraint from 
\eqref{eq:msdebconstr} because it ensures design consistency of the estimated
weights
\begin{equation}\label{eq:msdebconstr}
\sum_{i \in A_0} w_{i} g(d_i) q_i = \sum_{i \in U} g(d_i) q_i
\end{equation}
where $g(x) = \partial G(x) / \partial x$ and $d_i$ are the design weights for
sample $A_0$. This yields the optimization problem:
\begin{equation}\label{eq:dc3}
  \hat w_0 = \argmin_{w_0} \sum_{i \in A_0} G(w_{0i}) q_i \text{ such that }
  \widehat{\bf{T}} = \sum_{i \in A_0} w_{0i} \bf z_i.
\end{equation}
where $\widehat{\bf{T}} = ((\hat{\bf X}_{\rm{GLS}}^{(0)})^\top,
\sum_{i \in U} g(d_i) q_i)^\top$,
and $\bf z_i = ((\bf x_i^{(0)})^\top, g(d_i)q_i)^\top$. Then
the estimator is $\hat Y_{\rm MS} = \sum_{i \in A_0} \hat w_{0i} y_i$ and we can
construct this estimate minimizing the Lagrangian:
$$
L(w_{0}, \bm \lambda) = \sum_{i \in A_0} G(w_{0i}) q_i + {\bm \lambda}^\top
\left(\widehat{\bf{T}} - \sum_{i \in A_0} w_{0i} \bf z_i\right)
$$
where $\bm \lambda$ are the Lagrange multipliers from \eqref{eq:dc3}.
We now show that $\hat Y_{\rm MS}$ is asymptotically equivalent to a regression
estimator. This linearization technique shows design consistency.
\begin{theorem}[Design Consistency]
  Let $\bm \lambda^*$ be the probability limit of $\hat{\bm \lambda}$.
  Under regularity conditions,
  $$
  \hat Y_{\rm MS} = \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) + O_p(N / n_0)
  $$
  where
  $$
  \hat Y_{\ell, \rm MS}(\bm \lambda^*, \bm \phi^*) = 
  \sum_{i \in A_0} d_{0i} y_i + \left(
    \widehat{\bf T} - \sum_{i \in A_0} d_{0i} \bf z_i
  \right)^\top \bm \phi^*
  $$
  and 
  $$
  \bm \phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{0i} \bf z_i \bf z_i^\top q_i}{g'(d_{0i})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{0i} \bf z_i y_i}{g'(d_{0i})}.
  $$
\end{theorem}

This proof follows from the proof of Theorem~\ref{thm:dc2}. The only
difference is that we have $\hat{\bf X}_{\rm GLS}$ instead of $\hat{\bf X}_c$.
Since we have shown equivalence between our estimator and a regression estimator
we can estimate the variance by giving the variance of the regression estimator.
The next theorem gives a method for how to estimate the variance of 
$\hat Y_{\rm MS}$.
\begin{theorem}[Variance Estimation]
  Under regularity conditions,
  $$
  \begin{aligned}
    V(\hat Y_{\rm MS}) 
    &= (\bm \phi^{*})^\top \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\bm \phi^{*}) + 
      \sum_{i \in U} \sum_{j \in U} \Delta_{0ij}
      d_{0i}(y_i - \bf z_i\bm \phi^{*}q_i)
      d_{0j}(y_j - \bf z_j\bm \phi^{*}q_j). 
  \end{aligned}
  $$
  We can estimate the variance with
  $$
  \begin{aligned}
    \hat V(\hat Y_{\rm MS}) 
    &= (\hat{\bm \phi})^\top \hat \Var(\hat{\bf X}_{\rm GLS}^{(0)}) (\hat{\bm \phi})
  + \sum_{i \in A_0} \sum_{j \in A_0} d_{0ij}\Delta_{0ij}
    d_{0i}(y_i - \bf z_i \hat{\bm \phi}q_i)
    d_{0j}(y_j - \bf z_j \hat{\bm \phi}q_j)
  \end{aligned}
  $$
  where $\Delta_{0ij} = \pi_{0ij} - \pi_{0i}\pi_{0j}$, $d_{0ij} = \pi_{0ij}^{-1}$,
  and $\pi_{0ij} = \Pr(i \in A_0, j\in A_0)$.
\end{theorem}
The result follows the same argument as Theorem~\ref{thm:nnve}. The result
holds because each sample $A_m$ is independent from the other $A_{m'}$.

\section{Simulation Studies}

\subsection{Simulation Study 1}

We run a simulation testing the main proposed method. In this approach we have the
following simulation setup:
$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} \rm{Unif}(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta 0.5 X_{3i} + \varepsilon_i \\
\pi_{1i} &= n_1 / N \\
\pi_{2i|1} &= \max(\min(\Phi_3(x_{3i} - 1), 0.7), 0.02).
\end{aligned}
$$
where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
consider a finite population of size $N = 10,000$ with the Phase 1 sampling
being a simple random sample (SRS) and the 
Phase 2 sampling occurring under Poisson sampling. The Phase 1 sample has
$n_1 = 1200$ and the Phase 2 sample has an expected size of
$E[n_2] \approx 2588$. In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
The parameter $\delta \in \{0, 1\}$ controls the effect of model misspecification
in the simulation. Let $\bf x_i = (1, x_{1i}, x_{2i})^\top$. We compare the
proposed method for the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] $\pi^*$-estimator (PiStar): $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two-Phase Regression estimator (Reg): 
    $\hat Y_{\rm reg} = 
    \sum_{i \in A_1} \frac{\bf x_i' \widehat{\bm \beta}}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \widehat{\bm \beta})$ 
    where $\widehat{\bm \beta} = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$.
  \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
    solves

  \begin{equation}
    \argmin_{w_{2|1}} \sum_{i \in A_2} d_{1i} G(w_{2i}) \text{ such that}
    \sum_{i \in A_2} d_{1i} w_{2i} \bf z_i = \sum_{i \in U} \bf z_i.
  \end{equation}

  \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
    This solves \eqref{eq:primal} with $q_i = 1$.
\end{itemize}

%In addition to estimating the mean parameter $\bar Y_N$, we also construct
%variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
%approach we give the variance estimate in Table~\ref{tab:varforms}.
%% Move variance calculations into an appendix?
%
%\begin{table}[ht!]
%  \centering
%  \begin{tabular}{lcc}
%    \toprule
%    Estimator & Estimated Variance & Notes \\
%    \midrule
%    $\pi^*$ 
%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
%      \sum\limits_{i \in A_2} \frac{d_{1i}^2}{N^2} 
%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
%    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
%    \bar y_i)^2$} \\
%    TP-Reg  
%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
%      \sum\limits_{i \in A_2} \frac{d_{1i}}{N^2} 
%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
%    & {\scriptsize$ \eta_i = \bf x_i'\bm \beta + \delta_{2i} w_{2i}\varepsilon_i,\;
%        \varepsilon_i = (y_i - \bf x_i' \bm \beta)$} \\
%    DC-Pop  
%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
%      \sum\limits_{i \in A_2} \frac{d_{1i}}{N^2} 
%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i}\varepsilon_i,\;
%        \varepsilon_i = (y_i - \bf z_i'\bm \phi)$} \\
%    DC-Est  
%    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
%      \sum\limits_{i \in A_2} \frac{d_{1i}}{N^2} 
%        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
%    & {\scriptsize$\eta_i = \bf z_i' \bm \phi + \delta_{2i}w_{2i}\varepsilon_i,\;
%        \varepsilon_i = (y_i - \bf z_i' \bm \phi)$} \\
%    \bottomrule
%  \end{tabular}
%  \caption{This table gives the formulas for each variance estimator used in
%  Simulation 1. For the derivation of the estimated variance of the DC-Pop
%estimatator see Appendix A.}
%  \label{tab:varforms}
%\end{table}

We run the simulation $B = 1000$ times for each of these methods and compute the
Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
simulation replicate. We also include the Monte Carlo variance of the estimated
value, ($V_{MC} = (B - 1)^{-1} \sum_{b = 1}^B (\hat Y^{(b)} - \bar Y_N)^2$), the
mean of the estimated variance, 
($\bar{\hat V} = B^{-1} \sum_{b = 1}^B \hat V^{(b)}$), and the relative bias of
the estimated variance, ($(V_{MC} - \bar{\hat V}) / V_{MC}$). The results are in
Table~\ref{tab:tpdc0-mean} and Table~\ref{tab:tpdc1-mean}.

\begin{table}[ht!]
  \centering
  \input{tables/twophasesampd0.tex}
\caption{This table shows the results of Simulation Study 1 with $\delta = 0$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: PiStar, Reg,
  EstPop, and Est.}
\label{tab:tpdc0-mean}
\end{table}

\begin{table}[ht!]
  \centering
  \input{tables/twophasesampd1.tex}
\caption{This table shows the results of Simulation Study 1 with $\delta = 1$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: PiStar, Reg,
  EstPop, and Est.}
\label{tab:tpdc1-mean}
\end{table}

\subsection{Simulation Study 2}

We run a simulation testing the non-nested extension. This is very similar to
Simulation 1. We have the following simulation setup:
$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} \rm{Unif}(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta X_{3i} \varepsilon_i \\
\pi_{1i} &= n_1 / N \\
\pi_{2i} &= \max(\min(\Phi_3(X_{3i} - 2.5), 0.9), 0.01).
\end{aligned}
$$
where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
We consider a finite population of size $N = 10,000$ with the Phase 1 
sampling being a simple random sample (SRS) of size $n_1 = 1000$. The Phase 2
sample is a Poisson sample with an expected sample size of about 300.
In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. 
If $\delta = 0$, then there is no model misspecification. However, if $\delta =
1$, then there is some model misspecification. We estimate
the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] HT-estimator (HT): $\hat Y_{\rm HT} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{2i}},$
  \item[2.] Regression estimator (Reg): Let $\hat Y_{\rm NN, reg} = \hat Y_{\rm HT} + (
  \hat{\bf X}_c - \hat{\bf X}_{2, \rm HT}) \hat{\bm \beta_2}$ where $\hat Y_{HT} =
\sum_{i \in A_2} d_{2i} y_i$, $\hat{\bf X}_c = W \hat{\bf X}_{1, \rm HT} + (1 -
W)\hat{\bf X}_{2, \rm HT}$, $W = n_{1,\rm{eff}} / (n_{1,\rm{eff}} + n_{2, \rm{eff}})$,
$\hat{\bf X}_{1, \rm HT} = \sum_{i \in A_1} d_{1i} \bf x_{i}$, $\hat{\bf X}_{2,
    \rm HT} = \sum_{i \in A_2} d_{2i} \bf x_{i}$, $\bf x_i = (1, x_{1i}, x_{2i})^\top$
    and

    $$ \hat{\bm \beta_2} = \left(\sum_{i \in A_2} \bf x_i \bf x_i^\top\right)^{-1}
    \sum_{i \in A_2} \bf x_i y_i.$$

    Then $\hat{\bar Y}_{\rm NN, reg} = \hat Y_{\rm NN, reg} / N$.
  \item[3.] Debiased Calibration with Population Constraints (EstPop): This 
    solves

  \begin{align*}
    \hat w_2 &= \argmin_{w_2} \sum_{i \in A_2} G(w_{2i}) \\
    &\quad\text{ such that } 
  \sum_{i \in A_2} w_{2i} \bf x_i = \sum_{i \in U} \bf x_i \text{ and } 
  \sum_{i \in A_2} w_{2i} g(d_{2i}) = \sum_{i \in U} g(d_{2i})
  \end{align*}

  \item[4.] Debiased Calibration with Estimated Population Constraints (Est):
    This solves \eqref{eq:nnopt} with $q_i = 1$.
\end{itemize}

In addition to estimating the mean parameter $\bar Y_N$, we also construct
variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
We run the simulation $B = 1000$ times for each of these methods and compute the
Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
simulation replicate. We also include the Monte Carlo variance of the estimated
value, ($V_{MC} = (B - 1)^{-1} \sum_{b = 1}^B (\hat Y^{(b)} - \bar Y_N)^2$), the
mean of the estimated variance, 
($\bar{\hat V} = B^{-1} \sum_{b = 1}^B \hat V^{(b)}$), and the relative bias of
the estimated variance, ($(V_{MC} - \bar{\hat V}) / V_{MC}$). The results are in
Table~\ref{tab:nndc0-mean} and Table~\ref{tab:nndc1-mean}.

\begin{table}[ht!]
  \centering
  \input{tables/nonnestedtpd0.tex}
\caption{This table shows the results of Simulation Study 2 with $\delta = 0$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: HT, Reg,
  EstPop, and Est.}
\label{tab:nndc0-mean}
\end{table}

\begin{table}[ht!]
  \centering
  \input{tables/nonnestedtpd1.tex}
\caption{This table shows the results of Simulation Study 2 with $\delta = 1$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: HT, Reg,
  EstPop, and Est.}
\label{tab:nndc1-mean}
\end{table}

\subsection{Simulation Study 3}

This simulation study tests the multi-source extension.
We have the following superpopulation model with $N = 10,000$ elements:
$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} \rm{Unif}(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(5, 1) \\
Z_i &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \delta Z_{i} + \varepsilon_i \\
\pi_{0i} &= \min(\max(\Phi(Z_i - 2), 0.02), 0.9)\\
\pi_{1i} &= n_1 / N\\
\pi_{2i} &= \Phi(X_{2i} - 2) \\
\end{aligned}
$$
Like the previous simulation studies, when $\delta = 1$, there is model
misspecification for the outcome model because we observe the following columns
in each sample:
\begin{table}[ht!]
  \centering
  \begin{tabular}{lrrrr}
    \toprule
    Sample & $X_1$ & $X_2$ & $X_3$ & $Y$ \\
    \midrule
   $A_0$  & \checkmark     & \checkmark     & \checkmark     &  \checkmark \\    
   $A_1$  & \checkmark     &       & \checkmark     &    \\  
   $A_2$  & \checkmark     & \checkmark     &       &    \\  
   \bottomrule
  \end{tabular}
\end{table}

For the sampling mechanism both $A_0$ and $A_2$ are selected using a Poisson
sample with response probabilities $\pi_{0i}$ and $\pi_{1i}$ respectively. The
sample $A_1$ is a simple random sample with $n_1 = 2000$. We compare four
different estimators for $\theta = E[Y]$.
\begin{itemize}
  \item[1.] Horvitz-Thompson estimator (HT): $\hat Y = N^{-1} \sum_{i \in A_0}
    \frac{y_i}{\pi_{0i}}$,
  \item[2.] Non-nested regression (NNReg): This is the non-nested regression from
    \eqref{eq:nnopt} with only using information from Samples $A_0$ and
    $A_1$,
  \item[3.] Multi-Source proposed (Est): This is the proposed estimator from
    \eqref{eq:dc3}, and
  \item[4.] Multi-Source population (EstPop): This is the proposed estimator with
    using the true value of $T_1$ from the population.
\end{itemize}
The simulation results are displayed in Table~\ref{tab:msdc0-mean} and
Table~\ref{tab:msdc1-mean}.

\begin{table}[ht!]
  \centering
  \input{tables/multisourced0.tex}
\caption{This table shows the results of Simulation Study 3 with $\delta = 0$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: HT, NNReg,
  EstPop, and Est.}
\label{tab:msdc0-mean}
\end{table}

\begin{table}[ht!]
  \centering
  \input{tables/multisourced1.tex}
\caption{This table shows the results of Simulation Study 3 with $\delta = 1$.
  It displays the Bias, RMSE, empirical 95\% confidence interval, a t-statistic
  assessing the unbiasedness, the Monte Carlo variance, mean estimated variance
  and relative bias of the variance estimator for the estimators: HT, NNReg,
  EstPop, and Est.}
\label{tab:msdc1-mean}
\end{table}

As expected, EstPop has the lowest RMSE because it also uses the population
totals. The Est estimator outperforms
NNReg because it also uses information from $A_2$, even though it implicitly
uses a regression estimator with $X_3$ as a covariate, which is unnecessary.

\section{Real Data Analysis}

We apply our method to Medicare Current Beneficiary Survey (MCBS). The MCBS
is an ongoing representative national survey of the population of people on
Medicare in the United States, and in 2020, NORC started issuing an MCBS
supplement to identify the impact of COVID-19 on the Medicare population
\citep{norc2020mcbs}. This analysis looks at whether
respondents had obtained at least one dose of a COVID-19 vaccine.

Assessing the vaccine rate of Medicare patients is important from a public
health perspective. Older people\footnote{Medicare is available to people ages
65 years and older and select designated populations between the ages of 18 and
65.}
are known to have a higher all-cause mortality from
COVID-19 \citep{bonanad2020effect}. Thus, understanding what causes older people
to get vaccinated can help save lives and reduce medical costs.

The MCBS is a rotating panel survey in which participants are interviewed up to
three times within a four-year period \citep{cms2021mcbs}. The participants
are selected from Medicare enrollment data and enter a cohort in the fall round
of the survey. They then respond to the winter, summer,
and fall rounds of the survey for three consecutive years before exiting the
survey in the winter round of year four \citep{cms2021mcbs}. Each year a new
cohort is added. Participants are selected based on a three-stage cluster sample
design. The first stage of the sample consists of major metropolitan areas and
groups of rural counties \citep{cms2021mcbs}. The second stage are census
tracts within the primary sampling units. Finally, Medicare beneficiaries are
selected using a stratified systematic sample with random starts 
\citep{cms2021mcbs}. Medicare beneficiaries are selected from strata based on
age and whether the beneficiary is Hispanic. Strata with Hispanic beneficiaries
and strata with beneficiaries over the age of 85 and under the age of 65 are
oversampled.
The COVID supplement was added to the traditional MCBS survey in March of
2020 to assess the impact of COVID on the Medicare
population \citep{cms2021covid}. These questions about vaccines were added to
participants who were enrolled in the Summer 2021 cohort.

We combine the 2021 Summer MCBS survey with data from the 2021 American
Community Survey (ACS)
issued by the U.S. Census Bureau along with administrative data about people on
Medicare from the Center of Medicare and Medicaid Services (CMS). Like the MCBS
survey, the ACS data contains information about the overall distribution of
sex, race and ethnicity, education, and marital status for people on
Medicare. We combine this information with official totals of people on Medicare
by sex from the CMS. A complete display of the data available can be seen in
Table~\ref{tab:mcbscols}.

\begin{table}[ht!]
  \label{tab:mcbscols}
  \centering
  \begin{tabular}{lrrrrr}
  \toprule
  Data & Sex & Race and Ethnicity & Education & Marital Status & COVID Vaccine Dose \\
  \midrule
  MCBS & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
  ACS & \checkmark & \checkmark & \checkmark & \checkmark &  \\
  CMS & \checkmark &  &  &  & \\
  \bottomrule
  \end{tabular}
  \caption{This table shows the data available to us in each data set. A
  check mark in a particular column indicates that the data set listed in the
  left column contains information about the participant characteristic in the
  given column.}
\end{table}

To make the results between the ACS and MCBS compatible, we recode the 
ACS variables with the map found in Table~\ref{tab:acsvars}.

\begin{table}[ht!]
  \centering
  \label{tab:acsvars}
  \begin{tabular}{lr}
  \toprule
  Variable & Consolidated ACS Variables \\
  \midrule
  Education & SCHL:  01-15, 16-19, 20-24 \\
  Marital Status & MAR: 1, 2, 3-4, 5 \\
  \bottomrule
  \end{tabular}
  \caption{This table shows how we consolidated ACS variables. The right column
  displays the ACS category code and the values used in a particular category to
  match a MCBS category. For example, SCHL: 01-15 means that the values of 01 -
  15 of the SCHL ACS column were combined to match an individual MCBS value.}
\end{table}

While the MCBS data is largely free of missing values, there are a small number
that need to be imputed for the $Y$ variable about whether someone received at
least one COVID vaccination dose. This MCBS data was updated to include
information about whether a participant had received a second vaccine dose. 
Later, the survey included the total vaccine doses patients had received. If the
participant had received at least one more dose and their answer was missing
(because they refused to answer, did not know, or was missing) the missing value
was imputed to say that they yes, they had received a COVID vaccine dose.
Otherwise, the missing value was imputed to say no, the participant had not
received a COVID vaccine dose.

We analyze the data with three approaches: a Horvitz-Thompson method, a
regression algorithm, and our preposed multi-source two-phase sampling technique. 
The Horvitz-Thompson estimator uses the design weights, the MCBS observation
whether an individual received a COVID-19 vaccine dose, and the population total
from the CMS data about the total number of people on Medicare in 2021. The
regression estimator uses all of this and a population level total for the
number of males. Our multi-source algorithm also includes information from the
ACS about education and marital status. Due to the fact that we do not observe
the probability that an individual will be selected into the MCBS survey within
the population of all Medicare patients, we have to estimate the term $\sum_{i
\in U} g(d_i)$ from \eqref{eq:msdebconstr}. This technique is adopted
from \cite{kwon2024debiased} and described in more details in the appendix.
The results of the real data analysis are found in Table~\ref{tab:mcbsres}.

\begin{table}[ht!]
  \centering
  \label{tab:mcbsres}
  \begin{tabular}{lrr}
    \toprule
    Method & Point Estimate & Standard Error \\
    \midrule
    HT  & 0.817 & 0.01223 \\
    Reg & 0.877 & 0.00439 \\
    Est & 0.863 & 0.00436 \\ 
    \bottomrule
  \end{tabular}
  \caption{This table shows the point estimate and standard error of the
  percentage of people on Medicare who had at least one COVID-19 vaccine in
  2021 as estimated by a Horvitz-Thompson estimator (HT), a regression estimator
  (Reg) and the proposed multi-source estimator (Est).}
\end{table}

\section{Conclusion}

Overall, using debiased calibration for generalized two-phase sampling seems to
be a promising approach for combining multiple samples efficiently. Not only can
this method be used to combine surveys focusing on the same population frame,
but because it does not require individual responses levels from the
supplemental surveys, this method can also be used to construct domain level
estimates as exemplified in the real data analysis of Medicare patients.

\bibliographystyle{apa}
\bibliography{references.bib}

\newpage
\appendix

\section{Appendix: Regularity Conditions}

We need the following regularity conditions to prove Theorem 1.

\begin{itemize}
  \item[A1]: $\mathcal{V} \in \R$ is open,
  \item[A2]: $G(w)$ is strictly convex and twice continuously differentiable on
    $\mathcal{V}$ with a bounded second derivative,
  \item[A3]: There exist positives constants $c_1, c_2 \in \mathcal{V}$ such
    that $c_1 < \pi_{2i|1} < c_2$ for all $\pi_i$,
  \item[A4]: For $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
    $$
    \limsup_{N \to \infty} N \max_{i, j \in U, i \neq j} |\Delta_{2ij|1}| <
    \infty,
    $$
  \item[A5]: The matrix $\bm \Sigma_{\bf z} := 
    \lim_{N \to \infty} N^{-1} \sum_{i \in U} {\bf z}_i {\bf z}_i^\top$ exists
    and is positive definite, 
  \item[A6]: The average fourth moment of $({\bf x}_i^\top, y_i)^\top$ is finite
    such that 
    $$
    \limsup_{N \to \infty} N^{-1} \sum_{i = 1}^N ({\bf x}_i^\top, y_i)^\top < \infty,
    $$
  \item[A7]: $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$
\end{itemize}

Assumption A1 is needed for the differentiability of $G(w)$ in A2.
Assumption A2 guarentees a unique solution for $\hat{\bm \lambda}$ and ensures
that $g(w)$ is differentiable. The boundedness part of the assumption ensures
that $\frac{\partial}{\partial \bm \lambda} g^{-1}(\bm \lambda^\top {\bf z})$ is
also bounded. Conditions A3 and A4 are standard conditions which control the
first-order and second-order inclusion probability for the second phase sample.
These conditions are met by many traditional survey designs. Assumptions A5 and
A6 are technical conditions and common in the survey literature. Assumption A7
is added to simplify the proof. We could remove this condition and follow the
approach of Theorem 1 in \cite{kwon2024debiased}.

\subsection{Proof of Theorem 1}

In this proof, we derive the solution to \eqref{eq:primal} and show
that it is asymptotically equivalent to a regression estimator. Using the
method of Lagrange multipliers, to solve \eqref{eq:primal} we need to
minimize the Lagrangian in \eqref{eq:legragedc1}. 
The first order conditions show that at the solution,
$$
\frac{\partial \mathcal{L}}{\partial w_{2i}}: g(w_{2i}) d_{1i}q_i -
d_{1i} \hat{\bm \lambda}^\top \bf z_i = 0.
$$
Hence, $\hat w_{2i}(\hat{\bm \lambda}) = 
g^{-1}(\hat{\bm \lambda}^\top \bf z_i)$ where
$\hat{\bm \lambda}$ is determined by \eqref{eq:lamdc1}. When the
sample size gets large, we have $\hat w_{2i}(\hat{\bm \lambda}) \to
d_{2i|1}$ in probability because both terms in \eqref{eq:lamdc1} must
become equal by the Central Limit Theorem. Hence,
$\hat{\bm \lambda} \to \bm \lambda^*$ in probability and 
$\bm \lambda^* = (\bf 0^\top, 1)^\top$. 
Using the linearization technique of
\cite{randles1982asymptotic}, we can construct a regression estimator, 
$$
\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = 
\sum_{i \in A_2} d_{1i} \hat w_{2i} y_i + 
\left(\sum_{i \in A_1} d_{1i} \bm z_i  - \sum_{i \in A_2} d_{1i} 
\hat w_{2i} \bf z_i \right)^\top \bm \phi.
$$
Notice that $\hat Y_\ell(\hat{\bm \lambda}, \bm \phi) = \sum_{i \in A_2} d_{1i}
\hat w_{2i} y_i$ for all $\bm \phi$. We choose $\bm \phi^*$ such that
$$
E\left[\frac{\partial}{\partial \bm \lambda} \hat Y_\ell(\bm \lambda^*,
\bm \phi^*)\right]=\bf{0}.
$$
Using the fact that $g^{-1}((\bm \lambda^*)^\top \bf z_i) = g^{-1}(g(d_{2i|1})) = d_{2i|1}$
and $(g^{-1})'(x) = 1 / g'(g^{-1}(x))$, we have
\begin{align*}
  \bm \phi^*
  &= E\left[\sum_{i \in A_2} \frac{d_{1i}\bf z_i \bf z_i^\top }{g'(d_{2i|1})q_i}
  \right]^{-1}
  E\left[\sum_{i \in A_2} \frac{d_{1i}\bf z_i y_i}{g'(d_{2i|1})q_i}\right]\\
  &= \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i \bf z_i^\top }{g'(d_{2i|1})q_i}
  \right]^{-1}
  \left[\sum_{i \in U} \frac{\pi_{2i|1} \bf z_i y_i}{g'(d_{2i|1})q_i}\right]\\
\end{align*}
Thus, the linearization estimator is
$$\hat Y_\ell(\bm \lambda^*, \bm \phi^*) = 
\sum_{i \in A_1} d_{1i} \bf{z}_i^\top \bm \phi^* +
\sum_{i \in A_2} d_{1i} d_{2i|1} (y_i - \bf{z}_i^\top \bm \phi^*).$$
By construction using a Taylor expansion yields,
\begin{align*}
  \hat Y_{DCE}(\hat{\bm \lambda}) 
  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + 
  E\left[\frac{\partial}{\partial \bm \lambda}\hat Y_\ell(\bm \lambda^*,
  \bm \phi^*)\right](\hat{\bm \lambda} - \bm \lambda^*) + \frac{1}{2}
  E\left[\frac{\partial}{\partial \bm \lambda^2} \hat Y_{DCE}(\bm \lambda^*)\right] 
  (\hat{\bm \lambda} - \bm \lambda^*)^2\\
  &= \hat Y_\ell(\bm \lambda^*, \bm \phi^*) + O(N)O_p(n_2^{-1}).
\end{align*}
The final equality comes from the fact that 
$E\left[\frac{\partial}{\partial \bm \lambda}
\hat Y_\ell(\bm \lambda^*, \bm \phi^*)\right] = 0$,
$\frac{\partial}{\partial \bm \lambda^2} \hat w_{2i}(\bm \lambda^*)$ is
bounded and $||\hat{\bm \lambda} - \bm \lambda^*|| = O_p(n_2^{-1/2})$, which
proves our result.

\subsection{Proof of Theorem 2}

\begin{proof}
  By Theorem~\ref{thm:dc1}, $\hat Y_{\rm DCE}$ is asymptotically equivalent to
  $\hat Y_{\ell}$. Notice that we can define
  $$
  \eta_i^* = 
  \bf z_i \bm \phi^* + \delta_i d_{2i|1}(y_i - \bf{z}_i^\top \bm \phi^*)
  $$
  where $\delta_i$ is one if $i \in A_2$ and zero otherwise. Then,
  $$
  \hat Y_{\ell} = \sum_{i \in A_1} d_{1i} \eta_i^*.
  $$
  Using the reverse framework of \cite{fay1992inferences} and 
  \cite{shao1999variance}, and noting that 
  $\Delta_{1ij} = \pi_{1ij} - \pi_{1i}\pi_{1j}$ and 
  $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1}\pi_{2j|1}$,
  \begin{align*}
    \Var(\hat Y_{\ell, \rm{DCE}}) 
    &= \Var\left(\sum_{i \in A_1} d_{1i} \eta_i^*\right) \\
    &= \Var\left(E\left[\sum_{i \in A_1} d_{1i} \eta_i^* \mid A_2 \right]\right) 
    + E\left[\Var\left(\sum_{i \in A_1} d_{1i} \eta_i^* \mid A_2 \right)\right] \\
    &= \Var\left(\sum_{i \in U} \eta_i^*\right) 
    + E\left[\sum_{i \in U} \sum_{j \in U} \Delta_{1ij} d_{1i} \eta_i^* w_{1j}
      \eta_j^*\right]\\
  \end{align*}
  This means that we can estimate the variance of $\hat Y_{\rm DCE}$ with
  $$
  \hat V_{\rm DCE} 
= \sum_{i \in A_1} \sum_{j \in A_1} d_{1ij} \Delta_{1ij}
  \hat \eta_i \hat \eta_j
+ \sum_{i \in A_2} \sum_{j \in A_2} d_{1i} d_{2ij|1}\Delta_{2ij|1}
  d_{2i|1}(y_i - \bf z_i \hat{\bm \phi})
  d_{2j|1}(y_j - \bf z_j \hat{\bm \phi}).
  $$
\end{proof}

\subsection{Proof of Theorem 4}

\begin{proof}
  From Theorem~\ref{thm:dc2}, we know that $\hat Y_{\rm NNE}(\hat \lambda) = 
  \hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*) + O_p(Nn_2^{-1})$. Hence, the
  variance of $\hat Y_{\rm NNE}(\hat{\bm \lambda})$ is 
  \begin{align*}
    \Var(\hat Y_{\rm NNE}(\hat{\bm \lambda})) 
    &\doteq \Var(\hat Y_{\ell, \rm NNE}(\bm \lambda^*, \bm \phi^*)) \\ 
    &= \Var\left(\sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) y_i + 
      \left(\bf T - \sum_{i \in A_2} \hat w_{2i}(\bm \lambda^*) \bf z_i
    \right)^\top \bm \phi^*\right)\\
    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i} 
    (y_i - \bf{z}_i^\top \bm \phi^*)d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*)\\
    &\qquad+ 2 \Cov\left(\hat{\bf X}_c^\top \bm \phi_1^*, \sum_{i \in A_2} 
    d_{2i}(y_i - \bf{z}_i^\top \bm \phi^*)\right) \\
    &= (\bm \phi_1^*)^\top \Var(\hat{\bf X}_c) \bm \phi_1^* + 
    \sum_{i \in U} \sum_{j \in U} \Delta_{2ij}d_{2i}
    (y_i - \bf{z}_i^\top \bm \phi^*)d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*)\\
    &\qquad + (1 - W)\left(\sum_{i \in U} \sum_{j \in U} \Delta_{2ij} 
    d_{2i}\bf{x}_i d_{2j}(y_j - \bf{z}_j^\top \bm \phi^*_1)\right)^\top \bm \phi_1^* \\
  \end{align*}
where the last equality comes from the fact that $\hat{\bf X}_c = W\hat{\bf
X}_1 + (1 - W) \hat{\bf X}_2$. To have an unbiased estimator of the variance
  we can use:
  \begin{align*}
    \hat V_{\rm NNE} 
    &= (\hat{\bm \phi_1})^\top \Var(\hat{\bf X_c}) \hat{\bm \phi_1} + 
    \sum_{i \in A_2} \sum_{j \in A_2} d_{2ij}\Delta_{2ij}d_{2i}
    (y_i - \bf{z}^\top_i \hat{\bm \phi})d_{2j}(y_j - \bf{z}^\top_j \hat{\bm \phi}) \\
    &\qquad + (1 - W)\left(\sum_{i \in A_2} \sum_{j \in A_2} 
      d_{2ij} \Delta_{2ij} d_{2i}\bf{x}_i
    d_{2j}(y_j - \bf{z}^\top_j \hat{\bm \phi_1})\right)^\top \hat{\bm \phi}_1 
  \end{align*}
  where 
  $$\hat{\bm \phi} =
  \begin{bmatrix}
    \hat{\bm \phi_1} \\ \hat \phi_2
  \end{bmatrix} = 
  \left(\sum_{i \in A_2} \frac{1}{g'(d_{2i})q_i} 
  \begin{bmatrix}
    \bf x_i^2 & \bf x_i g(d_{2i}) q_i \\
    \bf x_i g(d_{2i}) q_i & g(d_{2i})^2 q_i^2
  \end{bmatrix}
  \right)^{-1}
  \sum_{i \in A_2} \frac{y_i}{g'(d_{2i})q_i} 
  \begin{bmatrix} \bf x_i \\ g(d_i) q_i \end{bmatrix}.
  $$

\end{proof}

\section{Estimating the Population Weights}

In \eqref{eq:primal}, \eqref{eq:nnopt}, and
\eqref{eq:msdebconstr}, we see the need for knowing $\sum_{i \in U}
g(d_i)$. (In the case of \eqref{eq:primal}, we only need $\sum_{i \in
A_1} g(d_{2i|1})$.) However, this quantity is not always known. When it is
unknown, we need to estimate it.
We follow the approach of \cite{kwon2024debiased}. Define 
$$N \alpha = \sum_{i \in U} g(d_i).$$
Assuming $N$ is known (or that we can estimate it), we need to estimate
$\alpha$. Since our estimating equation is convex for any fixed $\alpha$, we
adopt a two-step procedure for which we estimate $\alpha$ and $\bf w$. In this
approach we have the following loss function for the case of non-nested sampling:
\begin{align*}
  (\hat \alpha, \hat w_2) 
  &= 
  \argmin_{\alpha} \argmin_{w_2} \sum_{i \in A_2} \{G(w_{2i})\} - NK(\alpha), \\
  &\text{ such that } \sum_{i \in A_2} w_{2i} \bf x_i = \hat{\bf X}_c,
  \text{ and } \sum_{i \in A_2} w_{2i} g(d_{2i}) = N\alpha.
\end{align*}
The choice of $K(\alpha)$ is up to the analyst. We consider $K(\alpha) =
\alpha$.

\end{document}